---
title: 'Liver - gene network metadata association analysis in Moylan 2013 microarray data (new)'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
  pAdjMethod: "BH"
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

previous: liver_module_embed_in_moylan2013.Rmd

# Setup

## Load packages

```{r}
#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr"))
#library("Seurat")

library("data.table")
library("Matrix")
library("parallel")
#library("tidyverse")
library("here")
# library("boot")
library("magrittr")
# library("AUCell")
# library("car")
library("boot")
# source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))

# Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")

R=1e3-1
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

# Set constants

randomSeed = 12345
set.seed(randomSeed)
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_renamed"
colCellCluster = "cell_cluster_pres"
# naming convention:
# <prefixData>_<prefixRun>_<prefixOut>_<outputType>_<params>_<params>_..._<date>.pdf

```

load embeddings
this includes sc and bulk modules

```{r}
# df_embed = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mod_embed_gerhard2018.csv.gz") %>% setDF
df_embed = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mat_scaled_embed_moylan2013.csv.gz") %>% setDF

mat_scaled_embed = df_embed[,-1] %>% as.matrix
rownames(mat_scaled_embed) = df_embed$GEO_accession
mat_scaled_embed[0:4,0:4]
```

<!-- compare distribution of module activity to gerhard -->

<!-- ```{r} -->
<!-- df_embed_g = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mat_scaled_embed_gerhard2018.csv.gz") %>% setDF -->
<!-- ``` -->

<!-- ```{r} -->
<!-- mat_embed_g = df_embed_g[,-1] %>% as.matrix -->
<!-- rownames(mat_embed_g) = df_embed_g$run_accession -->
<!-- mat_embed_g[0:4,0:4] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- hist(mat_embed[,"Cholangiocytes_3"], breaks = 20) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- hist(mat_embed_g[,"Cholangiocytes_3"], breaks = 20) -->
<!-- ``` -->

## prepare table for linear model

lm can't deal with dashes in module names so replace with underscore

```{r}
colnames(mat_scaled_embed) <- gsub("-","_",colnames(mat_scaled_embed))
```

```{r}
dt_metadata <- fread(here("data", "moylan2013.metadata.csv"))
```

transform fibrosis to indicator variable

```{r}
dt_metadata[,condition:=factor(ifelse(fibrosis=="advanced",1,0)),]
```

<!-- load data.table with information on module and cell type of origin -->

<!-- ```{r} -->
<!-- path_geneMod_perslab <- paste0(dirWGCNA_tables,prefixData, "_",prefixRun_WGCNA,"_geneMod.csv.gz") -->

<!-- dt_geneMod_perslab <- fread(path_geneMod_perslab) -->

<!-- #head(dt_geneMod_perslab) -->

<!-- path_geneMod_gerhard2018 <- paste0(dirWGCNA_tables,"liver_gerhard2018_wgcna3_geneMod.csv.gz") -->

<!-- dt_geneMod_gerhard2018 <- fread(path_geneMod_gerhard2018) -->

<!-- #head(dt_geneMod_gerhard2018) -->

<!-- path_geneMod_moylan2013 <- paste0(dirWGCNA_tables,"liver_moylan2013_wgcna1_geneMod.csv.gz") -->

<!-- dt_geneMod_moylan2013 <- fread(path_geneMod_moylan2013) -->

<!-- #head(dt_geneMod_moylan2013) -->
<!-- ``` -->

# analysis

# Are any modules zero in all samples (because the genes aren't found in the bulk dataset) ?

<!-- ```{r} -->
<!-- vec_allZero <- apply(X=mat_embed, MARGIN=2, FUN = function(x) all(x==0)) -->
<!-- table(vec_allZero) -->
<!-- # FALSE  -->
<!-- #   129 -->
<!-- ``` -->

<!-- ```{r} -->

<!-- if (any(vec_allZero)) { -->
<!--   # get rid of all-zero columns, i.e. modules where none of the genes were matched in the bulk expression data -->
<!--   mat_embed <- mat_embed[,!apply(X=mat_embed, MARGIN=2, FUN = function(x) all(x==0))] -->

<!--   dim(mat_embed) -->
<!-- } -->
<!-- ``` -->


## run linear models

### all modules

### combine embeddings and metadata

run linear models 

```{r}
dt_model_tmp <- data.table(dt_metadata, mat_scaled_embed[dt_metadata$geo_accession,])
```

```{r}
set.seed(randomSeed)

lapply(colnames(mat_scaled_embed), function(module) {
  #formula_tmp <- formula(paste0(module, "~ condition"))
  ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp)
  #regr<-lm(formula = formula_tmp, dt_model_tmp)
  #bootOut = "error"
  #class(bootOut) = "try-error"
  if (all(is.na(mat_scaled_embed[,module]))) {
    return(NA) 
  } else {
    fnc_stat = function(data,indices) {
       lm(
         data = data[indices,], 
         formula = formula(paste0(module, "~ condition"))) %>% coefficients()
    }
    bootOut = "error"
    class(bootOut) = "try-error"  
    while (class(bootOut)=="try-error") {
       bootOut = try({
          set.seed(randomSeed+1)
          bootOut <- boot::boot(data=dt_model_tmp, 
                                 statistic = fnc_stat,
                                 R = R,
                                 sim = "ordinary",
                                 stype = "i",
                                 parallel = getOption("boot.parallel"),
                                 ncpus = getOption("boot.ncpus")) 
        })
      
    }
    return(bootOut)
  }
}) -> list_bootOut 

names(list_bootOut) = colnames(mat_scaled_embed)
```

extract coefficient tables 

get p-value, std error and confidence intervals from the $t

```{r}
mat_bootOut_stats <- t(sapply(names(list_bootOut), function(module) {
  bootOut = list_bootOut[[module]]
  vec_out = tryCatch({
    mat_boot_replicates = bootOut$t
    colnames(mat_boot_replicates) = names(bootOut$t0)
    idx_col = grep("condition",names(bootOut$t0))
    # compute bootstrap statistics
    bootBias = mean(mat_boot_replicates[,idx_col]) - bootOut$t0[idx_col]
    bootMean = mean(mat_boot_replicates[,idx_col]) 
    bootSE = sd(mat_boot_replicates[,idx_col])
    bootMed = median(mat_boot_replicates[,idx_col])
    # compute bootstrap percentile intervals 
    bootciOut = boot.ci(bootOut, index = 2)
    confInt_lower_percentile = bootciOut$percent[4]
    confInt_upper_percentile = bootciOut$percent[5]
    # compute Normal-theory intervals
    confInt_lower_norm = bootciOut$normal[2]
    confInt_upper_norm = bootciOut$normal[3]
    
    c("estimate_original"=bootOut$t0[idx_col],
     "bootSE"=bootSE,
     "estimate_bootMean"=bootMean,
     "estimate_bootMed" = bootMed,
     "confInt_lower_percentile"=confInt_lower_percentile,
     "confInt_upper_percentile"=confInt_upper_percentile,
     "confInt_lower_norm"=confInt_lower_norm,
     "confInt_upper_norm"=confInt_upper_norm,
      "p.value_percentile" = max(0.5,min(sum(mat_boot_replicates[,idx_col]>0), sum(mat_boot_replicates[,idx_col]<0)))*2/R,
     "p.value_norm" = pnorm(q=abs(bootMean), mean=0, sd=bootSE, lower.tail = F)*2)
  }, error = function(err) {
    warning(paste0(module, ": ", err))
    c("estimate_original"=NA_real_,
     "bootSE"=NA_real_,
     "estimate_bootMean"=NA_real_,
     "estimate_bootMed" = NA_real_,
     "confInt_lower_percentile"=NA_real_,
     "confInt_upper_percentile"=NA_real_,
     "confInt_lower_norm"=NA_real_,
     "confInt_upper_norm"=NA_real_,
      "p.value_percentile" = NA_real_,
     "p.value_norm"=NA_real_)
    })
  
}))
dt_bootOut_stats = data.table("condition"="fibrosis_severity", 
                              "module"=names(list_bootOut), 
                              mat_bootOut_stats)

colnames(dt_bootOut_stats)[colnames(dt_bootOut_stats)=="estimate_original.condition1"] <- "estimate_original" # for some reason, it appends ".condition" to these column names
```

<!-- ```{r} -->
<!-- sapply(list_bootOut, is.na) %>% sum -->
<!-- ``` -->

<!-- check the residuals .  -->

<!-- ```{r} -->
<!-- # n is the same within a condition -->
<!-- n <- length(list_regr[[1]]$residuals) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- regr <- list_regr[[69]] -->

<!-- stats::qqplot(y = summary(regr)$residuals/summary(regr)$sigma, x=qnorm(p=seq(from=0.01, to=1, by=1/n)), xlim=c(-3,3), ylim=c(-3,3)) -->

<!-- abline(coef=c(0,1)) -->
<!-- ``` -->


<!-- get coefficient matrices -->

<!-- ```{r} -->
<!-- lapply(list_regr, function(regr) { -->
<!--   regr %>% broom::tidy()  -->
<!-- }) -> list_tbl_tmp  -->

<!-- names(list_tbl_tmp) = colnames(mat_scaled_embed) -->
<!-- ``` -->

<!-- extract coefficients for the condition regressors  -->

<!-- ```{r} -->

<!-- mat_out <- sapply(list_tbl_tmp, function(tbl_regr) { -->
<!--     tbl_regr[tbl_regr$term=="condition1",colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")]  -->
<!--   }) %>% t -->

<!-- dt_condition_coef <- data.table("module"= rownames(mat_out), mat_out)  -->

<!-- ``` -->

<!-- ```{r} -->
<!-- dt_condition_coef[[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=dt_condition_coef$p.value,method=params$pAdjMethod) -->
<!-- ``` -->

<!-- how well do the betas correlate across modules with those for fibrosis in Gerhard? -->

<!-- ```{r} -->
<!-- dt_fibrosis_coef_gerhard2018 <- fread(paste0(dirWGCNA_tables, "liver_perslab_int_fibrosis_assoc_in_gerhard2018_201008.csv")) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- cor(dt_condition_coef[,estimate] %>% as.numeric, as.numeric(dt_fibrosis_coef_gerhard2018$estimate)) -->
<!-- # [1] 0.4491132 -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_fibrosis_severity_coef_gerhard2018 <- fread(paste0(dirWGCNA_tables, "liver_perslab_int_sc_modules_log2_fibrosis_severity_assoc_in_gerhard2018_201008.csv")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- condition = quote(!grepl("moylan|gerhard", dt_condition_coef$module)) -->
<!-- dt_condition_coef_perslab = dt_condition_coef[eval(condition)] -->
<!-- dt_fibrosis_severity_coef_gerhard2018_perslab = dt_fibrosis_severity_coef_gerhard2018[eval(condition)] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_condition_coef_perslab[,estimate] %>% as.numeric, as.numeric(dt_fibrosis_severity_coef_gerhard2018_perslab$estimate)) -->
<!-- # [1] 0.4491132 -->
<!-- ``` -->

<!-- check correlation of module betas for fibrosis in Moylan with module betas for lobular inflammation in gerhard 2018 -->

<!-- ```{r} -->
<!-- dt_lobinflam_coef_gerhard2018 <- fread(paste0(dirWGCNA_tables, "liver_perslab_int_sc_bulk_modules_scaled_lob_inflam_association_in_gerhard2018_200511.csv")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_condition_coef[,estimate] %>% as.numeric, as.numeric(dt_lobinflam_coef_gerhard2018$estimate)) -->
<!-- # [1] 0.386009 -->
<!-- ``` -->

<!-- check correlation between lob inflam and fibrosis betas in Gerhard 2018 -->

<!-- ```{r} -->
<!-- cor(dt_fibrosis_coef_gerhard2018[,estimate] %>% as.numeric, as.numeric(dt_lobinflam_coef_gerhard2018$estimate)) -->
<!-- # [1] 0.7230158 -->
<!-- ``` -->


## write out gene module assocation results

```{r}

fwrite(x=dt_bootOut_stats, file=paste0(dirWGCNA_tables, prefixData, "_mod_fibrosis_assoc_in_moylan2013_", flagDate,".csv"))

openxlsx::write.xlsx(x=dt_bootOut_stats,file = paste0(dirWGCNA_tables, prefixData, "_mod_fibrosis_assoc_in_moylan2013_", flagDate,".xlsx"))

```

# look at individual gene expression 

```{r}
vec_genes = c("S100A4", "S100A6", "S100A8",  "S100A9", "S100A10", "S100A12", "VCAN")
```

```{r}
condition = quote(match(vec_genes, dt_datExpr$gene))
 
mat_datExpr_sub = dt_datExpr[eval(condition)] %>% dplyr::select(!gene) %>% t  %>% scale

colnames(mat_datExpr_sub) = vec_genes
```

```{r}
dt_data_genes <- data.table(dt_metadata, mat_datExpr_sub)
```

```{r}
lapply(vec_genes, function(gene) {
  formula_tmp <- formula(paste0(gene, "~ fibrosis"))
  ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp)
  regr<-lm(formula = formula_tmp, dt_data_genes)
  regr %>% broom::tidy() 
}) -> list_tbl_regr

names(list_tbl_regr) <- vec_genes
```

extract coefficients for the regressors 

```{r}
mat_out <- sapply(list_tbl_regr, function(tbl_regr) {
    tbl_regr[tbl_regr$term=="fibrosis",colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")] 
  }) %>% t

dt_coef = data.table("gene"= rownames(mat_out), "condition"="fibrosis", mat_out)
```

```{r}

fwrite(x=dt_coef, file=paste0(dirWGCNA_tables, prefixData, "_", prefixRun, "_S100_genes_assoc_in_moylan2013_scaled_", flagDate,".csv"))
       
openxlsx::write.xlsx(x=list_dt_condition_coef[[name]],file = paste0(dirWGCNA_tables, prefixData, "_sc_bulk_modules_scaled_", name, "_association_in_gerhard2018_", flagDate,".xlsx"))


```

next: 

liver_module_sc_merge_to_metamodules.Rmd
liver_module_plots.Rmd
liver_module_GERR.Rmd
liver_module_gprofiler2.Rmd
