---
title: 'Liver - gene network metadata association analysis in Gerhard 2018 bulk data'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
  pAdjMethod: "BH"
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

Previous: liver_module_embed_in_gerhard2018.Rmd

# Setup

## Load packages

```{r}
#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr"))
#library("Seurat")

library("data.table")
library("Matrix")
library("parallel")
#library("tidyverse")
library("here")
library("boot")
library("magrittr")
#library("AUCell")
library("ggplot2")
#library("simpleboot")
library("boot")
#library("car")#https://socialsciences.mcmaster.ca/jfox/Books/Companion/appendices/Appendix-Bootstrapping.pdf
# source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))

# Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")

R=1e3-1
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

# Set constants


randomSeed = 12345
set.seed(randomSeed)
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 


prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_renamed"
colCellCluster = "cell_cluster_pres"
# naming convention:
# <prefixData>_<prefixRun>_<prefixOut>_<outputType>_<params>_<params>_..._<date>.pdf

```

## load data

<!-- ```{r} -->
<!-- dt_datExpr <- fread(here("data", "liver_gerhard2018_norm.counts.csv.gz")) -->
<!-- ``` -->
### patient metadata

```{r}
dt_metadata <- fread(here("data", "liver_gerhard2018_patient_metadata_qc.csv.gz"))
```

### embeddings

this includes sc and bulk modules

```{r}
# df_embed = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mod_embed_gerhard2018.csv.gz") %>% setDF
df_scaled_embed = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mat_scaled_embed_gerhard2018.csv.gz") %>% setDF

mat_scaled_embed = df_scaled_embed[,-1] %>% as.matrix
rownames(mat_scaled_embed) = df_scaled_embed$run_accession
mat_scaled_embed[0:4,0:4]
```

<!-- ```{r} -->
<!-- # also load unscaled, for comparison -->
<!-- df_unscaled_embed = fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mat_unscaled_embed_gerhard2018.csv.gz") %>% setDF -->

<!-- mat_unscaled_embed = df_unscaled_embed[,-1] %>% as.matrix -->
<!-- rownames(mat_unscaled_embed) = df_unscaled_embed$run_accession -->
<!-- mat_unscaled_embed[0:4,0:4] -->
<!-- ``` -->

<!-- load data.table with information on module and cell type of origin -->

<!-- ```{r} -->
<!-- path_geneMod_perslab <- paste0(dirWGCNA_tables,prefixData, "_",prefixRun_WGCNA,"_geneMod.csv.gz") -->
<!-- dt_geneMod_perslab <- fread(path_geneMod_perslab) -->
<!-- #head(dt_geneMod_perslab) -->

<!-- path_geneMod_gerhard2018 <- paste0(dirWGCNA_tables,"liver_gerhard2018_wgcna3_geneMod.csv.gz") -->
<!-- dt_geneMod_gerhard2018 <- fread(path_geneMod_gerhard2018) -->
<!-- #head(dt_geneMod_gerhard2018) -->

<!-- path_geneMod_moylan2013 <- paste0(dirWGCNA_tables,"liver_moylan2013_wgcna1_geneMod.csv.gz") -->
<!-- dt_geneMod_moylan2013 <- fread(path_geneMod_moylan2013) -->
<!-- #head(dt_geneMod_moylan2013) -->
<!-- ``` -->

# analysis

subset metadata and transform variables to 0,1,2

```{r}
table(dt_metadata$Diagnosis)
# Fibrosis 3 Fibrosis 3/4   Fibrosis 4 Lob Inflam 1 Lob Inflam 2       NORMAL  STEATOSIS 2  STEATOSIS 3 
#           19           12           28           38           13           24           34           16
```

```{r}
dt_metadata_steatosis = dt_metadata[Diagnosis %in% c("NORMAL", "STEATOSIS 2", "STEATOSIS 3"),]
dt_metadata_steatosis[,condition:=ifelse(Diagnosis=="NORMAL",0,1)]
dt_metadata_steatosis[Diagnosis=="STEATOSIS 3","condition"] <- 2

dt_metadata_lobinflam = dt_metadata[Diagnosis %in% c("NORMAL","Lob Inflam 1", "Lob Inflam 2"),]
dt_metadata_lobinflam[,condition:=ifelse(Diagnosis=="NORMAL",0,1)]
dt_metadata_lobinflam[Diagnosis=="Lob Inflam 2","condition"] <- 2

dt_metadata_fibrosis = dt_metadata[Diagnosis %in% c("NORMAL","Fibrosis 3"),]
dt_metadata_fibrosis[,condition:=ifelse(Diagnosis=="Fibrosis 3",1,0),]
# binary, so factor. NB: This will change the lm variable for case  to condition1
dt_metadata_fibrosis$condition <- as.factor(dt_metadata_fibrosis$condition)

dt_metadata_lobinflam_vs_fibrosis = dt_metadata[Diagnosis %in% c("Lob Inflam 1", "Lob Inflam 2","Fibrosis 3"),]
dt_metadata_lobinflam_vs_fibrosis[,condition:=ifelse(Diagnosis=="Fibrosis 3",1,0),]
# binary, so factor. NB: This will change the lm variable for case  to condition1
dt_metadata_lobinflam_vs_fibrosis$condition <- as.factor(dt_metadata_lobinflam_vs_fibrosis$condition)

dt_metadata_fibrosis_severity = dt_metadata[grepl("Fibrosis", Diagnosis),]
dt_metadata_fibrosis_severity[,condition:=ifelse(Diagnosis=="Fibrosis 3",0,1),]
# binary, so factor. NB: This will change the lm variable for case  to condition1
dt_metadata_fibrosis_severity$condition <- as.factor(dt_metadata_fibrosis_severity$condition)
```

<!-- # Are any modules zero in all samples (because the genes aren't found in the bulk dataset) ?  -->

<!-- ```{r} -->
<!-- vec_allZero <- apply(X=mat_scaled_embed, MARGIN=2, FUN = function(x) all(x==0)) -->
<!-- table(vec_allZero) -->
<!-- # FALSE  -->
<!-- #   107  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- if (any(vec_allZero)) { -->
<!--   # get rid of all-zero columns, i.e. modules where none of the genes were matched in the bulk expression data -->
<!--   mat_scaled_embed <- mat_scaled_embed[,!apply(X=mat_scaled_embed, MARGIN=2, FUN = function(x) all(x==0))] -->

<!--   dim(mat_scaled_embed) -->
<!-- } -->
<!-- ``` -->

### combine embeddings and metadata

lm can't deal with dashes in module names so replace with underscore

```{r}
colnames(mat_scaled_embed) <- gsub("-","_",colnames(mat_scaled_embed))
```

<!-- ```{r} -->
<!-- colnames(mat_unscaled_embed) <- gsub("-","_",colnames(mat_unscaled_embed)) -->
<!-- ``` -->

```{r}
list_dt_metadata = list(dt_metadata_steatosis, dt_metadata_lobinflam, dt_metadata_fibrosis, dt_metadata_lobinflam_vs_fibrosis, dt_metadata_fibrosis_severity)

names(list_dt_metadata) <- c("steatosis", "lob_inflam", "fibrosis", "fibrosis_vs_lob_inflam", "fibrosis_severity")
```

```{r}
list_dt_model_sub <- lapply(names(list_dt_metadata), function(condition) {
  dt_metadata_tmp = list_dt_metadata[[condition]]
  
  dt_tmp = data.frame(dt_metadata_tmp, mat_scaled_embed[ dt_metadata_tmp$run_accession,])
  # if (condition %in% c("steatosis", "lob_inflam")) {
  #   dt_tmp$condition <- as.numeric(dt_tmp$condition)
  # } else {
  #   dt_tmp$condition <- as.factor(dt_tmp$condition)
  # }
  # dt_tmp[,SEX := dt_metadata_tmp$SEX]
  # dt_tmp[,BMI_surg := dt_metadata_tmp$BMI_surg]
  # dt_tmp[,Age := dt_metadata_tmp$Age]
  # dt_tmp[,condition := dt_metadata_tmp$condition]
  #tmp = quote(colnames(mat_scaled_embed))
  #dt_tmp[, eval(tmp)] <- dt_tmp[, lapply(.SD, as.numeric)] #.SDcols=colnames(mat_scaled_embed)]
  return(dt_tmp)
})
names(list_dt_model_sub) <- names(list_dt_metadata)
```

<!-- ```{r} -->
<!-- list_dt_model_sub_unscaled <- lapply(list_dt_metadata, function(dt_sub) { -->
<!--   data.table(dt_sub, mat_unscaled_embed[dt_sub$run_accession,]) -->
<!-- }) -->
<!-- names(list_dt_model_sub_unscaled) <- names(list_dt_metadata) -->
<!-- ``` -->

plot the module expression by condition (not taking into account sex, age, BMI)

```{r}
dev.off()
```

```{r}
ggplot(data=list_dt_model_sub[["fibrosis"]], mapping = aes(x=data, y=Hepatocytes_4, fill=condition)) + 
  geom_violin(position = position_dodge(width = 1)) + geom_jitter(position = position_jitterdodge(seed = 1, dodge.width = 1))
  
```

```{r}
ggplot(data=list_dt_model_sub[["fibrosis"]], mapping = aes(x=data, y=Hepatocytes_8, fill=condition)) + 
 geom_violin(position = position_dodge(width = 1)) + geom_jitter(position = position_jitterdodge(seed = 1, dodge.width = 1))

```

run linear models 

scaled 

```{r}
set.seed(randomSeed)

system.time(list_list_bootOut <- lapply(names(list_dt_model_sub), function(condition) {
  print(paste0("condition: ", condition))
  
  dt_model_tmp <- setDF(list_dt_model_sub[[condition]])
  #dt_model_tmp$SEX <- factor(dt_model_tmp$SEX)
  lapply(colnames(mat_scaled_embed), function(module) {
    
    print(paste0("module: ", module))
    
    if (all(is.na(mat_scaled_embed[,module]))) {
      return(NA) 
    } else {
    fnc_stat = function(data,indices) {
         lm(
           data = data[indices,], 
           formula = formula(paste0(module, "~ SEX + BMI_surg + Age + condition"))) %>% coefficients()
    }
    bootOut = "error"
    class(bootOut) = "try-error"
    while (class(bootOut)=="try-error") {
      boot_out = try({
        set.seed(randomSeed+1)
        bootOut <- boot::boot(data=dt_model_tmp, 
                               statistic = fnc_stat,
                               R = R,
                               sim = "ordinary",
                               stype = "i",
                               parallel = getOption("boot.parallel"),
                               ncpus = getOption("boot.ncpus")) 
      })
    }
    return(bootOut)  
    }
  }) -> list_bootOut 
  names(list_bootOut) = colnames(mat_scaled_embed)
  return(list_bootOut)
  })
)

names(list_list_bootOut) <- names(list_dt_model_sub)
```

<!-- look at cook's distances  -->

<!-- ```{r} -->
<!-- regr <- list_list_regr[["fibrosis"]][[1]] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- regr_inflm = influence.measures(model = regr) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- summary(regr_inflm) -->
<!-- ``` -->

<!-- look at the studentized model residuals (not omitting each observation) -->

<!-- ```{r} -->

<!-- # n is the same within a condition -->
<!-- n <- length(list_list_regr[["fibrosis"]][[1]]$residuals) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- vec_sigma_del = sapply(1:n, function(i) { -->
<!--   regr$fitted.values -->
<!-- }) -->

<!--   summary(regr)$residuals/summary(regr)$sigma -->

<!-- stats::qqplot(y = , x=qnorm(p=seq(from=0.01, to=1, by=1/n)), xlim=c(-3,3), ylim=c(-3,3)) -->

<!-- abline(coef=c(0,1)) -->
<!-- ``` -->

<!-- uh uh. For the scaled (not log) embeddings we see an inverse sigmoid (overshooting Normality at low and high quantiles, suggesting that the model is skewed by a subset) -->

extract coefficient tables 

get p-value, std error and confidence intervals from the $t

```{r}
list_dt_bootOut_stats <- lapply(names(list_list_bootOut), function(condition) {
  list_bootOut = list_list_bootOut[[condition]]
  mat_bootOut_stats <- t(sapply(names(list_bootOut), function(module) {
    bootOut = list_bootOut[[module]]
    vec_out = tryCatch({
      mat_boot_replicates = bootOut$t
      colnames(mat_boot_replicates) = names(bootOut$t0)
      idx_col = grep("condition",names(bootOut$t0))
      # compute bootstrap statistics
      bootBias = mean(mat_boot_replicates[,idx_col]) - bootOut$t0[idx_col]
      bootMean = mean(mat_boot_replicates[,idx_col]) 
      bootSE = sd(mat_boot_replicates[,idx_col])
      bootMed = median(mat_boot_replicates[,idx_col])
      
      # compute bootstrap percentile intervals 
      bootciOut = boot.ci(bootOut, index = 5)
      # idx_lower = round((R+1)*0.025,0)
      # idx_upper = round((R+1)*0.975,0)
      # confInt_lower_boot = sort(mat_boot_replicates[,idx_col])[idx_lower]
      # confInt_upper_boot = sort(mat_boot_replicates[,idx_col])[idx_upper]
      confInt_lower_percentile = bootciOut$percent[4]
      confInt_upper_percentile = bootciOut$percent[5]
      # compute Normal-theory intervals
      # confInt_lower_norm = qnorm(p = 0.025, mean = bootMean, sd = bootSE, lower.tail = T)
      # confInt_upper_norm = qnorm(p = 0.975, mean = bootMean, sd = bootSE, lower.tail = T)
      confInt_lower_norm = bootciOut$normal[2]
      confInt_upper_norm = bootciOut$normal[3]
      
      c("estimate_original"=bootOut$t0[idx_col],
       #"bootBias"=bootBias,
       "bootSE"=bootSE,
       "estimate_bootMean"=bootMean,
       "estimate_bootMed" = bootMed,
       "confInt_lower_percentile"=confInt_lower_percentile,
       "confInt_upper_percentile"=confInt_upper_percentile,
       "confInt_lower_norm"=confInt_lower_norm,
       "confInt_upper_norm"=confInt_upper_norm,
       "p.value_percentile" = max(0.5,min(sum(mat_boot_replicates[,idx_col]>0), sum(mat_boot_replicates[,idx_col]<0)))*2/R,
       "p.value_norm" = pnorm(q=abs(bootMean), mean=0, sd=bootSE, lower.tail = F)*2)
    }, error = function(err) {
      warning(paste0(module, ": ", err))
      c("estimate_original"=NA_real_,
       #"bootBias"=NA_real_,
       "bootSE"=NA_real_,
       "estimate_bootMean"=NA_real_,
       "estimate_bootMed" = NA_real_,
       "confInt_lower_percentile"=NA_real_,
       "confInt_upper_percentile"=NA_real_,
       "confInt_lower_norm"=NA_real_,
       "confInt_upper_norm"=NA_real_,
        "p.value_percentile" = NA_real_,
       "p.value_norm"=NA_real_)
      })
    
  }))
  dt_bootOut_stats = data.table("condition"=condition, 
                                "module"=names(list_bootOut), 
                                mat_bootOut_stats)
  
  #dt_bootOut_stats[,p.value_percent_BH:=p.adjust(p.value_percent,method = params$pAdjMethod)]
  #dt_bootOut_stats[,p.value_born_BH:=p.adjust(p.value_norm, method = params$pAdjMethod)]
  colnames(dt_bootOut_stats)[3]="estimate_original" # for some reason, it appends ".condition" to these column names
  return(dt_bootOut_stats)
})

names(list_dt_bootOut_stats) <- names(list_dt_model_sub)

dt_bootOut_stats = data.table::rbindlist(list_dt_bootOut_stats)

#dt_bootOut_stats[,module := gsub("_","-", module)]
```

<!-- ```{r} -->
<!-- list_list_tbl_regr <- lapply(list_list_regr, function(list_regr) { -->
<!--   lapply(list_regr, function(regr) { -->
<!--     regr %>% broom::tidy()  -->
<!--   }) -> list_tbl_tmp  -->
<!--   names(list_tbl_tmp) = colnames(mat_scaled_embed) -->
<!--   return(list_tbl_tmp) -->
<!-- }) -->

<!-- names(list_list_tbl_regr) <- names(list_dt_model_sub) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- lapply(names(list_list_tbl_regr), function(condition) { -->
<!--   list_tbl_regr = list_list_tbl_regr[[condition]] -->
<!--   mat_out <- sapply(list_tbl_regr, function(tbl_regr) { -->
<!--     tbl_regr[grep("condition",tbl_regr$term),colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")]  -->
<!--   }) %>% t -->
<!--   data.table("module"= rownames(mat_out), "condition"=condition, mat_out) -->
<!-- }) -> list_dt_condition_coef -->

<!-- names(list_dt_condition_coef) <- names(list_list_tbl_regr) -->
<!-- ``` -->


<!-- ```{r} -->

<!-- lapply(names(list_list_boot_summary), function(condition) { -->
<!--   list_boot_summary = list_list_boot_summary[[condition]] -->
<!--   mat_out <- sapply(list_boot_summary, function(boot_summary) { -->
<!--     boot_summary[grep("condition",tbl_regr$term),colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")]  -->
<!--   }) %>% t -->
<!--   data.table("module"= rownames(mat_out), "condition"=condition, mat_out) -->
<!-- }) -> list_dt_condition_coef -->

<!-- names(list_dt_condition_coef) <- names(list_list_tbl_regr) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- #adjust p-values -->
<!-- list_dt_condition_coef[[1]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef[[1]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef[[2]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef[[2]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef[[3]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef[[3]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef[[4]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef[[4]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef[[5]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef[[5]]$p.value,method=params$pAdjMethod) -->
<!-- ``` -->

<!-- examine the model residuals -->




<!-- unscaled  -->

<!-- ```{r} -->
<!-- list_list_tbl_regr_unscaled <- lapply(names(list_dt_model_sub_unscaled), function(outcome) { -->
<!--   dt_model_tmp <- list_dt_model_sub_unscaled[[outcome]] -->
<!--   dt_model_tmp$SEX <- factor(dt_model_tmp$SEX) -->
<!--   lapply(colnames(mat_unscaled_embed), function(module) { -->
<!--     formula_tmp <- formula(paste0(module, "~ SEX + BMI_surg + Age + condition")) -->
<!--     ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp) -->
<!--     regr<-lm(formula = formula_tmp, dt_model_tmp) -->
<!--     regr %>% broom::tidy()  -->
<!--   }) -> list_tbl_tmp  -->
<!--   names(list_tbl_tmp) = colnames(mat_unscaled_embed) -->
<!--   return(list_tbl_tmp) -->
<!-- }) -->

<!-- names(list_list_tbl_regr_unscaled) <- names(list_dt_model_sub_unscaled) -->

<!-- #extract coefficients for the condition regressors  -->

<!-- lapply(names(list_list_tbl_regr_unscaled), function(condition) { -->
<!--   list_tbl_regr = list_list_tbl_regr_unscaled[[condition]] -->
<!--   mat_out <- sapply(list_tbl_regr, function(tbl_regr) { -->
<!--     tbl_regr[grep("condition",tbl_regr$term),colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")]  -->
<!--   }) %>% t -->
<!--   data.table("module"= rownames(mat_out), "condition"=condition, mat_out) -->
<!-- }) -> list_dt_condition_coef_unscaled -->

<!-- names(list_dt_condition_coef_unscaled) <- names(list_list_tbl_regr_unscaled) -->

<!-- #adjust p-values -->
<!-- list_dt_condition_coef_unscaled[[1]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef_unscaled[[1]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef_unscaled[[2]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef_unscaled[[2]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef_unscaled[[3]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef_unscaled[[3]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef_unscaled[[4]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef_unscaled[[4]]$p.value,method=params$pAdjMethod) -->

<!-- list_dt_condition_coef_unscaled[[5]][[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=list_dt_condition_coef_unscaled[[5]]$p.value,method=params$pAdjMethod) -->
<!-- ``` -->

<!-- look at results -->

<!-- scaled -->

<!-- ```{r} -->
<!-- sum(list_dt_condition_coef[["steatosis"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 0 -->
<!-- sum(list_dt_condition_coef[["lob_inflam"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 78 -->
<!-- sum(list_dt_condition_coef[["fibrosis"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 69 -->
<!-- sum(list_dt_condition_coef[["fibrosis_vs_lob_inflam"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- # [1] 30 -->
<!-- sum(list_dt_condition_coef[["fibrosis_severity"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 0 -->
<!-- ``` -->

<!-- unscaled -->

<!-- ```{r} -->
<!-- sum(list_dt_condition_coef_unscaled[["steatosis"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 0 -->
<!-- sum(list_dt_condition_coef_unscaled[["lob_inflam"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 83 -->
<!-- sum(list_dt_condition_coef_unscaled[["fibrosis"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 77 -->
<!-- sum(list_dt_condition_coef_unscaled[["fibrosis_vs_lob_inflam"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- # [1] 39 -->
<!-- sum(list_dt_condition_coef_unscaled[["fibrosis_severity"]][[paste0("p.value_",params$pAdjMethod)]]< 0.05) -->
<!-- #[1] 0 -->
<!-- ``` -->

<!-- compare scaled and unscaled embedding association estimates -->

<!-- ```{r} -->
<!-- cor(as.numeric(list_dt_condition_coef_unscaled[["fibrosis"]]$estimate), as.numeric(list_dt_condition_coef[["fibrosis"]]$estimate)) -->
<!-- #[1] 0.1755063 -->
<!-- ``` -->

<!-- compare past fibrosis results -->

<!-- ```{r} -->
<!-- dt_condition_coef_fibrosis_200325 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_seurat_7_sc_bulk_modules_fibrosis_association_in_gerhard2018_200325.csv") -->

<!-- dt_condition_coef_fibrosis_200511 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_seurat_7_sc_bulk_modules_scaled_fibrosis_association_in_gerhard2018_200511.csv") -->

<!-- dt_condition_coef_fibrosis_200911 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_seurat_7_sc_bulk_modules_scaled_fibrosis_association_in_moylan2013_200911.csv") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all.equal(dt_condition_coef_fibrosis_200325$module, dt_condition_coef_fibrosis_200511$module) -->
<!-- # [1] TRUE -->
<!-- all.equal(dt_condition_coef_fibrosis_200511$module, dt_condition_coef_fibrosis_200911$module) -->
<!-- # [1] TRUE -->
<!-- ``` -->

<!-- compare previous results  -->

<!-- ```{r} -->
<!-- cor(dt_condition_coef_fibrosis_200325$estimate, dt_condition_coef_fibrosis_200511$estimate) # unscaled vs scaled and centered -->
<!-- # [1] 0.344431 -->
<!-- cor(dt_condition_coef_fibrosis_200511$estimate, dt_condition_coef_fibrosis_200911$estimate) # scaled and centered vs log2 -->
<!-- # [1] 0.4491132 -->
<!-- cor(dt_condition_coef_fibrosis_200325$estimate, dt_condition_coef_fibrosis_200911$estimate) # unscaled vs log2 -->
<!-- # [1] 0.2738659 -->
<!-- ``` -->

<!-- compare old vs newest -->

<!-- ```{r} -->
<!-- dt_new_perslab_mods = list_dt_condition_coef[["fibrosis"]][!grepl("moylan|gerhard", module)] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_dict = openxlsx::read.xlsx("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_seurat_7_wgcna3_module_renamed_dict.xlsx") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_new_perslab_mods$module_original = dt_dict$original[match(dt_new_perslab_mods$module, gsub("-","_",dt_dict$new))] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- idx = match(dt_new_perslab_mods$module_original, dt_condition_coef_fibrosis_200325$module) -->
<!-- idx_notNA = idx[!is.na(idx)] -->
<!-- ``` -->

<!-- ```{r} -->

<!-- all.equal(dt_new_perslab_mods$module_original[!is.na(idx)],dt_condition_coef_fibrosis_200325$module[idx_notNA]) -->
<!-- # [1] TRUE -->
<!-- ``` -->

<!-- ```{R} -->
<!-- cor(as.numeric(dt_new_perslab_mods$estimate[!is.na(idx)]),dt_condition_coef_fibrosis_200325$estimate[idx_notNA]) -->
<!-- # [1] 0.1029858 -->

<!-- cor(as.numeric(dt_new_perslab_mods$estimate[!is.na(idx)]),dt_condition_coef_fibrosis_200511$estimate[idx_notNA]) -->
<!-- # [1] 0.5416771  -->

<!-- cor(as.numeric(dt_new_perslab_mods$estimate[!is.na(idx)]),dt_condition_coef_fibrosis_200911$estimate[idx_notNA]) -->
<!-- # [1] 0.07327647 -->
<!-- ``` -->

<!-- so scaling leads to very different results compared to normalized, and also taking log changes results again. -->

<!-- compare results to moylan -->

<!-- ```{r} -->
<!-- dt_condition_moylan = fread(paste0(dirWGCNA_tables, prefixData, "_mod_fibrosis_assoc_in_moylan2013_201012.csv")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all.equal(dt_condition_moylan$module, list_dt_condition_coef[["fibrosis"]]$module) -->
<!-- # [1] TRUE -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_condition_moylan$estimate, as.numeric(list_dt_condition_coef[["fibrosis"]]$estimate)) -->
<!-- # [1] 0.1607969 -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_condition_moylan$estimate, as.numeric(list_dt_condition_coef_unscaled[["fibrosis"]]$estimate)) -->
<!-- ``` -->

<!-- old gerhard results versus new Moylan (where was the high correspondance?) -->

<!-- ```{r} -->
<!-- dt_condition_moylan_perslab_mods = dt_condition_moylan[!grepl("moylan|gerhard", module)] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_condition_moylan_perslab_mods$module_original = dt_dict$original[match(dt_condition_moylan_perslab_mods$module, gsub("-","_",dt_dict$new))] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- idx = match(dt_condition_moylan_perslab_mods$module_original, dt_condition_coef_fibrosis_200325$module) -->
<!-- idx_notNA = idx[!is.na(idx)] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all.equal(dt_condition_moylan_perslab_mods$module_original[!is.na(idx)],dt_condition_coef_fibrosis_200325$module[idx_notNA]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_condition_moylan_perslab_mods$estimate[!is.na(idx)],dt_condition_coef_fibrosis_200325$estimate[idx_notNA]) -->
<!-- # [1] 0.05473216 -->
<!-- ``` -->

<!-- so the oldest gerhard results (unscaled?) don't match moylan at all. -->

<!-- ```{r} -->
<!-- cor(dt_condition_moylan_perslab_mods$estimate[!is.na(idx)],dt_condition_coef_fibrosis_200511$estimate[idx_notNA]) -->
<!-- # [1] 0.2637417 -->
<!-- ``` -->

<!-- there is a small correlation between moylan and the scaled? results -->

<!-- ```{r} -->
<!-- cor(dt_condition_moylan_perslab_mods$estimate[!is.na(idx)],dt_condition_coef_fibrosis_200911$estimate[idx_notNA]) -->
<!-- # 0.8042089 -->
<!-- ``` -->

<!-- very high correspondance when we use log2. Let's confirm this. -->

<!-- ### annotate modules in dt_geneMod -->

<!-- perslab -->

<!-- ```{r} -->
<!-- vec_mods_fibrosis = list_dt_condition_coef[["fibrosis"]][p.value_BH<0.05,module] -->
<!-- ``` -->


<!-- ```{r} -->
<!-- dt_geneMod_perslab[["cell_cluster_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_perslab[[colModule]] %in% vec_mods_fibrosis, dt_geneMod_perslab[[colCellCluster]], NA_character_) -->

<!-- dt_geneMod_perslab[["module_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_perslab[[colModule]] %in% vec_mods_fibrosis ,dt_geneMod_perslab[[colModule]], NA_character_) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- dt_geneMod_gerhard2018[["cell_cluster_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_gerhard2018[[colModule]] %in% vec_mods_fibrosis, dt_geneMod_gerhard2018[[colCellCluster]], NA_character_) -->

<!-- dt_geneMod_gerhard2018[["module_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_gerhard2018[[colModule]] %in% vec_mods_fibrosis, dt_geneMod_gerhard2018[[colModule]], NA_character_) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- dt_geneMod_moylan2013[["cell_cluster_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_moylan2013[[colModule]] %in% vec_mods_fibrosis, dt_geneMod_moylan2013[[colCellCluster]], NA_character_) -->

<!-- dt_geneMod_moylan2013[["module_fibrosis_gerhard2018"]] <- ifelse(dt_geneMod_moylan2013[[colModule]] %in% vec_mods_fibrosis, dt_geneMod_moylan2013[[colModule]], NA_character_) -->
<!-- ``` -->

## write out results

```{r}
fwrite(x=dt_bootOut_stats, file=paste0(dirWGCNA_tables, prefixData, "_mod_assoc_in_gerhard2018_", flagDate,".csv"))

openxlsx::write.xlsx(dt_bootOut_stats, file=paste0(dirWGCNA_tables, prefixData, "_mod_assoc_in_gerhard2018_", flagDate,".xlsx"))
```


<!-- ```{r} -->
<!-- for (name in names(list_dt_condition_coef)) { -->

<!--   fwrite(x=list_dt_condition_coef[[name]], file=paste0(dirWGCNA_tables, prefixData, "_mod_", name, "_assoc_in_gerhard2018_", flagDate,".csv")) -->
<!--   openxlsx::write.xlsx(x=list_dt_condition_coef[[name]],file = paste0(dirWGCNA_tables, prefixData, "_mod_", name, "_assoc_in_gerhard2018_", flagDate,".xlsx")) -->

<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- fwrite(dt_geneMod_perslab, file="/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz", compress="gzip") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- fwrite(dt_geneMod_gerhard2018, file="/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz", compress="gzip") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- fwrite(dt_geneMod_moylan2013, file="/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_moylan2013_wgcna1_geneMod.csv.gz", compress="gzip") -->
<!-- ``` -->

## Additional association analyses

look at individual gene expression 

```{r}
vec_genes = c("VCAN")
```

```{r}
condition = quote(match(vec_genes, dt_datExpr$gene))
 
mat_datExpr_sub = dt_datExpr[eval(condition)] %>% dplyr::select(!gene) %>% t  %>% scale 

colnames(mat_datExpr_sub) = vec_genes

rownames(mat_datExpr_sub) <- colnames(dt_datExpr)[-1]
``` 

run linear models 

```{r}
list_list_tbl_regr_genes <- lapply(list_dt_metadata, function(dt_metadata_sub) {
  
  dt_model_tmp = data.table(dt_metadata_sub, mat_datExpr_sub[dt_metadata_sub$run_accession,])
  colnames(dt_model_tmp)[(ncol(dt_metadata_sub)+1):ncol(dt_model_tmp)] <- vec_genes
  dt_model_tmp$SEX <- factor(dt_model_tmp$SEX)
  
  lapply(colnames(mat_datExpr_sub), function(gene) {
    
    formula_tmp <- formula(paste0(gene, "~ SEX + BMI_surg + Age + condition"))
    ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp)
    regr<-lm(formula = formula_tmp, dt_model_tmp)
    regr %>% broom::tidy() 
  }) -> list_tbl_tmp 
  names(list_tbl_tmp) = colnames(mat_datExpr_sub)
  return(list_tbl_tmp)
})

names(list_list_tbl_regr_genes) <- names(list_dt_metadata)
```

```{r}
lapply(vec_genes, function(gene) {
  formula_tmp <- formula(paste0(gene, "~ fibrosis"))
  ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp)
  regr<-lm(formula = formula_tmp, dt_data_genes)
  regr %>% broom::tidy() 
}) -> list_tbl_regr

names(list_tbl_regr) <- vec_genes
```

extract coefficients for the regressors 

```{r}
mat_out <- sapply(list_tbl_regr, function(tbl_regr) {
    tbl_regr[tbl_regr$term=="fibrosis",colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")] 
  }) %>% t

dt_coef = data.table("gene"= rownames(mat_out), "condition"="fibrosis", mat_out)
```

next: 

liver_module_sc_merge_to_metamodules.Rmd
