---
title: 'Liver - gene meta-network geneset tests'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

previous:
liver_module_sc_merge_to_metamodules.Rmd

# Setup

## Load packages

```{r}
#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr"))
#library("Seurat")

library("data.table")
library("Matrix")
library("parallel")
library("magrittr")
#library("readr")
#library("tidyverse")
#library("dplyr")
#library("ggplot2")
library("here")
#library("rlm")
if (!requireNamespace("gerr", quietly = TRUE)) {
  #devtools::install_github("TaoDFang/gerr")
  download.file(url="https://github.com/TaoDFang/gerr/archive/master.zip", destfile = "/projects/jonatan/R-package-repo/gerr_master.zip")
  system2(command = "unzip", args= "/projects/jonatan/R-package-repo/gerr_master.zip")
  install.packages(type="source", repos=NULL, pkgs = "/projects/jonatan/R-package-repo/gerr-master")
}
  
library("gerr")
library("glmnet")

library("qusage",lib.loc = "~/R/x86_64-pc-linux-gnu-library/")
library("GO.db")
library("gage") # provides GO genesets
#library("msigdbr") # provides genesets as tbl
#library("corrr")
#library("lme4")
#library("WGCNA")
#devtools::install_github(repo = "satijalab/seurat", ref = "develop")

#BiocManager::install("ReactomePA") 
#library("ReactomePA")
library("GSA")

# source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))


# Set options


options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

#options(datatable.WhenJisSymbolThenCallingScope=TRUE)

# Set constants

randomSeed = 12345
set.seed(randomSeed)

pval.adjust.method = "BH"
pValThreshold = 0.05#params$pValThreshold

genesetLengthMin = 5
genesetLengthMax = 1000
nRep = 10000 # bootstrap replicates
# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

alpha =  0.25 # for GERR calling glmnet elasticnet, alpha \in [0,1], alpha==0 => ridge regression, alpha==1 => LASSO regression. Defaults to 0.5
        # ?gerr::regression_selected_pathways

#generic project constants 


prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
#prefixOut <- "SCT"
#prefixOut <- "moylan2013_wgcna_assoc"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_meta"
colCellCluster = "cell_cluster_meta"
```

## load data

gene network data

```{r}
path_geneMod_perslab <- paste0(dirWGCNA_tables,"liver_perslab_int_wgcna3_geneMod.csv.gz")
dt_geneMod_perslab <- fread(path_geneMod_perslab)
head(dt_geneMod_perslab)
```

<!-- ```{r} -->
<!-- path_geneMod_gerhard2018 <- paste0(dirWGCNA_tables,"liver_gerhard2018_", "wgcna3","_geneMod_merged_200325.csv.gz") -->
<!-- dt_geneMod_gerhard2018 <-fread(file=path_geneMod_gerhard2018) -->
<!-- head(dt_geneMod_gerhard2018) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_geneMod_gerhard2018$module <- paste0("gerhard2018_",dt_geneMod_gerhard2018$module) -->
<!-- ``` -->



get a vector of all the genes as a background for gene test

```{r}
if (!file.exists(here("data","liver_perslab_int_genesBackground.RDS"))) {
  seuratObj <- load_obj("/projects/jonatan/pub-perslab/18-liver-fred/output/liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj.RDS.gz")

  Seurat::DefaultAssay(seuratObj) = "RNA"

  vec_genes_perslab <- rownames(seuratObj)

  length(vec_genes_perslab)
  #[1] 36116

  saveRDS(vec_genes_perslab, file = "/projects/jonatan/pub-perslab/18-liver-fred/data/liver_perslab_int_genesBackground.RDS")

  rm(seuratObj)

  } else {

  vec_genes_perslab <- readRDS(here("data","liver_perslab_int_genesBackground.RDS"))

}
```

<!-- ```{r} -->
<!-- if (!file.exists(here("data","liver_gerhard2018_genesBackground.RDS.gz"))) { -->
<!--   dt_datExpr_gerhard <- fread(here("data", "liver_gerhard2018_norm.counts.csv.gz")) -->

<!--   vec_genes_gerhard2018 <- dt_datExpr_gerhard$gene -->

<!--   length(vec_genes_gerhard2018) -->
<!--   #[1] 33660 -->

<!--   saveRDS(vec_genes_gerhard2018, file = "/projects/jonatan/pub-perslab/18-liver-fred/data/liver_gerhard2018_genesBackground.RDS") -->

<!--   rm(dt_datExpr_gerhard) -->

<!--   } else { -->

<!--   vec_genes_gerhard2018 <- readRDS(here("data","liver_gerhard2018_genesBackground.RDS")) -->

<!-- } -->
<!-- ``` -->

```{r}
vec_allGenes = vec_genes_perslab
```

<!-- ```{r} -->
<!-- vec_allGenes <- union(vec_genes_gerhard2018, vec_genes_perslab) -->
<!-- length(vec_allGenes) -->
<!-- # [1] 36131 -->
<!-- ``` -->

### genesets

#### download gene ontology genesets

```{r}
#data(bods)
#list_list_vec_genesetsGO <- bods(species = "human", pkg.name=NULL, id.type = "eg")
list_list_vec_genesetsGO <- go.gsets(species = "human", pkg.name=NULL, id.type = "eg")

list_vec_genesetsGO <- list_list_vec_genesetsGO[[1]]

# filter out very large or small genesets

list_vec_genesetsGO %>% sapply(., length) %>% summary
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
   #  1.0     2.0     6.0    98.8    24.0 19659.0

list_vec_genesetsGO <- list_vec_genesetsGO[sapply(list_vec_genesetsGO, length) >= genesetLengthMin & sapply(list_vec_genesetsGO, length) <= genesetLengthMax]

length(list_vec_genesetsGO)
# [1] 12044

#map the GO genesets from Entrez Gene to symbol

# https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/egSymb
# https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/eg2sym

data(egSymb)

list_vec_genesetsGO <- lapply(list_vec_genesetsGO, function(vec_geneset) {
  vec_geneset <- eg2sym(eg=vec_geneset)
  vec_geneset[!is.na(vec_geneset)]
})

# To get GO definitions we use GO.db package

dt_GO = as.data.table(GO.db::GOTERM)
```

#### Download KEGG genesets

```{r}
data(kegg.gs)
length(kegg.gs)
# [1] 177
# Hm it says 186 on the msigdb page http://software.broadinstitute.org/gsea/msigdb/genesets.jsp?collection=CP:KEGG

list_vec_genesetsKEGG = kegg.gs

rm(kegg.gs)

# translate 

data(egSymb)

list_vec_genesetsKEGG <- lapply(list_vec_genesetsKEGG, function(vec_geneset) {
  vec_geneset <- eg2sym(eg=vec_geneset)
  vec_geneset[!is.na(vec_geneset)]
})

```

```{r}
#### download Reactome genesets

destfile = "/projects/jonatan/data/genesets/ReactomePathways.gmt.zip"

if (!file.exists(destfile)) {
  download.file(url="https://reactome.org/download/current/ReactomePathways.gmt.zip", destfile)
}

reactome.gmt <- GSA.read.gmt(unzip("/projects/jonatan/data/genesets/ReactomePathways.gmt.zip"))

list_vec_genesetsREACT = reactome.gmt$genesets
list_vec_genesetsREACT = lapply(list_vec_genesetsREACT, toupper)
names(list_vec_genesetsREACT) = reactome.gmt$geneset.names
```

# Analysis

Make a list of vectors of genes in the WGCNA gene modules

combine dt_geneMod

```{r}
dt_geneMod_merged = dt_geneMod_perslab
```

<!-- ```{r} -->
<!-- dt_geneMod_merged = rbindlist(list(dt_geneMod_perslab,dt_geneMod_gerhard2018)) -->
<!-- ``` -->

```{r}

vec_mod_meta = dt_geneMod_merged[[colModule]][!is.na(dt_geneMod_merged[[colModule]]) & nchar(dt_geneMod_merged[[colModule]])>0] %>% unique

list_vec_modmetagenes <- lapply(vec_mod_meta, function(mod_meta) {
  x = dt_geneMod_merged[,..colModule][[1]]
  condition <- x==mod_meta
  vec_out = dt_geneMod_merged %>% dplyr::filter(., condition) %>% dplyr::select(genes) %>% '[['(1)
  vec_out = unique(vec_out)
})

names(list_vec_modmetagenes) <- vec_mod_meta
```


Gene-set Enrichment with Regularized Regression - GERR 

In the manuscript *Gene-set Enrichment with Regularized Regression*, we propose using regularized regression to model the relationship between $Y$, a binary dependent (target) variable indicating membership of genes in a set of genes of interest (GOI hereafter), and $\Omega$, a matrix of binary variables indicating membership of genes in gene-sets that are potentially overlapping or even identical with each other.

Classically, binary target variables are often modeled by logistic regression. Alternatively, they can also be modeled by simple linear regression [@agresti_introduction_2019], even when the target variable is a dichotomy, namely either $0$ or $1$ [@hellevik_linear_2009]. 

In this document, we illustrate how the two types of modelling can be constructed with `gerr`, the software package that we published along with the manuscript. In addition, we compare the results of elastic-net regression using either the linear regression or the logistic regression.

### Reguarlized linear regression

The link function of the generalized linear regression is specified by, the `family` parameter in the `glmnet` and `cv.glmnet` functions in the `glmnet` package. First, we construct a linear regression model, using the Gaussian family (`family="gaussian"`).


Convert list of genesets to sparse binary matrix: 

A binary background matrix whose columns are the pathways/gene sets and whose rows are all the genes from pathways/gene sets . It could be in sparse matrix format ((inherit from class "sparseMatrix" as in package Matrix) to save memory. For gene i and pathway j, the value of matrix(i,j) is 1 if gene i belonging to pathway j otherwise 0. 

**Users could leave it as default value then it will use pre-collected gene_pathway_matrix from GO Ontology and REACTOME database. Otherwise, they could use their own customized gene_pathway_matrix**

## Gene Ontology

```{r}
gene_pathway_matrixGO = sapply(list_vec_genesetsGO, function(vec_geneset) {
  (vec_allGenes %in% vec_geneset) %>% as.numeric
}, simplify = T)

rownames(gene_pathway_matrixGO) <- vec_allGenes
  
dim(gene_pathway_matrixGO)
#[1] 36116  12044

#convert to sparse matrix format 

gene_pathway_matrixGO %>% as(., "sparseMatrix") -> gene_pathway_matrixGO

# run regularised regression gene test 


list_gaussResGO <- lapply(list_vec_modmetagenes, function(vec_modGenes) {
  regression_selected_pathways(gene_input=vec_modGenes, 
                               gene_pathway_matrix = gene_pathway_matrixGO,
                               family="gaussian",
                               alpha=alpha)})
```

### gather the results in a data.table

```{r}
vec_module_GERR_GO <- rep(names(list_gaussResGO), sapply(list_gaussResGO, function(gaussRes) {gaussRes$selected_pathways_names %>% length}))


#vec_cell_cluster <- dt_metadataAssoc[,cell_cluster,][match(vec_module,dt_metadataAssoc[,module,] )] 
vec_coef_GO = lapply(list_gaussResGO, function(x) x[["selected_pathways_coef"]])  %>% unlist(use.names = F)


vec_p.value_fisher_GO = lapply(list_gaussResGO, function(x) x[["selected_pathways_fisher_pvalue"]])  %>% unlist(use.names = F)

vec_GOid <- lapply(list_gaussResGO, function(x) x[["selected_pathways_names"]] 
                     %>%  names 
                     %>% substr(.,start=1, stop = 10))  %>% unlist(use.names = F)

lapply(list_gaussResGO, function(x) {
  x[["selected_pathways_names"]] %>% names %>% substr(., start=12, stop=1000)
  })  %>% unlist(use.names = F) -> vec_GOterm


sapply(vec_GOid, function(id) {
  dt_GO$Definition[dt_GO$go_id == id] %>% unique
}) %>% unlist(use.names = F) -> vec_GOdef 


sapply(vec_GOid, function(id) {
  dt_GO$Ontology[dt_GO$go_id == id] %>% unique
}) %>% unlist(use.names = F) -> vec_ontology

# adjust p-values for totalt number of tests (strict!)
vec_p.value_fisher_GO_BH = p.adjust(vec_p.value_fisher_GO, 
                                    method="BH",
                                    n=length(list_vec_modmetagenes)*(length(list_vec_genesetsGO)+length(list_vec_genesetsKEGG)+length(list_vec_genesetsREACT)))

# for (module in unique(vec_module_GERR)) {
#   vec_logical = vec_module_GERR==module
#   vec_p.value_fisher_GO_BH = p.adjust(p=vec_p.value_fisher,
#                             method = "BH", 
#                             n = length(list_vec_modmetagenes))
# }

dt_resultsGO <- data.table(
  #cell_cluster = vec_cell_cluster,
  module = vec_module_GERR_GO,
  coef = vec_coef_GO,
  p.value_fisher = vec_p.value_fisher_GO,
  p.value_fisher_BH = vec_p.value_fisher_GO_BH,
  GO_id = vec_GOid,
  GO_term = vec_GOterm,
  GO_definition  = vec_GOdef,
  GO_ontology = vec_ontology
  )

# order results by fisher test p.value 

dt_resultsGO <- dt_resultsGO[order(p.value_fisher_BH)]
```

## KEGG

```{r}
gene_pathway_matrixKEGG = sapply(list_vec_genesetsKEGG, function(vec_geneset) {
  (vec_allGenes %in% vec_geneset) %>% as.numeric
}, simplify = T)

rownames(gene_pathway_matrixKEGG) <- vec_allGenes
  
dim(gene_pathway_matrixKEGG)
#[1] 36116  177

# convert to sparse matrix format 

gene_pathway_matrixKEGG %>% as(., "sparseMatrix") -> gene_pathway_matrixKEGG

# run regularised regression gene test 

list_gaussResKEGG <- lapply(list_vec_modmetagenes, function(vec_modGenes) {
  regression_selected_pathways(gene_input=vec_modGenes, 
                               gene_pathway_matrix = gene_pathway_matrixKEGG,
                               family="gaussian",
                               alpha=alpha)})
```

### gather results in a data.table

```{r}
vec_module_KEGG <- rep(names(list_gaussResKEGG), sapply(list_gaussResKEGG, function(gaussRes) {gaussRes$selected_pathways_names %>% length}))

#vec_cell_clusterKEGG <- dt_metadataAssoc[,cell_cluster,][match(vec_moduleKEGG,dt_metadataAssoc[,module,] )]

vec_coef_KEGG = lapply(list_gaussResKEGG, function(x) x[["selected_pathways_coef"]])  %>% unlist(use.names = F)

vec_p.value_fisher_KEGG = lapply(list_gaussResKEGG, function(x) x[["selected_pathways_fisher_pvalue"]])  %>% unlist(use.names = F)

vec_KEGGid <- lapply(list_gaussResKEGG, function(x) x[["selected_pathways_names"]] 
                     %>%  names 
                     %>% substr(.,start=1, stop = 8))  %>% unlist(use.names = F)

lapply(list_gaussResKEGG, function(x) {
  x[["selected_pathways_names"]] %>% names %>% substr(., start=10, stop=1000)
  })  %>% unlist(use.names = F) -> vec_KEGGterm

# adjust p-values for totalt number of tests (strict!)
vec_p.value_fisher_KEGG_BH = p.adjust(vec_p.value_fisher_KEGG, 
                                    method="BH",
                                n=length(list_vec_modmetagenes)*(length(list_vec_genesetsGO)+length(list_vec_genesetsKEGG)+length(list_vec_genesetsREACT)))
# for (module in unique(vec_moduleKEGG)) {
#   vec_logical = vec_moduleKEGG==module
#   vec_p.value_fisherKEGG_BH = p.adjust(p=vec_p.value_fisherKEGG_BH, 
#                                        method = "BH", 
#                                        n = length(list_vec_modmetagenes))
# }

dt_resultsKEGG <- data.table(
  #cell_cluster = vec_cell_clusterKEGG,
  module = vec_module_KEGG,
  coef = vec_coef_KEGG,
  p.value_fisher = vec_p.value_fisher_KEGG,
  p.value_fisher_BH = vec_p.value_fisher_KEGG_BH,
  KEGG_id = vec_KEGGid,
  KEGG_term = vec_KEGGterm
  )

# order results by fisher test p.value 

dt_resultsKEGG <- dt_resultsKEGG[order(p.value_fisher_BH)]
```

## Reactome 

```{r}
gene_pathway_matrixREACT = sapply(list_vec_genesetsREACT, function(vec_geneset) {
  (vec_allGenes %in% vec_geneset) %>% as.numeric
}, simplify = T)

rownames(gene_pathway_matrixREACT) <- vec_allGenes
  
dim(gene_pathway_matrixREACT)
#[1] 36116  2400

# convert to sparse matrix format 

gene_pathway_matrixREACT %>% as(., "sparseMatrix") -> gene_pathway_matrixREACT

# run regularised regression gene test 

list_gaussResREACT <- lapply(list_vec_modmetagenes, function(vec_modGenes) {
  regression_selected_pathways(gene_input=vec_modGenes, 
                               gene_pathway_matrix = gene_pathway_matrixREACT,
                               family="gaussian",
                               alpha=alpha)})
```

### gather results in a data.table

```{r}
vec_module_REACT <- rep(names(list_gaussResREACT), sapply(list_gaussResREACT, function(gaussRes) {gaussRes$selected_pathways_names %>% length}))

vec_coef_REACT = lapply(list_gaussResREACT, function(x) x[["selected_pathways_coef"]])  %>% unlist(use.names = F)

vec_p.value_fisher_REACT = lapply(list_gaussResREACT, function(x) x[["selected_pathways_fisher_pvalue"]])  %>% unlist(use.names = F)

vec_REACTnames <- lapply(list_gaussResREACT, function(x) x[["selected_pathways_names"]] 
                     %>%  names) %>% unlist(use.names = F)

lapply(list_gaussResREACT, function(x) {
  reactome.gmt$geneset.descriptions[match(names(x[["selected_pathways_names"]]), reactome.gmt$geneset.names)]}) %>% unlist -> vec_descriptionREACT

# adjust p-values for total number of tests (strict!)
vec_p.value_fisher_REACT_BH = p.adjust(vec_p.value_fisher_REACT, 
                                    method="BH",
                                n=length(list_vec_modmetagenes)*(length(list_vec_genesetsGO)+length(list_vec_genesetsKEGG)+length(list_vec_genesetsREACT)))

# for (module in unique(vec_moduleREACT)) {
#   vec_logical = vec_moduleREACT==module
#   vec_p.value_fisherREACT_BH = p.adjust(p=vec_p.value_fisherREACT_BH, 
#                                        method = "BH", 
#                                        n = length(list_vec_modmetagenes))
# }

dt_resultsREACT <- data.table(
  #cell_cluster = vec_cell_clusterREACT,
  module = vec_module_REACT,
  coef = vec_coef_REACT,
  p.value_fisher = vec_p.value_fisher_REACT,
  p.value_fisher_BH = vec_p.value_fisher_REACT_BH,
  REACT_name = vec_REACTnames,
  REACT_description = vec_descriptionREACT
  )

#order results by fisher test p.value 


dt_resultsREACT <- dt_resultsREACT[order(p.value_fisher_BH)]
```


# write results to disc

## GO 

```{r}
fwrite(x = dt_resultsGO, file = paste0(dirWGCNA_tables, prefixData, "_", prefixRun, "_", flagDate,"_GERR_GOenrichmentResults_metaMod.csv"))

# also as xlsx for convenience

openxlsx::write.xlsx(x=dt_resultsGO, file = paste0(dirWGCNA_tables,prefixData, "_", prefixRun,  "_", flagDate,"_GERR_GOenrichmentResults_metaMod.xlsx"))

## KEGG


fwrite(x = dt_resultsKEGG, file = paste0(dirWGCNA_tables, prefixData, "_", prefixRun,  "_", flagDate,"_GERR_KEGGenrichmentResults_metaMod.csv"))


# also as xlsx for convenience


openxlsx::write.xlsx(x=dt_resultsKEGG, file =  paste0(dirWGCNA_tables, prefixData, "_", prefixRun, "_",  flagDate,"_GERR_KEGGenrichmentResults_metaMod.xlsx"))

## REACTOME

fwrite(x = dt_resultsREACT, file = paste0(dirWGCNA_tables, prefixData, "_", prefixRun,  "_", flagDate,"_GERR_REACTOMEenrichmentResults_metaMod.csv"))


# also as xlsx for convenience


openxlsx::write.xlsx(x=dt_resultsREACT, file =  paste0(dirWGCNA_tables, prefixData, "_", prefixRun, "_",  flagDate,"_GERR_REACTOMEenrichmentResults_metaMod.xlsx"))
```


<!-- if not already done, also write out #gene module tables as xlsx to allow to query  -->

<!-- ```{r} -->
<!-- if (FALSE) { -->
<!--   openxlsx::write.xlsx(x=dt_geneMod_perslab, file = paste0(dirWGCNA_tables,"liver_perslab_int_wgcna3_geneMod_merged_200325.xlsx")) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- if (FALSE) { -->
<!-- openxlsx::write.xlsx(x=dt_geneMod_gerhard2018, file = paste0(dirWGCNA_tables,"liver_gerhard2018_wgcna3_geneMod_merged_200325.xlsx")) -->
<!-- } -->
<!-- ``` -->
