---
title: 'Liver sc data normalization and de-confounding'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

# Setup

## Set options

```{r}
options(stringsAsFactors = F,
        use="pairwise.complete.obs",
        warn=1,
        verbose=F,
        mc.cores=40# for parallel computation
        )
```

## Load packages

```{r}
#devtools::install_github(repo = "satijalab/seurat", ref = "develop") # dev version of Seurat

library(Seurat)
library("data.table")
library(dplyr)
library(ggplot2)
library(Matrix)
library(parallel)
#library(readr)
#library(tidyr)
#library(AUCell)

#library(Oscope)

library(openxlsx)
library(here)
library("sva") #https://static-content.springer.com/esm/art%3A10.1186%2Fs13059-019-1700-9/MediaObjects/13059_2019_1700_MOESM4_ESM.html

# https://www.bioconductor.org/packages/release/bioc/html/sva.html

#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr",  "AUCell", "Oscope", "openxlsx", "here"))
```


## source utility functions

```{r}
source(here("perslab-sc-library","utility_functions.R"))
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)

pValThreshold = 0.05
pval.adjust.method = "BH"
```

generic project constants 

```{r}
prefixData <- "liver_perslab_int"
prefixRun <- "seurat_7"
prefixOut <- "SCTint"
```

## specific options and params

```{r}
options(future.globals.maxSize = 8000 * 1024^2)
nCores=  20
npcs=60
res_primary=0.8
```

## load data

```{r}
pathSeuratObj <- "/nfsdata/projects/jonatan/pub-perslab/18-liver-fred/output/liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj.RDS.gz"
  #dir(path=here("output"), pattern = paste0(".*",prefixRun, ".*\\.RDS\\.gz"), full.names = T)
seuratObj <- load_obj(pathSeuratObj)
```

```{r}
head(seuratObj@meta.data)
```

```{r}
seuratObj@active.assay <- "RNA"
#seuratObj@active.assay <- "integrated"

```

## Cell-cycle analysis 
See https://satijalab.org/seurat/cell_cycle_vignette.html

```{r}
# list_cellCycleGenes <- lapply(paths_cellCycleGenes, function(x) load_obj(x)[,1])
# names(list_cellCycleGenes) <- names(paths_cellCycleGenes)

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase 
list_cellCycleGenes <- list()
list_cellCycleGenes[["s.genes"]] <- cc.genes$s.genes
list_cellCycleGenes[["g2m.genes"]] <- cc.genes$g2m.genes
```

Compute cell cycle scores using provided genesets

```{r}
# First get rid of zero-expressed genes
# seuratObj <- Seurat::CreateSeuratObject(counts=GetAssayData(seuratObj, slot= "counts"), 
#                                              project= prefixData, 
#                                              assay = "RNA", 
#                                              min.cells = 5, # filtering genes is the only reason to make a new Seurat object
#                                              min.features = 0,
#                                              meta.data = seuratObj@meta.data)

seuratObj <- Seurat::CellCycleScoring(object = seuratObj, 
                               s.features = if (!is.null(list_cellCycleGenes[["s.genes"]])) list_cellCycleGenes[["s.genes"]] else NULL, 
                               g2m.features = if (!is.null(list_cellCycleGenes[["g2m.genes"]])) list_cellCycleGenes[["g2m.genes"]] else NULL, 
                              set.ident = F)

seuratObj$CC.Difference <- seuratObj$S.Score - seuratObj$G2M.Score
```

Visualize cell cycle marker genes

```{r}
seuratObj@active.assay <- "integrated"
```

```{r, fig.width=12, fig.height= 10}
p<-FeaturePlot(object = seuratObj, features=c("S.Score"), blend = F, label = F, split.by = "sample_ID")
saveMeta(savefnc=ggsave, plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun,  "_", seuratObj@active.assay,"_S.Score_cellcycle_featPlot_splitClust.pdf")), width=40, height=8)

p<-FeaturePlot(object = seuratObj, features=c("S.Score"), blend = F, label = F)
saveMeta(savefnc=ggsave, plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun,  "_", seuratObj@active.assay,"_S.Score_cellcycle_featPlot.pdf")), width=12, height=12)

p<-FeaturePlot(object = seuratObj, features=c("G2M.Score"), blend = F, label = F, split.by = "sample_ID")
saveMeta(savefnc=ggsave, plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun, "_", seuratObj@active.assay, "_G2M.Score_cellcycle_featPlot_splitClust.pdf")), width=40, height=8)
  
p<-FeaturePlot(object = seuratObj, 
               features=c("G2M.Score"), 
               blend = F, 
               label = F)
saveMeta(savefnc=ggsave, plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun,  "_", seuratObj@active.assay,"_G2M.Score_cellcycle_featPlot.pdf")), width=12, height=12)

p<-FeaturePlot(object = seuratObj, 
               features=c("CC.Difference"),
               blend = F, 
               label = F, 
               split.by = "sample_ID")
saveMeta(savefnc=ggsave, plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun,  "_", seuratObj@active.assay,"_CC.Difference_cellcycle_featPlot_splitClust.pdf")), width=40, height=8)
2

p <- FeaturePlot(object = seuratObj, 
                 features=c("CC.Difference"), 
                 blend = F, 
                 label = F)

saveMeta(savefnc=ggsave,plot=p,filename = here("plots",paste0(prefixData, "_", prefixRun,  "_", seuratObj@active.assay,"_CC.Difference_cellcycle_featPlot.pdf")), width=12, height=12)

```

## Re-integrate data using sc-transform workflow

## SCTransform, then integrate

Regress out percent.mito and cell cycle difference

```{r}
seuratObj@active.assay <- "RNA"
```


```{r}
# NV: alterantively, could run SCtransform **before** integrating  https://satijalab.org/seurat/v3.0/integration.html
list_seuratObj <- SplitObject(seuratObj, split.by = "sample_ID")
```

```{r}
for (i in 1:length(list_seuratObj)) {
    list_seuratObj[[i]] <- SCTransform(list_seuratObj[[i]],
                              vars.to.regress = c("percent.mito",  "CC.Difference"), #  Variables to regress out in a second non-regularized linear regression. For example, percent.mito. Default is NULL
                              do.correct.umi=T,
                              do.scale=F,
                              do.center=T,
                              return.only.var.genes =F,
                              seed.use = randomSeed,#params$randomSeed,
                              verbose=T)
}
```

```{r}
features <- Seurat::SelectIntegrationFeatures(object.list = list_seuratObj, nfeatures = 3000)

list_seuratObj <- Seurat::PrepSCTIntegration(object.list = list_seuratObj,
                                     anchor.features = features,
                                     verbose = FALSE)
```

```{r}
anchors <- Seurat::FindIntegrationAnchors(object.list = list_seuratObj, 
                                  normalization.method = "SCT", 
                                  anchor.features = features, 
                                  verbose = FALSE)

seuratObjIntegrated <- Seurat::IntegrateData(anchorset =anchors, 
                                     normalization.method = "SCT", 
                                     verbose = FALSE)
```

## First check what data looks like when doing PCA and UMAP on SC Transformed data


```{r}

# for some reason we can't use the SCT data computed on the dataset split by sample above. So redo SCT normalization on the whole dataset
# Can we use vars.to.regress to remove batch effect? NO: We do not recommend using ScaleData for removing batch effects, for reasons discussed in https://www.nature.com/articles/nbt.4096. https://github.com/satijalab/seurat/issues/1351

seuratObjIntegrated <- SCTransform(seuratObjIntegrated, 
                                   assay = "RNA", 
                                   new.assay.name = "SCTall", 
                                   vars.to.regress = c("percent.mito",  "CC.Difference" ), #  Variables to regress out in a second non-regularized linear regression. For example, percent.mito. Default is NULL
                                  do.correct.umi=T,
                                  do.scale=F,
                                  do.center=T,
                                  return.only.var.genes =F,
                                  seed.use = randomSeed,#params$randomSeed,
                                  verbose=T)
seuratObjIntegrated <- RunPCA(seuratObjIntegrated, assay="SCTall", npcs = npcs, seed.use = randomSeed)
seuratObjIntegrated <- RunUMAP(object=seuratObjIntegrated, dims=1:npcs, reduction = "pca", assay=  "SCTall", seed.use = randomSeed)

p1 <-DimPlot(seuratObjIntegrated,  
             assay="SCTall",
             reduction = 'umap', 
             group.by = 'SCT_snn_res.0.8',
             label = FALSE)

saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_UMAP_SCT_FindClusters.pdf")), width = 20, height=12)

p1 <-DimPlot(seuratObjIntegrated,  
             assay="SCTall",
             reduction = 'umap', 
             group.by = 'sample_ID',
             label = FALSE)

saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_UMAP_SCT_sample_ID.pdf")), width = 20, height=12)
```

```{r}
saveMeta(savefnc = saveRDS, object =seuratObjIntegrated, file = here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "seuratObj.RDS.gz")), compress="gzip")
```

## standard clustering workflow on integrated data

```{r}
DefaultAssay(seuratObjIntegrated) <- "integrated"
```

```{r}

seuratObjIntegrated <- Seurat::RunPCA(seuratObjIntegrated, 
                                      assay = "integrated",
                                      npcs = npcs, 
                                      seed.use=randomSeed,
                                      verbose = FALSE)
```

```{r}
seuratObjIntegrated <- FindNeighbors(object = seuratObjIntegrated, dims= 1:npcs, verbose = T)
seuratObjIntegrated <- FindClusters(object = seuratObjIntegrated, verbose = T,resolution =res_primary, random.seed = randomSeed)
```

```{r}
seuratObjIntegrated <- RunUMAP(object=seuratObjIntegrated, dims=1:npcs, reduction = "pca", assay=  "SCT", seed.use = randomSeed)

```

```{r}
p1 <-DimPlot(seuratObjIntegrated, 
             reduction = 'umap', 
             group.by = 'SCT_snn_res.0.8',
             label = TRUE)

saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_UMAP_FindClusters.pdf")), width = 20, height=12)
```

```{r}
p2 <- DimPlot(seuratObjIntegrated, 
              reduction="umap",
              group.by = c("sample_ID"),
              label = F)

saveMeta(savefnc=ggsave, plot =p2, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_UMAP_sampleID.pdf")), width=15, height=15)
```

## Find cluster markers 

```{r}
clusters = seuratObjIntegrated@meta.data[[paste0("SCT_snn_res.", res_primary)]] %>% table %>% names 
Idents(seuratObjIntegrated) <-seuratObjIntegrated@meta.data[[paste0("SCT_snn_res.", res_primary)]]

list_iterable = list("X"=clusters)
fun = function(cluster) {tryCatch({
  FindMarkers(seuratObjIntegrated ,  
              #cells.1=colnames(seurat_obj)[Idents(seurat_obj)==cluster],
              #cells.2=NULL,
              ident.1 = cluster,
              only.pos = T,
              #ident.2 = clusters[clusters!=cluster],
              test.use  ="MAST",
              max.cells.per.ident=1000,
              random.seed=randomSeed,
              #latent.vars = if (!is.null(merge_specify) | !is.null(merge_group_IDs)) "sample_ID" else NULL,
              verbose = T)
}, 
error = function(err) {
  NA_character_
})}
list_markers=NULL
list_markers <- lapply(FUN=fun, "X"=list_iterable[[1]])

# add the gene and cluster as a column
list_markers <- mapply(function(df_markers, cluster) {
  if (!all(sapply(df_markers, is.na))) {
    cbind("gene" = rownames(df_markers), "cluster"=rep(cluster, nrow(df_markers)), df_markers)
  } else {
    NA_character_
  }
},
df_markers=list_markers, 
cluster=names(table(Idents(seuratObjIntegrated))), SIMPLIFY=F)

list_markers <- list_markers[!sapply(list_markers, function(markers) all(is.na(markers)))]
df_markers <- Reduce(x=list_markers, f=rbind)
rownames(df_markers) <- NULL
```

```{r}
saveMeta(savefnc=openxlsx::write.xlsx, x = df_markers, file=here("data", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_clustermarkers.xlsx")))
```

save in order to transfer labels using a separate script 

```{r}
saveMeta(savefnc = saveRDS, object =seuratObjIntegrated, file = here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "seuratObj.RDS.gz")), compress="gzip")
```

saveMeta(savefnc = saveRDS, object =seuratObj, file = here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_perslab_labels_seuratObj.RDS.gz")), compress="gzip")

