---
title: 'Liver - QC Gerhart 2018 bulk data'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

# References:

## Publications

Gerhard,...,DiStefano,2018, J Endocrine Soc, Transcriptomic Profiling of Obesity-Related
Nonalcoholic Steatohepatitis Reveals a Core Set of Fibrosis-Specific Genes
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6018672/

Leti,...DiStefano,2015, Trans. Res. High-throughput sequencing reveals altered expression of
hepatic miRNAs in non-alcoholic fatty liver disease-related fibrosis
https://www.sciencedirect.com/science/article/pii/S1931524415001401?via%3Dihub

DiStefano,...,Gerhard, 2015, Acta Diabetol., Genome-wide analysis of hepatic lipid content
in extreme obesity
https://link.springer.com/article/10.1007%2Fs00592-014-0654-3

Garhard,...,DiStefano, 2014, J. Obes., Identification of Novel Clinical Factors Associated with
Hepatic Fat Accumulation in Extreme Obesity
https://www.hindawi.com/journals/jobe/2014/368210/

## Raw data

NCBI SRA Raw sequence: https://www.ncbi.nlm.nih.gov/bioproject/512027~~

NCBI download page: https://www.ncbi.nlm.nih.gov/home/download/
NCBI SRA handbook: https://www.ncbi.nlm.nih.gov/books/NBK47528/

European Nucleotide Archive sequence: https://www.ebi.ac.uk/ena/browser/text-search?query=PRJNA512027
ENA browser tools: https://github.com/enasequence/enaBrowserTools # MUST BE INSTALLED!!

## metadata 

via email from Johanna DiStefano

# Setup

## Load packages

```{r}
library("data.table")
library("Matrix")
library("parallel")
library("dplyr")
library("ggplot2")
library("here")
library("corrr")
library("irlba")
library("DESeq2")
```

## source utility functions

```{r}
source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)

pval.adjust.method = "BH"
pValThreshold = 0.05#params$pValThreshold
```

generic project constants 

```{r}
#prefixData <- "aizarani"
prefixData = "gerhard2018"
prefixRun = "1"
#prefixOut <- "SCT"
prefixOut <- "QC"
```

## load data

```{r}
path_htseq_outs <- "/data/pub-others/gerhard-j_endocrine_soc-2018/data-htseq-count/"
list_htseq_outs <- fnc_merge_htseq_count(dir_htseq_count_outs = path_htseq_outs)
```

```{r}
mat_datExpr <- list_htseq_outs$mat_counts
dt_htseq_stats <- list_htseq_outs$dt_stats
rm(list_htseq_outs)
```

```{r}
openxlsx::read.xlsx(xlsxFile = here("data", "gerhard2018_patient_metadata.xlsx")) %>% setDT() -> dt_metadata
```

```{r}
dt_sampleID_SRR <- fread("/data/pub-others/gerhard-j_endocrine_soc-2018/Gerhard2018_PRJNA512027.SRR_sample_tsv.txt")
```

# QC 

add run_accession numbers to metadata

```{r}
vec_run_accession <- dt_sampleID_SRR$run_accession[match(dt_metadata$DCL.Patient.ID, dt_sampleID_SRR$`sample alias`)]
```

```{r}
dt_metadata_run_acc <- data.table("run_accession"=vec_run_accession, dt_metadata)
dt_metadata_run_acc <- dt_metadata_run_acc[order(run_accession)]
dt_metadata_run_acc
```

## metadata

from DiStefano: 
It appears that samples DLDR_0037-46 and DLDR_61-62 are listed as normal, but are actually fibrotic (F4) samples. Please keep this in mind when you are analyzing the data.

```{r}
(37:46) %>% paste0("DLDR_00",.) -> vec_fibrotic_1
  
(61:62) %>% paste0("DLDR_00",.) -> vec_fibrotic_2

vec_fibrotic <- c(vec_fibrotic_1, vec_fibrotic_2)
```

```{r}
dt_metadata_run_acc[DCL.Patient.ID %in% vec_fibrotic,"Diagnosis"] <- "Fibrosis 4"
dt_metadata_run_acc[DCL.Patient.ID %in% vec_fibrotic,]
```

```{r}
table(dt_metadata_run_acc$Diagnosis)
```

correct duplicate levels by removing whitespace

```{r}
dt_metadata_run_acc[Diagnosis =="STEATOSIS 2 ","Diagnosis"] <- "STEATOSIS 2" 
dt_metadata_run_acc[Diagnosis =="STEATOSIS 3 ","Diagnosis"] <- "STEATOSIS 3" 
```

```{r}
table(dt_metadata_run_acc$Diagnosis)
```

```{r}
dim(mat_datExpr)
# [1] 60623   185

dim(dt_metadata_run_acc)
# [1] 192   5
```

remove .htseq.counts from mat_datExpr colnames

```{r}
colnames(mat_datExpr) <- gsub("\\.htseq\\.counts","",colnames(mat_datExpr))
```

we filtered out 7 samples while aligning the RNA seq data. Let's remove them in the metadata as well 

```{r}
dt_metadata_run_acc <- dt_metadata_run_acc[run_accession %in% colnames(mat_datExpr)]
```

```{r}
nrow(dt_metadata_run_acc)
dim(mat_datExpr)
```

## expression data

check for duplicate features 

### pre-normalization QC 

```{r}
mat_datExpr %>% rownames %>% duplicated %>% table
# FALSE 
# 60623
```

check gene names

```{r}
mat_datExpr %>% rownames %>% nchar %>% summary
```

### normalize counts

DESeq method

https://hbctraining.github.io/DGE_workshop/lessons/02_DGE_count_normalization.html

create fake metadata (we don't intend to do DE analysis, but DESeq2 requires it.. )

```{r}
df_meta_fake <- data.frame(row.names = colnames(mat_datExpr), "sample_annotation"=factor(sample(x = as.character(0:9),size=ncol(mat_datExpr), replace=T)))
```

```{r}
# create DESeq2 data object
dds <- DESeqDataSetFromMatrix(countData = mat_datExpr, 
                              colData = df_meta_fake, 
                              design = ~ sample_annotation)
```

To perform the median of ratios method of normalization, DESeq2 has a single estimateSizeFactors() function that will generate size factors for us. We will use the function in the example below, but in a typical RNA-seq analysis this step is automatically performed by the DESeq() function, which we will see late

```{r}
dds <- estimateSizeFactors(dds)
```

```{r}
mat_datExpr_norm <- counts(dds, normalized=TRUE)

rm(dds)
mat_datExpr_norm[0:4,0:4]
```


## QC normalized counts

Outlier check : PCA

```{r}
set.seed(randomSeed)
irlba_out <- irlba(A = mat_datExpr_norm, nv = 6)
```

Check how the samples look on singular components 1 and 2, coloured by fibrosis

```{r}
df <- as.data.frame(irlba_out$v)
df[["Diagnosis"]] <- dt_metadata_run_acc$Diagnosis
df[["sample"]] <- colnames(mat_datExpr_norm)
```

Singular vectors 1 and 2
```{r}
ggplot(data = df, aes(x=V1, y=V2), label=sample) + 
  geom_point(size=5, aes(colour=Diagnosis)) +
  geom_text(aes(label=as.character(sample)),hjust=0,vjust=0)
```

Singular vectors 3 and 4
```{r}
ggplot(data = df, aes(x=V3, y=V4), label=sample) + 
  geom_point(size=5, aes(colour=Diagnosis)) +
  geom_text(aes(label=as.character(sample)),hjust=0,vjust=0)
```

Singular vectors 5 and 6
```{r}
ggplot(data = df, aes(x=V5, y=V6), label=sample) + 
  geom_point(size=5, aes(colour=Diagnosis)) +
  geom_text(aes(label=as.character(sample)),hjust=0,vjust=0)
```

sample SRR8378529 is an outlier on PC 1. How much variance do they account for?

```{r}
irlba_out$d
# [1] 70893897.7  5667674.4  4048015.5  1337544.8   917705.5   573675.3
```

```{r}
mat_datExpr_norm <- mat_datExpr_norm[,-which(colnames(mat_datExpr_norm)=="SRR8378529.htseq.counts")]
dim(mat_datExpr_norm)
# 1] 60623   184
```

```{r}
dt_metadata_run_acc <- dt_metadata_run_acc[run_accession!="SRR8378529",]
dim(dt_metadata_run_acc)
```

add a fake metadata 'dataset' variable for WGCNA script to subset by

```{r}
dt_metadata_run_acc$data <- "gerhard2018"
```

### make a copy of the datExpr replacing ensembl with hgnc gene names

```{r}
dt_datExpr_norm <- data.table("gene"=rownames(mat_datExpr_norm),mat_datExpr_norm)
```

```{r}
dt_datExpr_norm_hgnc <- gene_map(dataIn = dt_datExpr_norm, 
                                 colGene = "gene", 
                                 df_mapping = fread("/projects/jonatan/data/gene_remap/gene_annotation_hsapiens.txt.gz"), 
                                 from = "ensembl_gene_id", 
                                 to = "hgnc_symbol", 
                                 replace = T, 
                                 na.rm = T)
```

```{r}
dt_datExpr_perslab <- fread(here("output","liver_perslab_int_seurat_7_datExpr_combat.csv.gz"))
```

```{r}
dim(dt_datExpr_norm)
#[1] 60623   185
dim(dt_datExpr_norm_hgnc)
#[1] 57827   185
```

We only lost ~ 2800 genes

any duplicates?

```{r}
table(duplicated(dt_datExpr_norm_hgnc[[1]]))
# FALSE  TRUE 
# 37159 20668 
```

damn. Ok, better map perslab data to ensembl to avoid averaging..
(in liver_cells_prep_WGCNA.Rmd)

## wrap up 

datExpr ensembl 

```{r}
fwrite(dt_datExpr_norm, here("data", paste0(prefixData, "_norm.counts.csv.gz")), compress="gzip")
```

<!-- datExpr hgnc -->

<!-- ```{r} -->
<!-- fwrite(dt_datExpr_norm_hgnc, here("data", paste0(prefixData, "_norm.counts.hgnc.csv.gz")), compress="gzip") -->
<!-- ``` -->

metadata

```{r}
fwrite(dt_metadata_run_acc, here("data", paste0(prefixData, "_patient_metadata_qc.csv.gz")), compress="gzip")
```

