---
title: 'Liver - metamodule vs metadata association analysis in Moylan 2013 bulk data'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
  pAdjMethod: "BH"
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

previous:
liver_module_sc_merge_to_metamodules.Rmd

# Setup

## Load packages

```{r}
#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr"))
#library("Seurat")

library("data.table")
library("Matrix")
library("parallel")
#library("tidyverse")
library("here")
library("boot")
library("magrittr")
#library("AUCell")
```

## source utility functions

```{r}
source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)
```


generic project constants 

```{r}
prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_pres"
colCellCluster = "cell_cluster_pres"
# naming convention:
# <prefixData>_<prefixRun>_<prefixOut>_<outputType>_<params>_<params>_..._<date>.pdf

```

## load data

```{r}
dt_datExpr <- fread(here("data", "moylan2013.norm.expr.qc.csv.gz"))
```

```{r}
strTmp <- colnames(dt_datExpr[,-"gene"])
dt_datExpr[, (strTmp) := lapply(.SD, as.numeric), .SDcols = strTmp]
```

```{r}
dt_metadata <- fread(here("data", "moylan2013.metadata.csv"))
```

transform fibrosis to indicator variable
```{r}
dt_metadata[,fibrosis:=ifelse(fibrosis=="advanced",1,0),]
```


load data.table with information on module and cell type of origin

```{r}
path_geneMod_perslab_merged <- paste0(dirWGCNA_tables,prefixData, "_",prefixRun_WGCNA,"_geneMod_merged_200325.csv.gz")
  
dt_geneMod_perslab_merged <- fread(path_geneMod_perslab_merged)

head(dt_geneMod_perslab_merged)
```

<!-- ```{r} -->
<!-- path_geneMod_gerhard2018_merged <- paste0(dirWGCNA_tables,"liver_gerhard2018_wgcna3_geneMod_merged_200325.csv.gz") -->

<!-- dt_geneMod_gerhard2018_merged <- fread(path_geneMod_gerhard2018_merged) -->

<!-- head(dt_geneMod_gerhard2018_merged) -->
<!-- ``` -->

# analysis

## compute module embeddings in Moylan (they will be merged as metamodule embeddings)

note that the Moylan data is already normalized

```{r}

dt_geneMod_perslab_merged$module_pres %>% unique -> vec_mods_perslab
# 
vec_mods_perslab<- vec_mods_perslab[!is.na(vec_mods_perslab) & nchar(vec_mods_perslab)>0]
#vec_mods_perslab = c("perslab_honeydew3")
mat_embed_perslab <- sapply(vec_mods_perslab, function(mod) {

  # get gene weights
  condition = quote(dt_geneMod_perslab_merged[[colModule]]==mod)
  vec_pkMs <- dt_geneMod_perslab_merged[eval(condition), pkMs,]

  # normalize module gene weights to sum to 1
  vec_pkMs <- vec_pkMs/sum(vec_pkMs)

  # find corresponding rows of the expression matrix
  vec_idxRow <- match(dt_geneMod_perslab_merged[eval(condition), ..colGeneNames, ][[1]], dt_datExpr[,gene])

  # filter out mis-matches
  vec_pkMs <- vec_pkMs[!is.na(vec_idxRow)]
  vec_idxRow <- vec_idxRow[!is.na(vec_idxRow)]

  # compute weighted sum of normalized expression
  dt_datExpr[vec_idxRow,-("gene")] %>% as.matrix -> mat_sub

  vec_pkMs %*% mat_sub

})

rownames(mat_embed_perslab) <- colnames(dt_datExpr)[-1]

#mat_embed_perslab[0:4,0:4]
```

```{r}
mat_embed <- mat_embed_perslab#cbind(mat_embed_gerhard2018, mat_embed_perslab)
```

### scale embeddings to z-scores

```{r}
mat_embed_scaled = scale(mat_embed)
```

## take linear combinations of module embeddings to get metamodule embeddings

```{r}
vec_mods = unique(dt_geneMod_perslab_merged$module_pres)
vec_mods = vec_mods[nchar(vec_mods)>0]

```

```{r}
names(vec_mods) = dt_geneMod_perslab_merged$module_meta[sapply(vec_mods, function(mod) grep(mod, dt_geneMod_perslab_merged$module_pres)[1])]

vec_mods_meta = unique(names(vec_mods))

vec_mods_meta = sort(vec_mods_meta)
```

```{r}
mat_embed_meta = sapply(vec_mods_meta, function(mod_meta) {
  
  vec_mods_in_mod_meta = vec_mods[names(vec_mods)==mod_meta]
  # weight by number of genes
  vec_weights = sapply(vec_mods_in_mod_meta, function(each_mod) {sum(dt_geneMod_perslab_merged$module_pres==each_mod)})
  vec_weights = vec_weights/sum(vec_weights)
  names(vec_weights) = vec_mods_in_mod_meta
  mat_tmp = mat_embed_scaled[,vec_mods_in_mod_meta]
  if (!is.null(dim(mat_tmp))) {
      # scale the columns of the embeddings matrix by number of genes
    mat_tmp = sapply(colnames(mat_tmp), function(each_mod) mat_tmp[,each_mod]*vec_weights[each_mod])
    mat_tmp %>% rowMeans
  } else { # if the metamodule has a single module
    mat_tmp
  }
})

mat_embed_meta[0:3,0:3]
```



## run linear models 

### combine embeddings and metadata

```{r}
dt_model_tmp <- data.table(dt_metadata, mat_embed_meta[dt_metadata$geo_accession,])
```

```{r}

lapply(colnames(mat_embed_meta), function(metamodule) {
  formula_tmp <- formula(paste0(metamodule, "~ fibrosis"))
  ##mat_model <- model.matrix(object = formula_tmp, data=dt_model_tmp)
  regr<-lm(formula = formula_tmp, dt_model_tmp)
  regr %>% broom::tidy() 
}) -> list_tbl_tmp 

names(list_tbl_tmp) = colnames(mat_embed_meta)

```


extract coefficients for the condition regressors 

```{r}

mat_out <- sapply(list_tbl_tmp, function(tbl_regr) {
    tbl_regr[tbl_regr$term=="fibrosis",colnames(tbl_regr) %in% c("estimate","std.error","statistic","p.value")] 
  }) %>% t

dt_condition_coef <- data.table("module"= rownames(mat_out), mat_out) 

```

adjust p-values

```{r}
dt_condition_coef[[paste0("p.value_",params$pAdjMethod)]]   <- p.adjust(p=dt_condition_coef$p.value,method=params$pAdjMethod)
```

how well do the betas correlate across modules with those for fibrosis in Gerhard?

```{r}
dt_fibrosis_coef_gerhard2018_metamod <- fread(paste0(dirWGCNA_tables, "liver_perslab_int_metamods_fibrosis_association_in_gerhard2018_200914.csv"))
```

```{r}
cor(dt_condition_coef[,estimate] %>% as.numeric, as.numeric(dt_fibrosis_coef_gerhard2018_metamod$estimate))
# [1] 0.92
```

## write out results

```{r}

fwrite(x=dt_condition_coef, file=paste0(dirWGCNA_tables, prefixData, "_", prefixRun, "_metamods_fibrosis_association_in_moylan2013_", flagDate,".csv"))
openxlsx::write.xlsx(x=dt_condition_coef,file = paste0(dirWGCNA_tables, prefixData, "_metamods_fibrosis__association_in_moylan2013_", flagDate,".xlsx"))
  

```
