---
title: 'Liver - embed modules in moylan 2013 (linear combination)'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

previous: liver_module_bulk_overlap_with_sc.Rmd

# Setup

## Load packages

```{r}
library("dplyr")
library("Matrix")
library("parallel")
library("readr")
library("openxlsx")
library("data.table")
library("here")


# source utility functions


source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))

# Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

# Set constants

randomSeed = 12345
set.seed(randomSeed)

pAdjustMethod = "BH"
pValThreshold = 0.05#params$pValThreshold

nRep = 10000 # bootstrap replicates
# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf

flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_renamed"
colCellCluster = "cell_cluster_pres"
```

## load data

bulk expression data 

```{r}
dt_datExpr <- fread(here("data", "moylan2013.norm.expr.qc.csv.gz"))
```

convert columns to numeric 

```{r}
strTmp <- colnames(dt_datExpr[,-"gene"])
dt_datExpr[, (strTmp) := lapply(.SD, as.numeric), .SDcols = strTmp]
```

```{r}
mat_datExpr <- as.matrix(dt_datExpr[,-1])
rownames(mat_datExpr) = dt_datExpr$gene
mat_datExpr[0:3,0:3]
```

Load modules dataframe

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz")
#head(dt_geneMod_perslab)

dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz")
#head(dt_geneMod_gerhard2018)

dt_geneMod_moylan2013 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_moylan2013_wgcna1_geneMod.csv.gz")
#head(dt_geneMod_moylan2013 )

dt_geneMod_merged = rbindlist(list(dt_geneMod_perslab,dt_geneMod_gerhard2018, dt_geneMod_moylan2013), fill=TRUE)
```

```{r}
dt_geneMod_merged[[colModule]] %>% unique -> vec_mods
vec_mods <- vec_mods[!is.na(vec_mods) & nchar(vec_mods)>0]
```

## compute embeddings

compute module embeddings as weighted sums of normalized expression data

<!-- filter out genes with zero sd -->

<!-- ```{r} -->
<!-- vec_sd = apply(mat_datExpr, MARGIN=1, FUN = sd) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(vec_sd==0) -->
<!-- # FALSE  TRUE  -->
<!-- # 22084   796  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- mat_datExpr <- mat_datExpr[vec_sd!=0,] -->
<!-- ``` -->

#### using normalized data, but scaling first (no centering, though!)

filter out genes with zero sd

```{r}
vec_sd = apply(mat_datExpr, MARGIN=1, FUN = sd)
```

```{r}
table(vec_sd==0)
# FALSE  TRUE 
# 22084   796
```

```{r}
mat_datExpr <- mat_datExpr[vec_sd!=0,]
```

now scale by the standard deviation (without centering)

```{r}
mat_datExpr %>% t %>% scale(., center=FALSE, scale=apply(.,2,sd,na.rm=T)) %>% t -> mat_datExpr_scaled
```

compute embeddings

<!-- unscaled -->

<!-- ```{r} -->
<!-- mat_embed <- sapply(vec_mods, function(mod) { -->

<!--   # get gene weights -->
<!--   condition = quote(dt_geneMod_merged[[colModule]]==mod) -->
<!--   vec_pkMs <- dt_geneMod_merged[eval(condition), pkMs,] -->

<!--   # normalize module gene weights to sum to 1 -->
<!--   vec_pkMs <- vec_pkMs/sum(vec_pkMs) -->

<!--   # find corresponding rows of the expression matrix -->
<!--   vec_idxRow <- match(dt_geneMod_merged[eval(condition), ..colGeneNames, ][[1]], rownames(mat_datExpr)) -->

<!--   # filter out mis-matches -->
<!--   vec_pkMs <- vec_pkMs[!is.na(vec_idxRow)] -->
<!--   vec_idxRow <- vec_idxRow[!is.na(vec_idxRow)] -->

<!--   # compute weighted sum of normalized expression -->
<!--   mat_datExpr[vec_idxRow,] %>% as.matrix -> mat_datExpr_sub -->

<!--   if (nrow(mat_datExpr_sub)==0) { -->
<!--     warning(paste0(mod, ": no module genes found in datExpr")) -->
<!--     return(rep(NA_real_, ncol(mat_datExpr_sub))) -->
<!--   } else if (nrow(mat_datExpr_sub)>0 & nrow(mat_datExpr_sub)<length(vec_pkMs)/2) { -->
<!--     warning(paste0(mod, ": only ", nrow(mat_datExpr_sub), " module genes detected")) -->
<!--   } -->
<!--   vec_pkMs %*% mat_datExpr_sub -->

<!-- }) -->

<!-- rownames(mat_embed) <- colnames(mat_datExpr) -->

<!-- mat_embed[0:4,0:4] -->
<!-- ``` -->

```{r}
mat_scaled_embed <- sapply(vec_mods, function(mod) {

  # get gene weights
  condition = quote(dt_geneMod_merged[[colModule]]==mod)
  vec_pkMs <- dt_geneMod_merged[eval(condition), pkMs,]

  # normalize module gene weights to sum to 1
  vec_pkMs <- vec_pkMs/sum(vec_pkMs)

  # find corresponding rows of the expression matrix
  vec_idxRow <- match(dt_geneMod_merged[eval(condition), ..colGeneNames, ][[1]], rownames(mat_datExpr_scaled))

  # filter out mis-matches
  vec_pkMs <- vec_pkMs[!is.na(vec_idxRow)]
  vec_idxRow <- vec_idxRow[!is.na(vec_idxRow)]

  # compute weighted sum of normalized expression
  mat_datExpr_scaled[vec_idxRow,] %>% as.matrix -> mat_datExpr_scaled_sub
  
  if (nrow(mat_datExpr_scaled_sub)==0) {
    warning(paste0(mod, ": no module genes found in datExpr"))
    return(rep(NA_real_, ncol(mat_datExpr_scaled_sub)))
  } else if (nrow(mat_datExpr_scaled_sub)>0 & nrow(mat_datExpr_scaled_sub)<(length(vec_pkMs)/2)) {
    warning(paste0(mod, ": only ", nrow(mat_datExpr_scaled_sub), " module genes detected"))
    return(rep(NA_real_, ncol(mat_datExpr_scaled_sub)))
  }
  return(vec_pkMs %*% mat_datExpr_scaled_sub)

})

rownames(mat_scaled_embed) <- colnames(mat_datExpr_scaled)

mat_scaled_embed[0:4,0:4]
```

## rescale the embeddings to standard deviation 1

```{r}
mat_scaled_embed_rescaled <- scale(mat_scaled_embed, center=FALSE, scale=apply(mat_scaled_embed,2,sd,na.rm=T))
```

<!-- ```{r} -->
<!-- mat_cor = cor(mat_embed, mat_scaled_embed) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- vec_diag = mat_cor[row(mat_cor)==col(mat_cor)] -->
<!-- median(vec_diag) -->
<!-- # [1] 0.9981947 -->
<!-- ``` -->

scaling makes no difference here, i.e. the microarray measurements are already scaled

<!-- plot the distributions of activity of a few modules -->

<!-- ```{r} -->
<!-- hist(mat_embed[,64],breaks = 50) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- mat_embed_log2  = log2(mat_embed+1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- hist(mat_embed_log2[,10],breaks = 50) -->
<!-- ``` -->


write out scaled weighted sum module embeddings matrix

```{r}
dt_scaled_embed_rescaled = data.table("GEO_accession" = rownames(mat_scaled_embed_rescaled), mat_scaled_embed_rescaled)

fwrite(dt_scaled_embed_rescaled, file="/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mat_scaled_embed_moylan2013.csv.gz", compress="gzip")
```

next:
liver_module_association_moylan2013.Rmd

