---
title: 'Liver - analyse subsets of cells'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

>>> REVISE <<<
this script
* subsets the Seurat integrated single-cell data (Macrophages, hepatocytes, stellate cells, cholangiocytes..)
* re-computes PCA and subcluster labels (make UMAP plots)
* return silhouette of subcluster labellings
* add new subcluster labellings to common metadata 

# Setup

## Load packages

```{r}

library("Seurat")
library("data.table")
library("Matrix")
library("parallel")
library("dplyr")
library("ggplot2")
library("here")
# library("cluster")
# library("parallelDist")
library("gerr")
library("GO.db")
library("gage") # provides GO genesets
```

## source utility functions

```{r}
source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

#options(datatable.WhenJisSymbolThenCallingScope=TRUE)
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)

npcs=40
res_primary = 0.3
res_secondary = 0.5
res_tertiary = 0.8

pval.adjust.method = "BH"
pValThreshold = 0.05#params$pValThreshold

# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

genesetLengthMin = 5
genesetLengthMax = 1000

# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

alpha =  0.25 # for GERR calling glmnet elasticnet, alpha \in [0,1], alpha==0 => ridge regression, alpha==1 => LASSO regression. Defaults to 0.5
        # ?gerr::regression_selected_pathways
```

generic project constants 

```{r}

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
#prefixOut <- "SCT"
prefixOut <- "subset"


dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_assoc"
colModule_gerhard2018 = "module"
colCellCluster = "cell_cluster_assoc"
colCellCluster_gerhard2018 = "cell_cluster"
```

```{r}
celltype <- "Endothelial-cells"
```

## load seurat objects with expression and meta-data

```{r}
#seuratObj_sub <- readRDS(here("output","liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj_macroph.RDS.gz"))
seuratObj <- readRDS(here("output","liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj.RDS.gz"))
```

```{r}
seuratObj_macParland <- readRDS(here("data", "macparland_seurat_obj3.RDS.gz"))
```

all perslab single-cell genes (as background for enrichment tests)
```{r}
vec_allGenes <- readRDS(here("data","liver_perslab_int_genesBackground.RDS"))
```

gene modules 

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod_merged_200325.csv.gz")
head(dt_geneMod_perslab)
```

bulk data

```{r}
dt_datExpr_gerhard <- fread(here("data", "liver_gerhard2018_norm.counts.csv.gz"))
```

```{r}
dt_metadata_gerhard <- fread(here("data", "liver_gerhard2018_patient_metadata_qc.csv.gz"))
```

```{r}
dt_mod_embeddings_gerhard <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_mod_embed_gerhard2018.csv.gz")
```


## compare endothelial cells in perslab and in MacParland


```{r}
vec_genesPlot = c("S100A8","S100A9","S100A12") # top genes in "honeydew3" macrophage module
```

```{r}
p1 <- VlnPlot(object = seuratObj, features = vec_genesPlot,  group.by = "cluster_perslab")
ggsave(plot=p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_calgranulin_vlnplots_perslab_cluster_perslab.pdf")))
```

```{r}
p2 <- VlnPlot(object = seuratObj_macParland, features = vec_genesPlot,  group.by = "Cluster_annot")
ggsave(plot=p2, filename = here("plots", paste0(prefixData, "_", prefixRun, "_calgranulin_vlnplots_macparland_Cluster_annot.pdf")))
```


Note the differences in proprocessing: 

* MacParland et al: SCT with regress out mito
* Perslab: SCT with regress out mito and CC difference %>% ComBat %>% preserve zeros

```{r}
seuratObj_sub_endothelial <- subset(seuratObj, subset = cluster_perslab_coarse == celltype)
```

```{r}
p_endo_1 <- VlnPlot(object=seuratObj_sub_endothelial, features =  vec_genesPlot,  group.by = "Endothelial-cells_subclust")

ggsave(plot=p_endo_1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_calgranulin_vlnplots_perslab_endothelial_subclust.pdf")))
```


```{r}
p_endo_2 <- VlnPlot(object=seuratObj_sub_endothelial, features =  "perslab_darkorange",  group.by = "Endothelial-cells_subclust")

ggsave(plot=p_endo_2, filename = here("plots", paste0(prefixData, "_", prefixRun, "_darkorange_vlnplots_perslab_endothelial_subclust.pdf")))
```

```{r}
DefaultAssay(seuratObj_sub_endothelial) <-  "integrated"
```

```{r}
seuratObj_sub_endothelial <- seuratObj_sub_endothelial %>% ScaleData %>% RunPCA(object=., seed.use = randomSeed) %>% RunUMAP(object = ., seed.use = randomSeed, dims = 1:npcs)
```

```{r}
p_endo_3 <- DimPlot(seuratObj_sub_endothelial,reduction = "umap",group.by ="Endothelial-cells_subclust")#, label = T)
```


>>> everything below is unchanged from macrophages <<<

Now to Gerhard

```{r}
condition = quote(dt_datExpr_gerhard$gene %in% vec_genesPlot)
mat_plot = dt_datExpr_gerhard[eval(condition),-1] %>% t
colnames(mat_plot) = dt_datExpr_gerhard$gene[dt_datExpr_gerhard$gene %in% vec_genesPlot]
dt_plot <- data.table(mat_plot)
dt_plot$run_accession = rownames(mat_plot)
```

```{r}
dt_plot$Diagnosis <- dt_metadata_gerhard$Diagnosis[match(dt_plot$run_accession, dt_metadata_gerhard$run_accession)]
```

### inflammation

```{r}
dt_plot_inflam_sub <- dt_plot[Diagnosis %in% c("Lob Inflam 1", "Lob Inflam 2", "NORMAL") ]
dt_plot_inflam_sub[,plot_color:=ifelse(Diagnosis == "Lob Inflam 2", "red", ifelse(Diagnosis == "Lob Inflam 1", "orange", "blue"))]
```

```{r}
p3 <- ggplot(data = dt_plot_inflam_sub, aes(x=S100A8, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p3, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A8_gerhard_inflam_histo.pdf")))
```

```{r}
p4 <- ggplot(data = dt_plot_inflam_sub, aes(x=S100A9, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p4, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A9_gerhard_inflam_histo.pdf")))

```

```{r}
p5 <- ggplot(data = dt_plot_inflam_sub, aes(x=S100A12, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p5, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A12_gerhard_inflam_histo.pdf")))

```

It seems the three calgranulin genes differ only in a few inflamed patients


### fibrosis

```{r}
dt_plot_fibr_sub <- dt_plot[grepl("Fibrosis|NORMAL", Diagnosis)]
#dt_plot_fibr_sub[,plot_color:=ifelse(Diagnosis == "Fibrosis4", "red", ifelse(Diagnosis == "Lob Inflam 1", "orange", "blue"))]
```

```{r}
p6 <- ggplot(data = dt_plot_fibr_sub, aes(x=S100A8, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p6, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A8_gerhard_fibrosis_histo.pdf")))
```

```{r}
p7 <- ggplot(data = dt_plot_fibr_sub, aes(x=S100A9, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p7, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A9_gerhard_fibrosis_histo.pdf")))

```

```{r}
p8 <- ggplot(data = dt_plot_fibr_sub, aes(x=S100A12, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p8, filename = here("plots", paste0(prefixData, "_", prefixRun, "_S100A12_gerhard_fibrosis_histo.pdf")))

```

There is even less of a pattern in fibrosis

### Look at the whole honeydew3 module

```{r}
dt_plot_honeydew3 <- dt_mod_embeddings_gerhard[,.(run_accession,perslab_honeydew3)]
dt_plot_honeydew3$Diagnosis <- dt_metadata_gerhard$Diagnosis[match(dt_plot_honeydew3$run_accession, dt_metadata_gerhard$run_accession)]
```

#### lob inflam 

```{r}
dt_plot_honeydew3_inflam_sub <- dt_plot_honeydew3[Diagnosis %in% c("Lob Inflam 1", "Lob Inflam 2", "NORMAL") ]
```


```{r}
p9 <- ggplot(data = dt_plot_honeydew3_inflam_sub, aes(x=perslab_honeydew3, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p9, filename = here("plots", paste0(prefixData, "_", prefixRun, "_honeydew3_gerhard_inflam_histo.pdf")))
```

#### fibrosis


```{r}
dt_plot_honeydew3_fibr_sub <- dt_plot_honeydew3[grepl("Fibrosis|NORMAL", Diagnosis)]
#dt_plot_fibr_sub[,plot_color:=ifelse(Diagnosis == "Fibrosis4", "red", ifelse(Diagnosis == "Lob Inflam 1", "orange", "blue"))]
```

```{r}
p10 <- ggplot(data = dt_plot_honeydew3_fibr_sub, aes(x=perslab_honeydew3, fill=Diagnosis)) + geom_histogram(bins=50, alpha = 0.5)
ggsave(plot=p10, filename = here("plots", paste0(prefixData, "_", prefixRun, "_honeydew3_gerhard_fibrosis_histo.pdf")))
```


<!-- ```{r} -->
<!-- dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod_merged_200325.csv.gz") -->
<!-- dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod_merged_200325.csv.gz") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_geneMod_gerhard2018$module <- paste0("gerhard2018_",dt_geneMod_gerhard2018$module) -->
<!-- ``` -->

### GERR gene set enrichment on subcluster DE ('marker') genes

```{r}
dt_markers <- openxlsx::read.xlsx(xlsxFile=here("output", paste0(prefixData, "_", prefixRun, "_", "subclust_markergenes.xlsx"))) %>% setDT 
```

```{r}
dt_markers <- dt_markers[celltype=="Macrophages"]
```

```{r}
list_vec_DEgenes <- lapply(unique(dt_markers$cluster), function(clust) {
  #vecOut <- -log10(dt_markers$p_val[dt_markers$cluster==clust])
  #names(vecOut) <- dt_markers$gene[dt_markers$cluster==clust]
  #vecOut <- sort(vecOut, decreasing = T)
  
  vec_out <- dt_markers$gene[dt_markers$cluster==clust]
  vec_out[1:(as.integer(length(vec_out)))]
  #vecOut
})
names(list_vec_DEgenes)<-unique(dt_markers$cluster)

```

#### Gene Ontology

download gene ontology genesets

```{r}
#data(bods)
#list_list_vec_genesetsGO <- bods(species = "human", pkg.name=NULL, id.type = "eg")
list_list_vec_genesetsGO <- go.gsets(species = "human", pkg.name=NULL, id.type = "eg")

list_vec_genesetsGO <- list_list_vec_genesetsGO[[1]]
```

filter out very large or small genesets

```{r}
list_vec_genesetsGO %>% sapply(., length) %>% summary
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
   #  1.0     2.0     6.0    98.8    24.0 19659.0
```

```{r}
list_vec_genesetsGO <- list_vec_genesetsGO[sapply(list_vec_genesetsGO, length) >= genesetLengthMin & sapply(list_vec_genesetsGO, length) <= genesetLengthMax]

length(list_vec_genesetsGO)
# [1] 12044
```

map the GO genesets from Entrez Gene to symbol

https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/egSymb
https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/eg2sym

```{r}
data(egSymb)

list_vec_genesetsGO <- lapply(list_vec_genesetsGO, function(vec_geneset) {
  vec_geneset <- eg2sym(eg=vec_geneset)
  vec_geneset[!is.na(vec_geneset)]
})
```

To get GO definitions we use GO.db package

```{r}
dt_GO = as.data.table(GO.db::GOTERM)
```

```{r}
gene_pathway_matrixGO = sapply(list_vec_genesetsGO, function(vec_geneset) {
  (vec_allGenes %in% vec_geneset) %>% as.numeric
}, simplify = T)

rownames(gene_pathway_matrixGO) <- vec_allGenes
  
dim(gene_pathway_matrixGO)
#[1] 36116  12044
```

convert to sparse matrix format 

```{r}
gene_pathway_matrixGO %>% as(., "sparseMatrix") -> gene_pathway_matrixGO
```

run regularised regression gene test 

```{r gaussianFit}
list_gaussResGO <- lapply(list_vec_DEgenes, function(vec_DEGenes) {
  regression_selected_pathways(gene_input=vec_DEGenes, 
                               gene_pathway_matrix = gene_pathway_matrixGO,
                               family="gaussian",
                               alpha=alpha)})
```

### gather the GO results in a data.table

```{r}
vec_clust <- rep(names(list_gaussResGO), sapply(list_gaussResGO, function(gaussRes) {gaussRes$selected_pathways_names %>% length}))
```

```{r}
vec_coef = lapply(list_gaussResGO, function(x) x[["selected_pathways_coef"]])  %>% unlist(use.names = F)
```

```{r}
vec_p.value_fisher = lapply(list_gaussResGO, function(x) x[["selected_pathways_fisher_pvalue"]])  %>% unlist(use.names = F)
```

```{r}
vec_GOid <- lapply(list_gaussResGO, function(x) x[["selected_pathways_names"]] 
                     %>%  names 
                     %>% substr(.,start=1, stop = 10))  %>% unlist(use.names = F)
```

```{r}
lapply(list_gaussResGO, function(x) {
  x[["selected_pathways_names"]] %>% names %>% substr(., start=12, stop=1000)
  })  %>% unlist(use.names = F) -> vec_GOterm
```

```{r}

sapply(vec_GOid, function(id) {
  dt_GO$Definition[dt_GO$go_id == id] %>% unique
}) %>% unlist(use.names = F) -> vec_GOdef 

```

```{r}

sapply(vec_GOid, function(id) {
  dt_GO$Ontology[dt_GO$go_id == id] %>% unique
}) %>% unlist(use.names = F) -> vec_ontology

```

```{r}
dt_resultsGO <- data.table(
  cluster = vec_clust,
  coef = vec_coef,
  p.value_fisher = vec_p.value_fisher,
  GO_id = vec_GOid,
  GO_term = vec_GOterm,
  GO_definition  = vec_GOdef,
  GO_ontology = vec_ontology
  )
```

KEGG analysis

Download KEGG genesets

```{r}
data(kegg.gs)
length(kegg.gs)
# [1] 177
# Hm it says 186 on the msigdb page http://software.broadinstitute.org/gsea/msigdb/genesets.jsp?collection=CP:KEGG

list_vec_genesetsKEGG = kegg.gs; rm(kegg.gs)
```

map the KEGG genesets from Entrez Gene to symbol

https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/egSymb
https://www.rdocumentation.org/packages/gage/versions/2.22.0/topics/eg2sym

```{r}
data(egSymb)

list_vec_genesetsKEGG <- lapply(list_vec_genesetsKEGG, function(vec_geneset) {
  vec_geneset <- eg2sym(eg=vec_geneset)
  vec_geneset[!is.na(vec_geneset)]
})
```

```{r}
gene_pathway_matrixKEGG = sapply(list_vec_genesetsKEGG, function(vec_geneset) {
  (vec_allGenes %in% vec_geneset) %>% as.numeric
}, simplify = T)

rownames(gene_pathway_matrixKEGG) <- vec_allGenes
  
dim(gene_pathway_matrixKEGG)
# [1] 3000  177
```

convert to sparse matrix format 

```{r}
gene_pathway_matrixKEGG %>% as(., "sparseMatrix") -> gene_pathway_matrixKEGG
```

run regularised regression gene test 

```{r gaussianFit}
list_gaussResKEGG <- lapply(list_vec_DEgenes, function(vec_DEGenes) {
  regression_selected_pathways(gene_input=vec_DEGenes, 
                               gene_pathway_matrix = gene_pathway_matrixKEGG,
                               family="gaussian",
                               alpha=alpha)})
```

### gather the KEGG results in a data.table

```{r}
vec_clust <- rep(names(list_gaussResKEGG), sapply(list_gaussResKEGG, function(gaussRes) {gaussRes$selected_pathways_names %>% length}))
```

```{r}
vec_coef = lapply(list_gaussResKEGG, function(x) x[["selected_pathways_coef"]])  %>% unlist(use.names = F)
```

```{r}
vec_p.value_fisher = lapply(list_gaussResKEGG, function(x) x[["selected_pathways_fisher_pvalue"]])  %>% unlist(use.names = F)
```

```{r}
vec_KEGGid <- lapply(list_gaussResKEGG, function(x) x[["selected_pathways_names"]] 
                     %>%  names 
                     %>% substr(.,start=1, stop = 8))  %>% unlist(use.names = F)
```

```{r}
lapply(list_gaussResKEGG, function(x) {
  x[["selected_pathways_names"]] %>% names %>% substr(.,start=10,stop=10000)
  })  %>% unlist(use.names = F) -> vec_KEGGterm
```

<!-- ```{r} -->

<!-- sapply(vec_KEGGid, function(id) { -->
<!--   dt_KEGG$Definition[dt_GO$go_id == id] %>% unique -->
<!-- }) %>% unlist(use.names = F) -> vec_GOdef -->

<!-- ``` -->

<!-- ```{r} -->

<!-- sapply(vec_GOid, function(id) { -->
<!--   dt_GO$Ontology[dt_GO$go_id == id] %>% unique -->
<!-- }) %>% unlist(use.names = F) -> vec_ontology -->

<!-- ``` -->

```{r}
dt_resultsKEGG <- data.table(
  cluster = vec_clust,
  coef = vec_coef,
  p.value_fisher = vec_p.value_fisher,
  KEGG_id = vec_KEGGid,
  KEGG_term = vec_KEGGterm
  )
```

### Examine module expression within celltype

<!-- ```{r} -->
<!-- list_seuratObj[[celltype]] %>% GetAssayData(object=., assay="SCT_combat") %>% as.data.table  -> dt_datExpr_celltype -->

<!-- dt_datExpr_celltype <- data.table("gene" = rownames(list_seuratObj[[celltype]]), dt_datExpr_celltype) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- mat_datExpr_celltype <- as.matrix(dt_datExpr_celltype[,-1]) -->
<!-- rownames(mat_datExpr_celltype) <- dt_datExpr_celltype[["gene"]] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- vec_mods_perslab = unique(dt_geneMod_perslab[[colModule]]) -->
<!-- vec_mods_perslab <- vec_mods_perslab[nchar(vec_mods_perslab)>0] -->
<!-- ``` -->

<!-- use **all** Gerhard modules - not only the ones associated with inflammation or fibrosis -->

<!-- ```{r} -->
<!-- vec_mods_gerhard2018 <- unique(dt_geneMod_gerhard2018[[colModule_gerhard2018]]) -->
<!-- vec_mods_gerhard2018 <- vec_mods_gerhard2018[nchar(vec_mods_gerhard2018)>0] -->
<!-- ``` -->

<!-- ```{r} -->

<!-- list_mat_embed_perslab <- lapply(vec_mods_perslab, function(module) { -->
<!--   vec_geneWeights <- dt_geneMod_perslab[[colGeneWeights]][dt_geneMod_perslab[[colModule]]==module] -->
<!--   names(vec_geneWeights) <- dt_geneMod_perslab[[colGeneNames]][dt_geneMod_perslab[[colModule]]==module] -->
<!--   vec_geneWeights <- vec_geneWeights/sum(vec_geneWeights) # normalise gene Weights -->
<!--   vec_geneWeights <- vec_geneWeights/sum(vec_geneWeights) # normalise gene Weights -->
<!--   vec_genes_intersect = intersect(names(vec_geneWeights),rownames(mat_datExpr_celltype)) -->
<!--   mat_out <- t(as.matrix(vec_geneWeights[vec_genes_intersect])) %*% mat_datExpr_celltype[vec_genes_intersect,] -->
<!--   return(mat_out) -->
<!-- }) -->
<!-- list_mat_embed_perslab %>% Reduce(x=., f = rbind) %>% t -> mat_embed_perslab -->
<!-- colnames(mat_embed_perslab) <- vec_mods_perslab -->

<!-- mat_embed_perslab[0:4,0:4] -->

<!-- ``` -->

<!-- ```{r} -->

<!-- list_mat_embed_gerhard2018<- lapply(vec_mods_gerhard2018, function(module) { -->
<!--   vec_geneWeights <- dt_geneMod_gerhard2018[[colGeneWeights]][dt_geneMod_gerhard2018[[colModule_gerhard2018]]==module] -->
<!--   names(vec_geneWeights) <- dt_geneMod_gerhard2018[[colGeneNames]][dt_geneMod_gerhard2018[[colModule_gerhard2018]]==module] -->
<!--   vec_geneWeights <- vec_geneWeights/sum(vec_geneWeights) # normalise gene Weights -->
<!--   vec_genes_intersect = intersect(names(vec_geneWeights),rownames(mat_datExpr_celltype)) -->
<!--   mat_out <- t(as.matrix(vec_geneWeights[vec_genes_intersect])) %*% mat_datExpr_celltype[vec_genes_intersect,] -->
<!--   return(mat_out) -->
<!-- }) -->
<!-- list_mat_embed_gerhard2018 %>% Reduce(x=., f = rbind) %>% t -> mat_embed_gerhard2018 -->
<!-- colnames(mat_embed_gerhard2018) <- vec_mods_gerhard2018 -->

<!-- mat_embed_gerhard2018[0:4,0:4] -->

<!-- ``` -->

<!-- ```{r} -->
<!-- mat_embed = cbind(mat_embed_gerhard2018, mat_embed_perslab) -->
<!-- ``` -->

<!-- ### add module embeddings to celltype metadata -->

<!-- ```{r} -->
<!-- condition = quote(!is.na(dt_metadata[[paste0(celltype,"_subclust")]])) -->
<!-- dt_metadata_celltype <- dt_metadata[eval(condition)] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_metadata_celltype <- data.table(dt_metadata_celltype, mat_embed) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_embed <- data.table("cell_id"=rownames(mat_embed),  -->
<!--                        dt_metadata_celltype[[paste0(celltype,"_subclust")]],  -->
<!--                        mat_embed) -->
<!-- colnames(dt_embed)[2] <- paste0(celltype,"_subclust") -->
<!-- ``` -->

<!-- ## compute celltype subcluster median embeddings -->

<!-- ```{r} -->
<!-- # /!\ -->
<!-- # NB: here we must manually type in the variable.. -->
<!-- colnames(dt_embed) <- gsub("-","_",colnames(dt_embed)) -->
<!-- dt_embed_clustMedian <- dt_embed[, lapply(.SD, median),  -->
<!--                                  by = .(Endothelial_cells_subclust),  -->
<!--                                  .SDcols = c(vec_mods_gerhard2018,vec_mods_perslab) ] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_embed_clustMedian_melted <- melt.data.table(data = dt_embed_clustMedian, id.vars=paste0(gsub("-","_",celltype),"_subclust"),  variable.name = "module", value.name = "median_embedding") -->
<!-- ``` -->

<!-- ## add mod embeddings to seurat object metadata as well  -->

<!-- ```{r} -->
<!-- list_seuratObj[[celltype]] <- Seurat::AddMetaData(object=list_seuratObj[[celltype]],metadata = data.frame(dt_embed,row.names = 1)) -->
<!-- colnames(list_seuratObj[[celltype]]@meta.data)[grep(gsub("-","_",celltype),colnames(list_seuratObj[[celltype]]@meta.data))]  <- paste0(celltype,"_subclust") -->
<!-- ``` -->

##  make violin plots of celltype- and gerhard modules in celltype subclusters

>>> THIS SECTION NEEDS UPDATING <<<

```{r}
vec_mods_celltype <- dt_geneMod_perslab[[colModule]][dt_geneMod_perslab[[colCellCluster]]==celltype] %>% unique
vec_mods_vlnplot = c(vec_mods_celltype, vec_mods_gerhard2018)
```

```{r}
#list_seuratObj[[1]]$macs_subclust <- factor(list_seuratObj[[1]]$macs_subclust, levels = 0:6, ordered=T)

plots <- Seurat::VlnPlot(object = list_seuratObj[[celltype]], features = vec_mods_vlnplot, sort = F, group.by = paste0(celltype,"_subclust"), combine = T)

ggsave(plot = plots, filename = here("plots", paste0(prefixData, "_",prefixRun, "_wgcna3_",celltype,"_subclust_modexpr_vlnplots.pdf")), width = 20, height=20)
```

```{r}
#list_seuratObj[[1]]$macs_subclust <- factor(list_seuratObj[[1]]$macs_subclust, levels = 0:6, ordered=T)

plots <- Seurat::VlnPlot(object = list_seuratObj[[celltype]], features = vec_mods_vlnplot, sort = F, group.by = "cluster_perslab", combine = T)

ggsave(plot = plots, filename = here("plots", paste0(prefixData, "_",prefixRun, "_wgcna3_",celltype,"_cluster_perslab_modexpr_vlnplots.pdf")), width = 20, height=25)
```

>>> END OF SECTION NEEDING UPDATING <<<

### macrophage subcluster 2

```{r}
seuratObj_sub = subset(x = seuratObj,
                       subset = cluster_perslab_coarse == "Macrophages")
```

```{r}
DefaultAssay(seuratObj_sub) <- "integrated"
```

```{r}
seuratObj_sub %>%  RunPCA(object=., npcs = npcs, seed.use = randomSeed) %>% RunUMAP(object = ., dims=1:npcs, seed.use = randomSeed) -> seuratObj_sub
```

```{r}
DimPlot(object = seuratObj_sub, group.by = "Macrophages_subclust", reduction = "umap", label = T)
```

```{r}
dt_tmp <- seuratObj_sub@meta.data %>% setDT
```

```{r}
dt_tmp[,lapply(perslab_honeydew3, mean), by=Macrophages_subclust]
```

seuratObj_macParland

seuratObj


# write out results 

GO results 

```{r}
saveMeta(savefnc=openxlsx::write.xlsx, x = dt_resultsGO, file=here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype, "_clustermarkers_GERR.xlsx")))

```

KEGG results 

```{r}

saveMeta(savefnc=openxlsx::write.xlsx, x = dt_resultsKEGG, file=here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype, "_clustermarkers_GERR.xlsx")))
  
```

<!-- ```{r} -->
<!-- fwrite(dt_metadata, here("data","liver_perslab_int_seurat_7_metadata_full.csv.gz")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- fwrite(dt_metadata_celltype, here("data",paste0(prefixData,"_", prefixRun, "_metadata_",celltype,".csv.gz"))) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- path_datExpr_celltype = paste0('/projects/jonatan/pub-perslab/18-liver-fred/output/liver_perslab_int_seurat_7_datExpr_combat_', celltype,'.csv.gz') -->
<!-- if (!file.exists(path_datExpr_celltype)) fwrite(x = dt_datExpr_celltype, file=path_datExpr_celltype) -->
<!-- ``` -->


<!-- save **all** subcluster seurat objects -->

<!-- ```{r} -->
<!-- for (celltype in names(list_seuratObj)) { -->
<!--   seuratObj_sub = list_seuratObj[[celltype]] -->
<!--   saveRDS(seuratObj_sub, here("output",paste0("liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj_", prefixOut, "_",celltype,".RDS.gz"))) -->
<!-- } -->
<!-- ``` -->

