---
title: 'Liver - analyse subsets of cells'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

this script
* subsets the Seurat integrated single-cell data (Macrophages, hepatocytes, stellate cells, cholangiocytes..)
* re-computes PCA and subcluster labels (make UMAP plots)
* return silhouette of subcluster labellings
* add new subcluster labellings to common metadata 

# Setup

## Load packages

```{r}

library("Seurat")
library("data.table")
library("Matrix")
library("parallel")
library("dplyr")
library("ggplot2")
library("here")
library("cluster")
library("parallelDist")

```

## source utility functions

```{r}
source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

#options(datatable.WhenJisSymbolThenCallingScope=TRUE)
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)

npcs=40
res_primary = 0.3
res_secondary = 0.5
res_tertiary = 0.8

pval.adjust.method = "BH"
pValThreshold = 0.05#params$pValThreshold

# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

genesetLengthMin = 5
genesetLengthMax = 1000

# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

alpha =  0.25 # for GERR calling glmnet elasticnet, alpha \in [0,1], alpha==0 => ridge regression, alpha==1 => LASSO regression. Defaults to 0.5
        # ?gerr::regression_selected_pathways
```

generic project constants 

```{r}

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
#prefixOut <- "SCT"
prefixOut <- "subset"


dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_assoc"
colModule_gerhard2018 = "module"
colCellCluster = "cell_cluster_assoc"
colCellCluster_gerhard2018 = "cell_cluster"
```

## load seurat object with expression data

```{r}
#seuratObj_sub <- readRDS(here("output","liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj_macroph.RDS.gz"))
seuratObj <- readRDS(here("output","liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj.RDS.gz"))
```

<!-- ## load metadata with -->

<!-- ```{r} -->
<!-- #dt_datExpr_combat <- fread(here("output", "liver_perslab_int_seurat_7_datExpr_combat_ensembl.csv.gz")) -->
<!-- dt_metadata <- fread(here("data","liver_perslab_int_seurat_7_metadata_full.csv.gz")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod_merged_200325.csv.gz") -->
<!-- dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod_merged_200325.csv.gz") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_geneMod_gerhard2018$module <- paste0("gerhard2018_",dt_geneMod_gerhard2018$module) -->
<!-- ``` -->

## subset integrated data and run standard clustering workflow on integrated data

```{r}
Idents(seuratObj) <- seuratObj$cluster_perslab_coarse
DefaultAssay(seuratObj) <- "integrated"
```

```{r}
vec_celltype = c("Cholangiocytes", "Endothelial-cells", "Hepatic-stellate-cells", "Hepatocytes",  "Macrophages", "NK-like-cells", "T-cells-alpha-beta","T-cells-gamma-delta")

list_seuratObj <- lapply(vec_celltype, function(id) subset(seuratObj, idents=id))

names(list_seuratObj) <- vec_celltype
```

subset PCA, UMAP and clustering

```{r}
list_seuratObj <- lapply(list_seuratObj, function(seuratObj_sub){

  # dimensional reduction and subclustering
  seuratObj_sub <- Seurat::RunPCA(seuratObj_sub, 
                                assay = "integrated",
                                npcs = npcs, 
                                seed.use=randomSeed,
                                verbose = FALSE)
 seuratObj_sub <- FindNeighbors(object = seuratObj_sub, dims= 1:npcs, verbose = T)
 seuratObj_sub <- FindClusters(object = seuratObj_sub, verbose = T,resolution =res_primary, random.seed = randomSeed) 
 seuratObj_sub <- RunUMAP(object=seuratObj_sub,  
                               dims=1:npcs, 
                               reduction = "pca",
                               assay=  "integrated", 
                               seed.use = randomSeed)
 
 
})

names(list_seuratObj) <- vec_celltype
```

add a second and third round of clustering

```{r}
list_seuratObj <- lapply(list_seuratObj, function(seuratObj_sub){
  seuratObj_sub <- FindClusters(object = seuratObj_sub, verbose = T,resolution =res_secondary, random.seed = randomSeed) 
  seuratObj_sub <- FindClusters(object = seuratObj_sub, verbose = T,resolution =res_tertiary, random.seed = randomSeed)
})

names(list_seuratObj) <- vec_celltype
```

subcluster UMAP plots
colored by subcluster

```{r}
for (celltype in names(list_seuratObj)) {
  
  seuratObj_sub = list_seuratObj[[celltype]]
  
  # plot subclusters
  p1 <-DimPlot(seuratObj_sub, 
             reduction = 'umap', 
             group.by = paste0(DefaultAssay(seuratObj_sub),'_snn_res.',res_secondary),
             label = TRUE)
  saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype,"_UMAP_clust",res_primary,".pdf")), width = 20, height=12)
  
  p1 <-DimPlot(seuratObj_sub, 
           reduction = 'umap', 
           group.by = paste0(DefaultAssay(seuratObj_sub),'_snn_res.',res_secondary),
           label = TRUE)
  saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype,"_UMAP_clust",res_secondary,".pdf")), width = 20, height=12)
  p1 <-DimPlot(seuratObj_sub, 
           reduction = 'umap', 
           group.by = paste0(DefaultAssay(seuratObj_sub),'_snn_res.',res_tertiary),
           label = TRUE)
  saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype,"_UMAP_clust",res_tertiary,".pdf")), width = 20, height=12)

}
```

```{r}
for (celltype in names(list_seuratObj)) {
  
  seuratObj_sub = list_seuratObj[[celltype]]
  
  # plot subclusters
  p1 <-DimPlot(seuratObj_sub, 
             reduction = 'umap', 
             group.by = "sample_ID",
             label = FALSE,
             pt.size=2)

  saveMeta(savefnc= ggsave, plot= p1, filename = here("plots", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype,"_UMAP_sample_ID.pdf")), width = 20, height=12)

}
```

compute cluster silhouette

```{r}
cluster_res = c(0.5,0.5,0.3,0.3,0.3,0.3,0.5,0.5)
names(cluster_res) = names(list_seuratObj)
```

```{r}

list_seuratObj <- lapply(1:length(list_seuratObj), function(i) {
  
  
  seuratObj_sub = list_seuratObj[[i]]
  
  DefaultAssay(seuratObj_sub) <- "integrated"
   
  d <- parDist(Embeddings(seuratObj_sub, reduction = "pca")[,1:npcs])
  print(names(list_seuratObj)[i])
  ident<-as.numeric(seuratObj_sub@meta.data[[paste0(DefaultAssay(seuratObj_sub),'_snn_res.',cluster_res[i])]])
  
  names(ident)<-names(Idents(seuratObj_sub))
  
  s2 <- silhouette(ident, d)
  
  # data.frame(cellscore=s2[,3], ident=as.character(ident)) %>%
  #   group_by(ident) %>% summarise_at(vars(cellscore), list(~median(., na.rm=TRUE))) -> silscore
  # 
  data.frame(silscore=s2[,3], row.names = names(ident)) -> df_sillow
  
  colnames(df_sillow) = paste0(names(list_seuratObj)[i], "_subclust_silhouette")
    
  seuratObj_sub <- AddMetaData(object = seuratObj_sub, metadata = df_sillow)
  
  seuratObj_sub
  
})

names(list_seuratObj) <- c("Cholangiocytes", "Endothelial-cells", "Hepatic-stellate-cells", "Hepatocytes","Macrophages","NK-like-cells","T-cells-alpha-beta","T-cells-gamma-delta")
```

get subcluster marker genes

```{r}

list_dt_markers_sub = list()

for (i in 1:length(list_seuratObj)) {
  
  celltype= names(list_seuratObj)[i]
  
  seuratObj_sub = list_seuratObj[[celltype]]
  
  Seurat::DefaultAssay(seuratObj_sub) <-  "SCT_combat"
  # get subcluster marker genes
  clusters = seuratObj_sub@meta.data[[paste0("integrated_snn_res.", cluster_res[i])]] %>% table %>% names 
  
  Seurat::Idents(seuratObj_sub) <-seuratObj_sub@meta.data[[paste0("integrated_snn_res.", cluster_res[i])]]

  list_iterable = list("X"=clusters)
  fun = function(cluster) {tryCatch({
    FindMarkers(seuratObj_sub ,  
            #cells.1=colnames(seurat_obj)[Idents(seurat_obj)==cluster],
            #cells.2=NULL,
            ident.1 = cluster,
            only.pos = T,
            #ident.2 = clusters[clusters!=cluster],
            test.use  ="wilcox",
            max.cells.per.ident=1000,
            random.seed=randomSeed,
            #latent.vars = if (!is.null(merge_specify) | !is.null(merge_group_IDs)) "sample_ID" else NULL,
            verbose = T)
            }, 
            error = function(err) {
              NA_character_
              warning(paste0("findmarkers failed for ", cluster, " with error "))
            })}
  
    list_markers=NULL
    list_markers <- lapply(FUN=fun, "X"=list_iterable[[1]])
    
    # add the gene and cluster as a column
    list_markers <- mapply(function(df_markers, cluster) {
      if (!all(sapply(df_markers, is.na))) {
        cbind("gene" = rownames(df_markers), "cluster"=rep(cluster, nrow(df_markers)), df_markers)
      } else {
        NA_character_
      }
    },
    df_markers=list_markers, 
    cluster=names(table(Idents(seuratObj_sub))), SIMPLIFY=F)
    
    list_markers <- list_markers[!sapply(list_markers, function(markers) all(is.na(markers)))]
    df_markers <- Reduce(x=list_markers, f=rbind)
    rownames(df_markers) <- NULL
    
    dt_markers_tmp <- data.table("celltype"=celltype, df_markers) 
    list_dt_markers_sub[[celltype]] <- dt_markers_tmp
    # saveMeta(savefnc=openxlsx::write.xlsx, x = df_markers, file=here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_", celltype, "_clustermarkers.xlsx")))
    
}

dt_markers <- data.table::rbindlist(list_dt_markers_sub)
```

add subclustering labels and silhouette to full seuratObj

```{r}
for (i in 1:length(list_seuratObj)){
  
  celltype = names(list_seuratObj)[i]
  seuratObj_sub = list_seuratObj[[celltype]]
  idx = na.omit(match(colnames(seuratObj_sub), colnames(seuratObj)))
  col_header <- paste0(celltype, "_subclust")
  col_header_sil <- paste0(celltype, "_subclust_silhouette")
  
  vec_subclust <- rep(NA_character_, ncol(seuratObj))
  vec_subclust_silhouette <- rep(NA_character_, ncol(seuratObj))
  
  vec_subclust[idx] <- seuratObj_sub@meta.data[[paste0(DefaultAssay(seuratObj_sub),"_snn_res.", cluster_res[i])]]
  vec_subclust_silhouette[idx] <- seuratObj_sub@meta.data[[col_header_sil]]
  
  seuratObj <- AddMetaData(object=seuratObj, metadata = vec_subclust, col.name= col_header)
  seuratObj <- AddMetaData(object=seuratObj, metadata = vec_subclust_silhouette, col.name = col_header_sil)
}
```

# write out results 

marker genes

```{r}
openxlsx::write.xlsx(x=dt_markers, file = here("output",paste0(prefixData, "_",prefixRun,  "_subclust_markergenes.xlsx")))
```

seurat object

```{r}
saveRDS(seuratObj, here("output",paste0(prefixData, "_",prefixRun,  "_SCTint_perslab_labels_seuratObj.RDS.gz")), compress="gzip")
```

