---
title: 'Liver - assocication output postprocessing'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
  pAdjMethod: "BH"
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

Previous: 
liver_module_association_moylan2013.Rmd
liver_module_association_gerhard2018.Rmd

# Setup

## Load packages

```{r}

#ipak(c("Seurat", "dplyr", "ggplot2", "Matrix", "parallel", "readr", "tidyr"))
#library("Seurat")

library("data.table")
library("Matrix")
library("parallel")
#library("tidyverse")
library("here")
library("magrittr")
#library("AUCell")
#library("ggplot2")
#library("simpleboot")
#library("boot")
#library("car")#https://socialsciences.mcmaster.ca/jfox/Books/Companion/appendices/Appendix-Bootstrapping.pdf
# source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))

# Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")

R=1e3-1
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

# Set constants


randomSeed = 12345
set.seed(randomSeed)
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 


prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_renamed"
colCellCluster = "cell_cluster_pres"
# naming convention:
# <prefixData>_<prefixRun>_<prefixOut>_<outputType>_<params>_<params>_..._<date>.pdf

```

## load data


load data.table with information on module and cell type of origin

```{r}
path_geneMod_perslab <- paste0(dirWGCNA_tables,prefixData, "_",prefixRun_WGCNA,"_geneMod.csv.gz")
dt_geneMod_perslab <- fread(path_geneMod_perslab)
#head(dt_geneMod_perslab)

path_geneMod_gerhard2018 <- paste0(dirWGCNA_tables,"liver_gerhard2018_wgcna3_geneMod.csv.gz")
dt_geneMod_gerhard2018 <- fread(path_geneMod_gerhard2018)
#head(dt_geneMod_gerhard2018)
# 
# path_geneMod_moylan2013 <- paste0(dirWGCNA_tables,"liver_moylan2013_wgcna1_geneMod.csv.gz")
# dt_geneMod_moylan2013 <- fread(path_geneMod_moylan2013)
# #head(dt_geneMod_moylan2013)
```



## bootstrap association stats

```{r}
# dt_bootOut_stats_moylan2013 = fread(paste0(dirWGCNA_tables, prefixData, "_mod_fibrosis_assoc_in_moylan2013_201020.csv"))

dt_bootOut_stats_gerhard2018 <- fread(paste0(dirWGCNA_tables, prefixData,"_mod_assoc_in_gerhard2018_201101.csv"))
```


fix previous slipup in string renaming
```{r}
vec_number = str_split(string = dt_bootOut_stats_gerhard2018$module, pattern = "-") %>% sapply(., FUN=function(vec) vec[length(vec)])

vec_rest = gsub("-\\d+$", "", dt_bootOut_stats_gerhard2018$module)

dt_bootOut_stats_gerhard2018$module = paste0(vec_rest, "_", vec_number)
```

# analysis

<!-- ## check how well the fibrosis associations match between moylan and gerhard -->

<!-- #### moylan fibrosis severity versus gerhard fibrosis severity -->

<!-- ```{r} -->
<!-- dt_bootOut_stats_gerhard2018_fibrosis_severity = dt_bootOut_stats_gerhard2018[condition=="fibrosis_severity"] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all.equal(dt_bootOut_stats_gerhard2018_fibrosis_severity$module, dt_bootOut_stats_moylan2013$module) -->
<!-- #[1] TRUE  -->
<!-- ``` -->

<!-- #### all modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis_severity$estimate_original, dt_bootOut_stats_moylan2013$estimate_original, use = "pairwise.complete.obs") -->
<!-- # [1] -0.06896993 -->
<!-- ``` -->

<!-- #### perslab modules -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis_severity[!grepl("gerhard|moylan", module), estimate_original], dt_bootOut_stats_moylan2013[!grepl("gerhard|moylan", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] 0.02471545 -->
<!-- ``` -->

<!-- #### gerhard modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis_severity[grepl("gerhard", module), estimate_original], dt_bootOut_stats_moylan2013[grepl("gerhard", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] -0.5218904 -->
<!-- ``` -->

<!-- #### moylan modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis_severity[grepl("moylan", module), estimate_original], dt_bootOut_stats_moylan2013[grepl("moylan", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] -0.2899675 -->
<!-- ``` -->

<!-- gerhard fibrosis versus moylan fibrosis -->

<!-- ```{r} -->
<!-- dt_bootOut_stats_gerhard2018_fibrosis = dt_bootOut_stats_gerhard2018[condition=="fibrosis"] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all.equal(dt_bootOut_stats_gerhard2018_fibrosis$module, dt_bootOut_stats_moylan2013$module) -->
<!-- #[1] TRUE  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis$estimate_original, dt_bootOut_stats_moylan2013$estimate_original, use = "pairwise.complete.obs") -->
<!-- # [1] 0.1231598 -->
<!-- ``` -->

<!-- #### perslab modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis[!grepl("gerhard|moylan", module), estimate_original], dt_bootOut_stats_moylan2013[!grepl("gerhard|moylan", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] 0.1908229 -->
<!-- ``` -->

<!-- #### gerhard modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis[grepl("gerhard", module), estimate_original], dt_bootOut_stats_moylan2013[grepl("gerhard", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] 0.2086069 -->
<!-- ``` -->

<!-- #### moylan modules  -->

<!-- ```{r} -->
<!-- cor(dt_bootOut_stats_gerhard2018_fibrosis[grepl("moylan", module), estimate_original], dt_bootOut_stats_moylan2013[grepl("moylan", module), estimate_original], use = "pairwise.complete.obs") -->
<!-- # [1] 0.05892578 -->
<!-- ``` -->


### adjust p-values for multiple testing separately for each condition



gerhard

```{r}
list_dt_bootOut_stats_gerhard2018_sub_BH = lapply(unique(dt_bootOut_stats_gerhard2018$condition), function(condition) {
  print(condition)
  filterCondition <- dt_bootOut_stats_gerhard2018$condition == condition
  dt_tmp <- dt_bootOut_stats_gerhard2018 %>% dplyr::filter(., filterCondition) 
  
  dt_tmp$p.value_percentile_BH= p.adjust(dt_tmp$p.value_percentile, method = "BH")#params$pAdjMethod)
  dt_tmp$p.value_norm_BH = p.adjust(dt_tmp$p.value_norm, method = "BH")#params$pAdjMethod)
  return(dt_tmp)
})

dt_bootOut_stats_gerhard2018_BH = data.table::rbindlist(list_dt_bootOut_stats_gerhard2018_sub_BH)
```

<!-- moylan -->

<!-- ```{r} -->
<!-- dt_bootOut_stats_moylan2013_BH = dt_bootOut_stats_moylan2013[!grepl("gerhard", module)] -->

<!-- dt_bootOut_stats_moylan2013_BH[,p.value_percentile_BH := p.adjust(p.value_percentile, method = params$pAdjMethod)] -->

<!-- dt_bootOut_stats_moylan2013_BH[,p.value_norm_BH := p.adjust(p.value_norm, method = params$pAdjMethod)] -->
<!-- ``` -->

```{r}
dt_bootOut_stats_gerhard2018_BH = data.table(dataset="gerhard2018", 
                                             dt_bootOut_stats_gerhard2018_BH)
# dt_bootOut_stats_moylan2013_BH = data.table(dataset="moylan2013",
#                                             dt_bootOut_stats_moylan2013_BH)
```

<!-- ```{r} -->
<!-- dt_bootOut_stats_combined_summary = data.table::rbindlist(list(dt_bootOut_stats_gerhard2018_BH, dt_bootOut_stats_moylan2013_BH)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dt_bootOut_stats_gerhard2018_BH_summary <- dt_bootOut_stats_gerhard2018_BH[,list(dataset, condition, module, estimate_original, estimate_bootMean, estimate_bootMed, confInt_lower_percentile, confInt_upper_percentile, p.value_percentile, p.value_percentile_BH)] -->
<!-- ``` -->

```{r}
vec_mods_fibrosis_gerhard = unique(dt_bootOut_stats_gerhard2018_BH[p.value_percentile_BH<params$pValThreshold & condition=="fibrosis",module])

vec_mods_fibrosis_severity_gerhard = unique(dt_bootOut_stats_gerhard2018_BH[p.value_percentile_BH<params$pValThreshold & condition=="fibrosis_severity",module])

vec_mods_lob_inflam_gerhard = unique(dt_bootOut_stats_gerhard2018_BH[p.value_percentile_BH<params$pValThreshold & condition=="lob_inflam",module])
```

<!-- ```{r} -->
<!-- vec_mods_fibrosis_severity_moylan = unique(dt_bootOut_stats_moylan2013_BH[p.value_percentile_BH<params$pValThreshold,module]) -->
<!-- ``` -->

## annotate in genemod tables

```{r}

# perslab modules
dt_geneMod_perslab$cell_cluster_fibrosis_gerhard2018 = ifelse(dt_geneMod_perslab$module_renamed %in% vec_mods_fibrosis_gerhard, dt_geneMod_perslab$cell_cluster_pres, NA_character_)

dt_geneMod_perslab$module_fibrosis_gerhard2018 = ifelse(dt_geneMod_perslab$module_renamed %in% vec_mods_fibrosis_gerhard, dt_geneMod_perslab$module_renamed, NA_character_)

dt_geneMod_perslab$cell_cluster_lob_inflam_gerhard2018 = ifelse(dt_geneMod_perslab$module_renamed %in% vec_mods_lob_inflam_gerhard, dt_geneMod_perslab$cell_cluster_pres, NA_character_)

dt_geneMod_perslab$module_lob_inflam_gerhard2018 = ifelse(dt_geneMod_perslab$module_renamed %in% vec_mods_lob_inflam_gerhard, dt_geneMod_perslab$module_renamed, NA_character_)
# 
# dt_geneMod_perslab$cell_cluster_fibrosis_moylan2013 = ifelse(gsub("-", "_", dt_geneMod_perslab$module_renamed) %in% vec_mods_fibrosis_moylan, dt_geneMod_perslab$cell_cluster_pres, NA_character_)
# 
# dt_geneMod_perslab$module_fibrosis_moylan2013 = ifelse(gsub("-", "_", dt_geneMod_perslab$module_renamed) %in% vec_mods_fibrosis_moylan, dt_geneMod_perslab$module_renamed, NA_character_)

# gerhard2018 modules

dt_geneMod_gerhard2018$cell_cluster_fibrosis_gerhard2018 = ifelse(dt_geneMod_gerhard2018$module_renamed %in% vec_mods_fibrosis_gerhard, dt_geneMod_gerhard2018$cell_cluster_pres, NA_character_)

dt_geneMod_gerhard2018$module_fibrosis_gerhard2018 = ifelse( dt_geneMod_gerhard2018$module_renamed %in% vec_mods_fibrosis_gerhard, dt_geneMod_gerhard2018$module_renamed, NA_character_)

dt_geneMod_gerhard2018$cell_cluster_lob_inflam_gerhard2018 = ifelse( dt_geneMod_gerhard2018$module_renamed %in% vec_mods_lob_inflam_gerhard, dt_geneMod_gerhard2018$cell_cluster_pres, NA_character_)

dt_geneMod_gerhard2018$module_lob_inflam_gerhard2018 = ifelse( dt_geneMod_gerhard2018$module_renamed %in% vec_mods_lob_inflam_gerhard, dt_geneMod_gerhard2018$module_renamed, NA_character_)

# dt_geneMod_gerhard2018$cell_cluster_fibrosis_moylan2013 = NA_character_
# 
# dt_geneMod_gerhard2018$module_fibrosis_moylan2013 = NA_character_

# moylan2013 modules

# dt_geneMod_moylan2013$cell_cluster_fibrosis_gerhard2018 = NA_character_
# 
# dt_geneMod_moylan2013$module_fibrosis_gerhard2018 = NA_character_
# 
# dt_geneMod_moylan2013$cell_cluster_lob_inflam_gerhard2018 = NA_character_
# 
# dt_geneMod_moylan2013$module_lob_inflam_gerhard2018 = NA_character_
# 
# dt_geneMod_moylan2013$cell_cluster_fibrosis_moylan2013 = ifelse(gsub("-", "_", dt_geneMod_moylan2013$module_renamed) %in% vec_mods_fibrosis_moylan, dt_geneMod_moylan2013$cell_cluster_pres, NA_character_)
# 
# dt_geneMod_moylan2013$module_fibrosis_moylan2013 = ifelse(gsub("-", "_", dt_geneMod_moylan2013$module_renamed) %in% vec_mods_fibrosis_moylan, dt_geneMod_moylan2013$module_renamed, NA_character_)

```

```{r}
all.equal(colnames(dt_geneMod_perslab),colnames(dt_geneMod_gerhard2018))
#[1] TRUE

#all.equal(colnames(dt_geneMod_perslab),colnames(dt_geneMod_moylan2013))
#[1] TRUE
```

## write files to disk 

## association datatables 


```{r}
fwrite(x=dt_bootOut_stats_gerhard2018_BH, file=paste0(dirWGCNA_tables, prefixData, "_mod_assoc_in_gerhard2018_BH_", flagDate,".csv"))

openxlsx::write.xlsx(dt_bootOut_stats_gerhard2018_BH, file=paste0(dirWGCNA_tables, prefixData, "_mod_assoc_in_gerhard2018_BH_", flagDate,".xlsx"))
```

<!-- ```{r} -->
<!-- fwrite(x=dt_bootOut_stats_gerhard2018_BH_summary, file=paste0(dirWGCNA_tables, prefixData, "_mod_assoc_in_gerhard2018_Bsummary_", flagDate,".csv")) -->

<!-- openxlsx::write.xlsx(x=dt_bootOut_stats_gerhard2018_BH_summary,file = paste0(dirWGCNA_tables, prefixData, "_mod_assoc_summary_", flagDate,".xlsx")) -->
<!-- ``` -->


## geneMod tables 

```{r}
fwrite(dt_geneMod_perslab, file="/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz", compress="gzip")

# openxlsx::write.xlsx(x = dt_geneMod_perslab, file = "/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.xlsx")

fwrite(dt_geneMod_gerhard2018, file = "/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz", compress="gzip")

# openxlsx::write.xlsx(x = dt_geneMod_gerhard2018, file = "/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.xlsx")
# 
# fwrite(dt_geneMod_moylan2013, file = "/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_moylan2013_wgcna1_geneMod.csv.gz", compress="gzip")

# openxlsx::write.xlsx(x = dt_geneMod_moylan2013, file = "/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_moylan2013_wgcna1_geneMod.xlsx")
```

