---
title: 'Liver - compute embeddings of single cell modules in MacParland data and add to MacParland seurat object'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

previous: ran WGCNA on perslab, moylan and gerhard separately



# Setup

## Load packages

```{r}
library("dplyr")
library("Matrix")
library("parallel")
library("readr")
library("openxlsx")
library("data.table")
library("here")

## source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))


## Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

## Set constants


randomSeed = 12345
set.seed(randomSeed)

pAdjustMethod = "BH"
pValThreshold = 0.05#params$pValThreshold

nRep = 10000 # bootstrap replicates
# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf

flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixOut <- "SCTint"

prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_renamed"
colCellCluster = "cell_cluster_pres"

```

## load data

Seurat object contains expression data and cell + patient-level metadata

```{r}
path_seuratObj_macparland <- here("data", "macparland_seurat_obj3.RDS.gz")

seuratObj_macparland <- load_obj(path_seuratObj_macparland)
```

Load modules dataframe

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz")
head(dt_geneMod_perslab)

dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz")
# head(dt_geneMod_gerhard2018)
```


```{r}
DefaultAssay(seuratObj_macparland) <- "SCT"
```

<!-- ```{r} -->
<!-- seuratObj <- ScaleData(seuratObj, features = rownames(seuratObj)) -->
<!-- ``` -->

```{r}
mat_datExpr_macparland <- as.matrix(GetAssayData(object = seuratObj_macparland, assay = "SCT", slot="data"))
rownames(mat_datExpr_macparland) <- rownames(seuratObj_macparland)
mat_datExpr_macparland[0:4,0:4]
```

<!-- ```{r} -->
<!-- vec_sd = apply(mat_datExpr_macparland, MARGIN=1, FUN = sd) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(vec_sd==0) -->
<!-- # FALSE  TRUE  -->
<!-- # 18686    29  -->
<!-- ``` -->

only keep genes that are part of modules

```{r}
mat_datExpr_macparland = mat_datExpr_macparland[rownames(mat_datExpr_macparland) %in% dt_geneMod_perslab$genes,]

dim(mat_datExpr_macparland)
# [1] 6038 8444
```


now scale (without centering)

```{r}
mat_datExpr_macparland %>% t %>% scale(., center=FALSE, scale=apply(.,2,sd,na.rm=T)) %>% t -> mat_datExpr_macparland_scaled
```


### compute module embeddings in sc data

```{r}
dt_geneMod_merged = dt_geneMod_perslab

dt_geneMod_merged[[colModule]] %>% unique -> vec_mods

vec_mods <- vec_mods[!is.na(vec_mods) & nchar(vec_mods)>0]
```

```{r}
mat_scaled_embed_macparland <- sapply(vec_mods, function(mod) {
 
  # get gene weights
  condition = quote(dt_geneMod_merged[[colModule]]==mod)
  vec_pkMs <- dt_geneMod_merged[eval(condition), pkMs,]

  # normalize module gene weights to sum to 1
  vec_pkMs <- vec_pkMs/sum(vec_pkMs)

  # find corresponding rows of the expression matrix
  vec_idxRow <- match(dt_geneMod_merged[eval(condition), ..colGeneNames, ][[1]], rownames(mat_datExpr_macparland_scaled))

  # filter out mis-matches
  vec_pkMs <- vec_pkMs[!is.na(vec_idxRow)]
  vec_idxRow <- vec_idxRow[!is.na(vec_idxRow)]
  
  #print(length(vec_pkMs)/2)
  # compute weighted sum of normalized and scaled expression
  mat_datExpr_macparland_scaled[vec_idxRow,] %>% as.matrix -> mat_datExpr_macparland_scaled_sub
  
  if (nrow(mat_datExpr_macparland_scaled_sub)==0) {
  warning(paste0(mod, ": no module genes found in datExpr"))
  return(rep(NA_real_, ncol(mat_datExpr_macparland_scaled_sub)))
  } else if (nrow(mat_datExpr_macparland_scaled_sub)>0 & nrow(mat_datExpr_macparland_scaled_sub)<(length(vec_pkMs)/2)) {
    warning(paste0(mod, ": only ", nrow(mat_datExpr_macparland_scaled_sub), " module genes detected"))
    return(rep(NA_real_, ncol(mat_datExpr_macparland_scaled_sub)))
  }
  return(vec_pkMs %*% mat_datExpr_macparland_scaled_sub)

})

rownames(mat_scaled_embed_macparland) <- colnames(mat_datExpr_macparland_scaled)

mat_scaled_embed_macparland[0:4,0:4]
```

scale embeddings 


## rescale the embeddings to standard deviation 1

```{r}
mat_scaled_embed_macparland_rescaled <- scale(mat_scaled_embed_macparland, center=FALSE, scale=apply(mat_scaled_embed_macparland,2,sd,na.rm=T))
```

## add new embeddings to Seurat object

<!-- ### first remove the old -->

<!-- ```{r} -->
<!-- vec_mods_old <- unique(dt_geneMod_merged$module_filter) -->
<!-- for (mod_old in vec_mods_old){  -->
<!--   if (mod_old %in% colnames(seuratObj_macparland@meta.data)) { -->
<!--     seuratObj_macparland[[mod_old]] <- NULL   -->
<!--   } -->
<!-- } -->
<!-- ``` -->

```{r}
seuratObj_macparland <- AddMetaData(object=seuratObj_macparland, metadata = data.frame(row.names=rownames(mat_scaled_embed_macparland_rescaled), mat_scaled_embed_macparland_rescaled))
```

```{r}
saveRDS(object = seuratObj_macparland, file = path_seuratObj_macparland, compress = "gzip")
```

next: liver_module_filter_prune.Rmd


