---
title: 'Liver - QC bulk data'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

references: 

* Moylan CA, Pang H, Dellinger A, Suzuki A et al. Hepatic gene expression profiles differentiate presymptomatic patients with mild versus severe nonalcoholic fatty liver disease. Hepatology 2014 Feb;59(2):471-82.  https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE31803

# Setup

## Load packages

```{r}
library("data.table")
library("Matrix")
library("parallel")
library("dplyr")
library("ggplot2")
library("here")
library("corrr")
library("irlba")

```

## source utility functions

```{r}
source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()
```

## Set constants

```{r}
randomSeed = 12345
set.seed(randomSeed)

pval.adjust.method = "BH"
pValThreshold = 0.05#params$pValThreshold
```


generic project constants 

```{r}
#prefixData <- "aizarani"
prefixData = "liver_moylan2013"
prefixRun = "1"
#prefixOut <- "SCT"
prefixOut <- "QC"
```

## load data

```{r}
dt_datExpr <- fread(here("data", "moylan2013.norm.expr.csv.gz"))
```

```{r}
dt_metadata  <- fread(here("data", "moylan2013.metadata.csv"))
```

# QC 

check for duplicate features 

```{r}
table(duplicated(dt_datExpr$gene))
# FALSE  TRUE 
# 22881   640 
```

```{r}
vec_samplenames <- colnames(dt_datExpr)[-1]
```

average gene expression of duplicate genes

```{r}
dt_datExpr <- dt_datExpr[,lapply(.SD,mean),by="gene", .SDcols = vec_samplenames]
```

```{r}
dim(dt_datExpr)
# ] 22881    73
```

```{r}
dt_datExpr[0:5,0:5]
```

check gene names

```{r}
summary(nchar(dt_datExpr$gene))
```

```{r}
dt_datExpr <- dt_datExpr[nchar(dt_datExpr$gene)>0,]
```

```{r}
dim(dt_datExpr)
# ] 22880    73
```

## Outlier check : PCA

```{r}
irlba_out <- irlba(A = as.matrix(dt_datExpr[,-c("gene")]), 
                   nv = 3)
```

Check how the samples look on singular components 1 and 2, coloured by fibrosis

```{r}
df <- as.data.frame(irlba_out$v)
df[["fibrosis"]] <- dt_metadata[["fibrosis"]]
```

```{r}
ggplot(data = df, aes(x=V1, y=V2)) + 
  geom_point(size=5, aes(colour=fibrosis)) 
```

Looks pretty good, no extreme outliers, either

## wrap up 

```{r}
fwrite(dt_datExpr, here("data", "moylan2013.norm.expr.qc.csv.gz"))
```

