---
title: 'Liver - compute gene metamodule embeddings in single cell dataset'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

add cell embeddings on prioritised modules to metadata

# Setup

## Load packages

```{r}
library("dplyr")
library("Matrix")
library("parallel")
library("readr")
library("openxlsx")
library("data.table")
library("here")

## source utility functions

source(here("/perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))


## Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

RNGkind("L'Ecuyer-CMRG")
# ?boot
# For most of the boot methods the resampling is done in the master process, but not if simple = TRUE nor sim = "parametric". In those cases (or where statistic itself uses random numbers), more care is needed if the results need to be reproducible. Resampling is done in the worker processes by censboot(sim = "wierd") and by most of the schemes in tsboot (the exceptions being sim == "fixed" and sim == "geom" with the default ran.gen).
# 
# Where random-number generation is done in the worker processes, the default behaviour is that each worker chooses a separate seed, non-reproducibly. However, with parallel = "multicore" or parallel = "snow" using the default cluster, a second approach is used if RNGkind("L'Ecuyer-CMRG") has been selected. In that approach each worker gets a different subsequence of the RNG stream based on the seed at the time the worker is spawned and so the results will be reproducible if ncpus is unchanged, and for parallel = "multicore" if parallel::mc.reset.stream() is called: see the examples for mclapply. 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

## Set constants


randomSeed = 12345
set.seed(randomSeed)

pAdjustMethod = "BH"
pValThreshold = 0.05#params$pValThreshold

nRep = 10000 # bootstrap replicates
# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf

flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)

# generic project constants 

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
prefixOut <- "SCTint"

prefixRun_WGCNA = "wgcna3"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneWeights = "pkMs"
colGeneNames = "genes"
colModule = "module_meta"
colCellCluster = "cell_cluster_meta"

```

## load data

Seurat object contains expression data and cell + donor-level metadata

```{r}
path_seuratObj <- here("output", paste0(prefixData, "_", prefixRun, "_", prefixOut, "_perslab_labels_seuratObj.RDS.gz"))

seuratObj <- load_obj(path_seuratObj)
```

Load modules dataframe

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz")
head(dt_geneMod_perslab)
```

<!-- ```{r} -->
<!-- dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz") -->
<!-- head(dt_geneMod_gerhard2018) -->
<!-- ``` -->


```{r}
DefaultAssay(seuratObj) <- "SCT_combat"
```

retrieve module embeddings

TODO: Need to translate old to new module names
```{r}
vec_mods_perslab = dt_geneMod_perslab[["module_renamed"]] %>% unique 

vec_mods_perslab <- vec_mods_perslab[nchar(vec_mods_perslab)>0]
```


```{r}

df_scaled_embed = seuratObj@meta.data[,gsub("\\.","-",colnames(seuratObj@meta.data)) %in% vec_mods_perslab]

mat_scaled_embed = df_scaled_embed %>% as.matrix
colnames(mat_scaled_embed) <- gsub("\\.","-",colnames(mat_scaled_embed))
mat_scaled_embed[0:4,0:20]
```


### compute metamodule embeddings in sc data by taking weighted sums of the module embeddings

```{r}
dt_geneMod_perslab[["module_renamed"]] -> vec_mods
names(vec_mods) = dt_geneMod_perslab[[colModule]]
vec_mods <- vec_mods[!is.na(vec_mods) & nchar(vec_mods)>0]
vec_mods <- vec_mods[!duplicated(vec_mods)]
#vec_mods = gsub("-","_", vec_mods)
names(vec_mods) %>% unique -> vec_mods_meta
```

```{r}
mat_embed_meta = sapply(vec_mods_meta, function(mod_meta) {
  
  vec_mods_in_mod_meta = vec_mods[names(vec_mods)==mod_meta]
  # weight by number of genes
  vec_weights = sapply(vec_mods_in_mod_meta, function(each_mod) {sum(dt_geneMod_perslab[["module_renamed"]]==each_mod)})
  vec_weights = vec_weights/sum(vec_weights)
  names(vec_weights) = vec_mods_in_mod_meta
  mat_tmp = mat_scaled_embed[,vec_mods_in_mod_meta]
  if (!is.null(dim(mat_tmp))) {
      # scale the columns of the embeddings matrix by number of genes and take the mean across modules
    mat_tmp = sapply(colnames(mat_tmp), function(each_mod) mat_tmp[,each_mod]*vec_weights[each_mod])
    mat_tmp %>% rowMeans
  } else { # if the metamodule has a single module
    mat_tmp
  }
})

mat_embed_meta[0:3,0:3]


```

```{r}
sd(mat_embed_meta[,1])

```

the metamodules have much lower variance than the modules. 
Scale metamodules to standard deviation 1 just for consistency

```{r}
mat_embed_meta_rescaled <- scale(mat_embed_meta, center=F, scale= apply(mat_embed_meta, 2, sd, na.rm = TRUE)) 
```

## add new embeddings to Seurat object

```{r}
seuratObj <- AddMetaData(object=seuratObj, metadata = data.frame(row.names=rownames(mat_embed_meta_rescaled), mat_embed_meta_rescaled))
```







```{r}
saveRDS(object = seuratObj, file = path_seuratObj, compress = "gzip")
```





