---
title: 'Liver - miscellenaneous analyses'
author: "Jon Thompson, Pers lab, rkm916 at ku dot dk"
date: "`r Sys.time()`" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
# output:
#   html_notebook:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
#   html_document:
#     df_print: paged
#     number_sections: yes
#     toc: yes
#     toc_depth: 3
#     toc_float: yes
---

# Setup

## Load packages

```{r}

# workaround until we fix the project library AND the setting of environment variables
.libPaths(new=c(.libPaths(),"/nfsdata/tools/R/envs/shared/4.0_20201209"))

library("here")
library("data.table")
library("Matrix")
library("parallel")

library("Seurat")#, lib.loc = "/nfsdata/tools/R/envs/rkm916/x86_64-redhat-linux-gnu-library/4.0/")

library("dplyr")
library("ggplot2")
library("ggrepel")
library("igraph")
library("oaqc")
library("tidyr")
library("tidygraph")
library("ggraph")
library("wesanderson")
library("colorspace")
library("patchwork")
library("ggsci")
library("ggpubr")
library("graphlayouts")
library("pheatmap")

library("eulerr")
library("lattice")
# https://github.com/yanwu2014/swne
library("NNLM")
library("swne")

library("Biobase")
# source utility functions
source(here("perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))

# Set options

options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1, 
        verbose=F,
        boot.parallel="multicore", # see ?boot: each worker process inherits the environment of the current session, including the workspace and the loaded namespaces and attached packages (but not the random number seed: see below). 
        boot.ncpus = 20,
        mc.cores=40 # for parallel computation
        ) 

# data.table
setDTthreads(threads = NULL, 
             restore_after_fork = NULL)
getDTthreads()

#options(datatable.WhenJisSymbolThenCallingScope=TRUE)

# Set constants

randomSeed = 12345
set.seed(randomSeed)
npcs = 60
# maxit = 100 # for rlm, see users.stat.umn.edu/~sandy/courses/8053/handouts/robust.pdf
flagDate =substr(gsub("-","",as.character(Sys.Date())),3,1000)


# generic project constants 

prefixData = "liver_perslab_int"
prefixRun = "seurat_7"
dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

colGeneNames = "genes"
colGeneWeights = "pkMs"
```

```{r}
fontSize_plotTitle = 14

fontSize_label_lg = 9
fontSize_label_md = 7
fontSize_label_sm = 4


fontSize_axisTitle_lg = 24
fontSize_axisTitle_md = 20
fontSize_axisTitle_sm = 16

fontSize_axisText_xlg = fontSize_legend_xlg = 20
fontSize_axisText_lg = fontSize_legend_lg = 15
fontSize_axisText_md = fontSize_legend_md = 12
fontSize_axisText_sm = fontSize_legend_sm = 9
fontSize_axisText_xsm = fontSize_legend_xsm = 7


alpha_unicode = "\u03B1"
Beta_unicode = "\u03B2"
gamma_unicode = "\u03B3"
delta_unicode = "\u03B4"



vec_celltype_abbrev = c("Chol", 
                        "DC", 
                        "Endo", 
                        "HSC", 
                        "Stem", 
                        "Hep", 
                        "Mac", 
                        "NK", 
                        paste0(alpha_unicode, Beta_unicode, "T"),#bquote(alpha*Beta~"T"),#expression(paste(alpha, beta, "T")), 
                        paste0(gamma_unicode,delta_unicode,"T"),#bquote(gamma*delta~"T"),#expression(paste(gamma*delta,~"T")),
                        "Bcell",
                        "Plasma")


names(vec_celltype_abbrev) = c("Cholangiocytes","Dendritic-cells", "Endothelial-cells","Hepatic-stellate-cells","Hepatic-stem-cells","Hepatocytes","Macrophages","NK-like-cells",paste0("T-cells-",alpha_unicode, Beta_unicode),paste0("T-cells-",gamma_unicode,delta_unicode),"B-cells-mature","Plasma-cells")
```


## load data

```{r}
seuratObj <- readRDS(here("output","liver_perslab_int_seurat_7_SCTint_perslab_labels_seuratObj.RDS.gz"))
```

```{r}
path_seuratObj_macparland <- here("data", "macparland_seurat_obj3.RDS")

seuratObj_macparland <- load_obj(path_seuratObj_macparland)
```

Load modules dataframes

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod.csv.gz")
#head(dt_geneMod_perslab)

dt_geneMod_gerhard <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna3_geneMod.csv.gz")

```

## Analyses

### check CKB, GLUL, GLS and mir14 in perslab vs macparland

```{r}
seuratObj_macparland$sample_ID <- seuratObj_macparland$Sample
```

```{r}
seuratObj$dataset <- "perslab"
seuratObj_macparland$dataset <- "macparland"
```

```{r}
seuratObj_merged = merge(x=seuratObj, 
                         y=seuratObj_macparland,
                         add.cell.ids = c("perslab", "macparland"), 
                         project = "liver")
```

## Cell-cycle analysis 
See https://satijalab.org/seurat/cell_cycle_vignette.html

```{r}
# list_cellCycleGenes <- lapply(paths_cellCycleGenes, function(x) load_obj(x)[,1])
# names(list_cellCycleGenes) <- names(paths_cellCycleGenes)

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase 
list_cellCycleGenes <- list()
list_cellCycleGenes[["s.genes"]] <- cc.genes$s.genes
list_cellCycleGenes[["g2m.genes"]] <- cc.genes$g2m.genes
```

Compute cell cycle scores using provided genesets

```{r}
# First get rid of zero-expressed genes
# seuratObj <- Seurat::CreateSeuratObject(counts=GetAssayData(seuratObj, slot= "counts"), 
#                                              project= prefixData, 
#                                              assay = "RNA", 
#                                              min.cells = 5, # filtering genes is the only reason to make a new Seurat object
#                                              min.features = 0,
#                                              meta.data = seuratObj@meta.data)

seuratObj_merged <- Seurat::CellCycleScoring(object = seuratObj_merged, 
                               s.features = if (!is.null(list_cellCycleGenes[["s.genes"]])) list_cellCycleGenes[["s.genes"]] else NULL, 
                               g2m.features = if (!is.null(list_cellCycleGenes[["g2m.genes"]])) list_cellCycleGenes[["g2m.genes"]] else NULL, 
                              set.ident = F)


seuratObj_merged$CC.Difference <- seuratObj_merged$S.Score - seuratObj_merged$G2M.Score
```


```{r}
DefaultAssay(seuratObj_merged) = "RNA"
```

```{r}
seuratObj_merged <- Seurat::SCTransform(
  seuratObj_merged, 
  vars.to.regress = c("percent.mito", "CC.Difference"))
```

PCA and UMAP 

```{r}
seuratObj_merged %>% RunPCA(object = ., npcs = npcs, verbose = F, seed.use = randomSeed) %>% RunUMAP(object = ., dims=1:npcs,seed.use=randomSeed) -> seuratObj_merged
```

combine cluster annotation

```{r}
seuratObj_merged$cluster_coarse_merge <- ifelse(seuratObj_merged$dataset=="perslab", seuratObj_merged$cluster_perslab_coarse, seuratObj_merged$Cluster_annot_merged)
```

```{r}
Idents(seuratObj_merged) = "cluster_coarse_merge"
```


```{r}
FeaturePlot(object = seuratObj_merged, 
            features=features,
            cells = colnames(seuratObj_merged)[seuratObj_merged$cluster_coarse_merge=="Hepatocytes"],
            split.by = "dataset") 
```

```{r}
list_subset_idents = list(
  "Hepatocytes"=c("Hepatocytes"),
  "Macrophages"=c("Macrophages"),
  "Hepatic-stellate-cells"=c("Hepatic-stellate-cells","Hepatic_stellate_cells"),
  "Endothelial-cells"=c("Endothelial-cells","Endothelial_cells"),
  "Cholangiocytes"=c("Cholangiocytes")
)
```

```{r}
features = c("CKB", "GLUL", "GLS")#, "mir14")
```

```{r}
A = lapply(list_subset_idents, function(subset_idents) {
  seuratObj_sub = subset(seuratObj_merged, idents = subset_idents)
  try(DE_MAST_RE_seurat(
    object=seuratObj_sub,
    random_effect.var="sample_ID",
    ident.1="perslab",
    ident.2 = "macparland",
    group.by = "dataset",
    logfc.threshold = 0.01,
    assay=NULL,
    slot="data",
    features = features,
    min.pct = 0.01,
    max.cells.per.ident = 5000,
    random.seed = randomSeed,
    latent.vars = NULL,
    n_cores=NULL,
    verbose=TRUE))
})

```

```{r}
df_features = list_df_DE[[5]][rownames(list_df_DE[[5]]) %in% features,]
```


```{r}
list_df_DE_2 = lapply(names(list_subset_idents), function(key) {
  subset_idents = list_subset_idents[[key]]
  seuratObj_sub = subset(seuratObj_merged, idents = subset_idents)
  df_out = try(Seurat::FindMarkers(
      object=seuratObj_sub,
      test.use="MAST",
      ident.1="perslab",
      ident.2 = "macparland",
      group.by = "dataset",
      logfc.threshold = 0.01,
      assay=NULL,
      slot="data",
      features = features,#if (key!="Hepatocytes") features[1:3] else features[c(1,3)],
      min.pct = 0.01,
      max.cells.per.ident = 5000,
      random.seed = randomSeed,
      latent.vars = NULL,
      verbose=TRUE)) 
  # list_df = Filter(f= function(df) class(df)=="data.frame", x=list_df)
  # df_out = Reduce(f=rbind, x=list_df)
  if (class(df_out)=="data.frame") df_out$cluster = key
  return(df_out)
})

names(list_df_DE_2) = names(list_subset_idents)
```



```{r}
df_DE_2 = Reduce(rbind,list_df_DE_2[3:5])
df_DE_2$p_val_adj <- p.adjust(df_DE_2$p_val, method="fdr")
df_DE_2 <- df_DE_2[order(df_DE_2$p_val, -df_DE_2$avg_logFC),]

df_DE_2 = data.frame("gene"=rownames(df_DE_2),df_DE_2)
```

```{r}
openxlsx::write.xlsx(df_DE_2, file = here("output", paste0(prefixData,"_", prefixRun, "_DEgenes_perslab_macparland_", flagDate, ".xlsx")))
```

Are macrophage modules specific to particular patients?

```{r}
VlnPlot(seuratObj, features = paste0("Macrophages_",1:4), group.by = "sample_ID")
```

why is M-Mac-2 specific to a subset of patients?

* How is module activity computed?
* M-Mac-2 corresponds more or less to a macrophage population. How specific is it?

```{r}
Idents(seuratObj) <- "cluster_perslab_coarse"
seuratObj_macs <- subset(seuratObj, idents="Macrophages")
```

```{r}
VlnPlot(seuratObj_macs, group.by = "cluster_perslab", features = paste0("Macrophages_",1:4))
```

Top Reactome terms for M-Mac-2:
Respiratory electron transport
Respiratory electron transport, ATP synthesis by chemiosmotic coupling, and heat production by uncoupling proteins.

Top KEGG terms for M-Mac-2: Oxidative phosphorylation

```{r}
VlnPlot(seuratObj, features = c("nCount_RNA", "nFeature_RNA", "percent.mito", "percent.ribo"), group.by = "sample_ID")
```

```{r}
dt_covars_tech_macs = seuratObj_macs@meta.data[,c("Macrophages_2", "nCount_RNA", "nFeature_RNA", "percent.mito", "percent.ribo", "CC.Difference")]

cor(dt_covars_tech_macs)
```

```{r}
dt_covars_patient_macs = seuratObj_macs@meta.data[,c("Macrophages_2", colnames(seuratObj_macs@meta.data)[c(39:43,45:46)])]

mat_covars_patient_macs = apply(dt_covars_patient_macs,2,as.numeric)

dt_covars_patient_macs_numeric <- data.table(mat_covars_patient_macs)
dt_covars_patient_macs_numeric$sample_ID <- seuratObj_macs$sample_ID
dt_covars_patient_macs_numeric_mean <- dt_covars_patient_macs_numeric[,lapply(.SD, mean), by=sample_ID, .SDcols=colnames(mat_covars_patient_macs)]
```

```{r}
cor(dt_covars_patient_macs_numeric_mean[,2:9])
```

genes to stain for as a complement to the MS (probably for a revision). 
check the expression of the individual HLA genes in Gerhard 
and in our macrophage subclusters (compared to MacParland)?:

Gerhard 2018 bulk expression data 

```{r}
dt_datExpr <- fread(here("data", "liver_gerhard2018_norm.counts.csv.gz"))

mat_datExpr <- as.matrix(dt_datExpr[,-1])
rownames(mat_datExpr) = dt_datExpr$gene
```

### patient metadata

```{r}
dt_metadata <- fread(here("data", "liver_gerhard2018_patient_metadata_qc.csv.gz"))
```

```{r}
vec_genes_HLA = c("HLA-DRA", "HLA-DQA2", "HLA-DQB2")
```

```{r}
all.equal(colnames(mat_datExpr), dt_metadata$run_accession)
# [1] TRUE
```


```{r}
mat_datExpr[vec_genes_HLA,] %>% t %>% scale(., center=FALSE, scale=apply(.,2,sd,na.rm=T)) -> mat_datExpr_scaled_sub_t
```

```{r}
dt_comb = cbind(dt_metadata, mat_datExpr_scaled_sub_t)
```

```{r}
colnames(dt_comb) = gsub("-","_",colnames(dt_comb))
```
```{r}
library(colorspace)
```

boxplots

```{r}
vec_conditions = c("STEATOSIS", "Fibrosis", "Lob Inflam")
vec_colors_plot = wes_palette("Darjeeling1", n =length(vec_conditions))
names(vec_colors_plot) = vec_conditions

list_row_boxplot = lapply(vec_conditions, function(condition) {
  dt_boxplot <- dt_comb[,.(DCL.Patient.ID,Diagnosis,HLA_DRA,HLA_DQA2,HLA_DQB2)]
  filter_condition = quote(grepl(paste0(condition,"|NORMAL"), dt_boxplot$Diagnosis))
  dt_boxplot <- dt_boxplot[eval(filter_condition),]
  #dt_boxplot <- dt_boxplot[!grepl("Fibrosis 3/4", Diagnosis)]
  dt_boxplot <- melt(dt_boxplot,id.vars = c("DCL.Patient.ID", "Diagnosis"))
  dt_boxplot$Diagnosis <- ifelse(dt_boxplot$Diagnosis=="NORMAL", "NORMAL", condition)
  colnames(dt_boxplot)[c(3,4)] <- c("gene","expression")
  # dt_boxplot$Diagnosis = factor(dt_boxplot$Diagnosis, levels = c("NORMAL","Fibrosis 3","Fibrosis 3/4","Fibrosis 4"), ordered=TRUE)
  dt_boxplot$Diagnosis = factor(dt_boxplot$Diagnosis, levels = c("NORMAL",condition), ordered=TRUE)
  
  vec_colors_plot_condition = rep(vec_colors_plot[condition], length(unique(dt_boxplot$Diagnosis)))
    
  names(vec_colors_plot_condition) = levels(dt_boxplot$Diagnosis)
    
  list_plots = lapply(1:length(vec_genes_HLA), function(i) {
    gene_plot = gsub("-","_",vec_genes_HLA)[i]
    #j must be provided when with=FALSE 
    dt_boxplot_sub = dt_boxplot[gene==gene_plot,colnames(dt_boxplot),with=F,]
    
    
    p_box = ggplot(
      data=dt_boxplot_sub, mapping=aes(x=Diagnosis, y=expression, fill=Diagnosis)) +
      geom_boxplot()  +
      scale_fill_manual(values=vec_colors_plot_condition) +
      geom_jitter(width=0.1,alpha=0.5) +
      ggtitle(vec_genes_HLA[i]) +
      theme(
          legend.position = "none",
          axis.text.x = element_text(size = fontSize_axisText_sm, angle=90), 
          axis.text.y = element_text(size = fontSize_axisText_sm), 
          axis.line = element_blank(), 
          axis.ticks = element_blank(), 
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = fontSize_axisTitle_sm),
          panel.background = element_blank(), 
          plot.background = element_blank(), 
          panel.grid = element_blank(),
          legend.title=element_blank(),
          legend.text = element_text(size=fontSize_legend_lg),
          aspect.ratio = 1.5) + 
      labs(y="gene expression (scaled)")
    
  })
  patchwork::wrap_plots(... = list_plots, 
                            ncol = length(list_plots)
                            )
})

grid_boxplot = patchwork::wrap_plots(... = list_row_boxplot, 
                            nrow = length(list_row_boxplot)
                            )

ggsave(plot = grid_boxplot, filename = paste0(dirWGCNA_plots,prefixData,"_", prefixRun,  "_HLAgenes_gerhard_boxplots_", params$date,".pdf"), height=15, width =15)
```

# prep for regression 
```{r}
list_dt_comb = list()
list_dt_comb[["steatosis"]] = dt_comb[Diagnosis %in% c("NORMAL", "STEATOSIS 2", "STEATOSIS 3"),]
list_dt_comb[["steatosis"]][,condition:=ifelse(Diagnosis=="NORMAL",0,1)]
list_dt_comb[["steatosis"]][Diagnosis=="STEATOSIS 3","condition"] <- 2

list_dt_comb[["lobinflam"]] = dt_comb[Diagnosis %in% c("NORMAL","Lob Inflam 1", "Lob Inflam 2"),]
list_dt_comb[["lobinflam"]][,condition:=ifelse(Diagnosis=="NORMAL",0,1)]
list_dt_comb[["lobinflam"]][Diagnosis=="Lob Inflam 2","condition"] <- 2

list_dt_comb[["fibrosis"]] = dt_comb[Diagnosis %in% c("NORMAL","Fibrosis 3"),]
list_dt_comb[["fibrosis"]][,condition:=ifelse(Diagnosis=="Fibrosis 3",1,0),]
# binary, so factor. NB: This will change the lm variable for case  to condition1
list_dt_comb[["fibrosis"]]$condition <- as.factor(dt_comb_fibrosis$condition)

```

```{r}
# https://stats.stackexchange.com/questions/242404/how-does-the-boot-package-in-r-handle-collecting-bootstrap-samples-if-strata-are
set.seed(randomSeed)
R=9999

system.time(list_list_bootOut <- lapply(names(list_dt_comb), function(condition) {
  
  print(paste0("condition: ", condition))
  
  dt_model_tmp <- setDF(list_dt_comb[[condition]])

  lapply(gsub("-","_",vec_genes_HLA), function(gene_HLA) {
    
    print(paste0("gene_HLA: ", gene_HLA))
    
    fnc_stat = function(data,indices) {
         lm(
           data = data[indices,], 
           formula = formula(paste0(gene_HLA, "~ SEX + BMI_surg + Age + condition"))) %>% coefficients()
    }
    
    bootOut = "error"
    class(bootOut) = "try-error"
    while (class(bootOut)=="try-error") {
      boot_out = try({
        set.seed(randomSeed+1)
        bootOut <- boot::boot(data=dt_model_tmp, 
                               statistic = fnc_stat,
                               strata = as.factor(dt_model_tmp$Diagnosis),
                               R = R,
                               sim = "ordinary",
                               stype = "i",
                               parallel = getOption("boot.parallel"),
                               ncpus = getOption("boot.ncpus")
                              ) 
      })
    }
    return(bootOut)  
    
  }) -> list_bootOut 
  names(list_bootOut) = vec_genes_HLA
  return(list_bootOut)
  })
)

names(list_list_bootOut) <- names(list_dt_comb)
```

extract coefficient tables 

get p-value, std error and confidence intervals from the $t

```{r}
list_dt_bootOut_stats <- lapply(names(list_list_bootOut), function(condition) {
  list_bootOut = list_list_bootOut[[condition]]
  mat_bootOut_stats <- t(sapply(names(list_bootOut), function(gene_HLA) {
    bootOut = list_bootOut[[gene_HLA]]
    vec_out = tryCatch({
      mat_boot_replicates = bootOut$t
      colnames(mat_boot_replicates) = names(bootOut$t0)
      idx_col = grep("condition",names(bootOut$t0))
      # compute bootstrap statistics
      bootBias = mean(mat_boot_replicates[,idx_col]) - bootOut$t0[idx_col]
      bootMean = mean(mat_boot_replicates[,idx_col]) 
      bootSE = sd(mat_boot_replicates[,idx_col])
      bootMed = median(mat_boot_replicates[,idx_col])
      
      # compute bootstrap percentile intervals 
      bootciOut = boot.ci(bootOut, index = 5)
      # idx_lower = round((R+1)*0.025,0)
      # idx_upper = round((R+1)*0.975,0)
      # confInt_lower_boot = sort(mat_boot_replicates[,idx_col])[idx_lower]
      # confInt_upper_boot = sort(mat_boot_replicates[,idx_col])[idx_upper]
      confInt_lower_percentile = bootciOut$percent[4]
      confInt_upper_percentile = bootciOut$percent[5]
      # compute Normal-theory intervals
      # confInt_lower_norm = qnorm(p = 0.025, mean = bootMean, sd = bootSE, lower.tail = T)
      # confInt_upper_norm = qnorm(p = 0.975, mean = bootMean, sd = bootSE, lower.tail = T)
      confInt_lower_norm = bootciOut$normal[2]
      confInt_upper_norm = bootciOut$normal[3]
      
      c("estimate_original"=bootOut$t0[idx_col],
       #"bootBias"=bootBias,
       "bootSE"=bootSE,
       "estimate_bootMean"=bootMean,
       "estimate_bootMed" = bootMed,
       "confInt_lower_percentile"=confInt_lower_percentile,
       "confInt_upper_percentile"=confInt_upper_percentile,
       "confInt_lower_norm"=confInt_lower_norm,
       "confInt_upper_norm"=confInt_upper_norm,
       "p.value_percentile" = ifelse(bootOut$t0[idx_col]>0,  max(1,sum(mat_boot_replicates[,idx_col]<=0))/R, max(1, sum(mat_boot_replicates[,idx_col]>=0))/R),
       "p.value_norm" = pnorm(q=abs(bootMean), mean=0, sd=bootSE, lower.tail = F))
    }, error = function(err) {
      warning(paste0(gene_HLA, ": ", err))
      c("estimate_original"=NA_real_,
       #"bootBias"=NA_real_,
       "bootSE"=NA_real_,
       "estimate_bootMean"=NA_real_,
       "estimate_bootMed" = NA_real_,
       "confInt_lower_percentile"=NA_real_,
       "confInt_upper_percentile"=NA_real_,
       "confInt_lower_norm"=NA_real_,
       "confInt_upper_norm"=NA_real_,
        "p.value_percentile" = NA_real_,
       "p.value_norm"=NA_real_)
      })
    
  }))
  dt_bootOut_stats = data.table("condition"=condition, 
                                "gene_HLA"=names(list_bootOut), 
                                mat_bootOut_stats)
  
  #dt_bootOut_stats[,p.value_percent_BH:=p.adjust(p.value_percent,method = params$pAdjMethod)]
  #dt_bootOut_stats[,p.value_born_BH:=p.adjust(p.value_norm, method = params$pAdjMethod)]
  colnames(dt_bootOut_stats) = gsub("\\.condition\\d*$", "", colnames(dt_bootOut_stats)) # for some reason, it appends ".condition" to these column names
  return(dt_bootOut_stats)
})

names(list_dt_bootOut_stats) <- names(list_dt_model_sub)

dt_bootOut_stats = data.table::rbindlist(list_dt_bootOut_stats)

#dt_bootOut_stats[,gene_HLA := gsub("_","-", gene_HLA)]
```


## Check for PPI enrichment in figure 3 heatmap overlapping module genes

```{r}
dt_geneMod_perslab$module_renamed %>% unique -> vec_mods

vec_mods <- vec_mods[!is.na(vec_mods) & nchar(vec_mods)>0]

vec_mods_ribo_overlap = c(
  "Endothelial-cells_2", 
  "T-cells-gamma-delta_1", 
  "Dendritic-cells_2",
  "Hepatic-stem-cells_1",
  "Plasma-cells_1",
  "B-cells-mature_1",
  "Macrophages_4",
  "T-cells-alpha-beta_1",
  "Cholangiocytes_2",
  "Hepatic-stellate-cells_3"
  )

all(vec_mods_ribo_overlap %in% vec_mods)
```

```{r}
list_genelists_ribo_overlap = lapply(vec_mods_ribo_overlap, function(mod) {
  condition = quote(dt_geneMod_perslab$module_renamed==mod)
  vec_mod_genes = dt_geneMod_perslab[eval(condition), ..colGeneNames, ][[1]] %>% unique
  # filter out moduke genes absent in expression data
  #vec_idxRow <- match(vec_mod_genes, dt_datExpr[,gene])
  #vec_mod_genes <- vec_mod_genes[!is.na(vec_idxRow)]
  return(vec_mod_genes)
})

names(list_genelists_ribo_overlap) = vec_mods_ribo_overlap
```


```{r}
  
for (i in 1:4) {
 
names(list_genelists_ribo_overlap)=   gsub(pattern=c("alpha-", "beta", "gamma-", "delta")[i], 
  replacement=c(alpha_unicode, Beta_unicode, gamma_unicode, delta_unicode)[i],
  x=names(list_genelists_ribo_overlap))

}

vec_mod_number = gsub("\\D","",names(list_genelists_ribo_overlap))
names(list_genelists_ribo_overlap) = paste0("M-", vec_celltype_abbrev[match(gsub("_\\d$","",names(list_genelists_ribo_overlap)), names(vec_celltype_abbrev))], "-", vec_mod_number)
```

```{r}
vec_genes_all = unlist(list_genelists_ribo_overlap, use.names=F) %>% unique
```

```{r}
vec_gene_module_membership = sapply(vec_genes_all, function(gene) sapply(list_genelists_ribo_overlap, function(genelist) {gene %in% genelist}) %>% sum)  %>% sort(.,decreasing = T)

vec_genes_most_shared = names(vec_gene_module_membership)[vec_gene_module_membership>5]

vec_genes_most_shared %>% kength
# [1] 77
```

PPI analysis

```{r}
library("STRINGdb")
```

```{r}
data_organism = "hsapiens"
```

```{r}
if (data_organism == "hsapiens") STRINGdb_species <- 9606 else if (data_organism == "mmusculus") STRINGdb_species <- 10090
```

```{r}
ppi <- data.frame(gene = vec_genes_most_shared) # extract the genes with the corresponding color to dataframe
modulePPIrow <- c('p.value' = 1, 'expected.interactions' = NA)

# generate a STRINGdb object instance
string_db <- STRINGdb$new(version="11",
                          species = STRINGdb_species,
                          score_threshold=200, #threshold for the combined scores of the interactions, such that any interaction below that threshold is not loaded in the object (by default the score threshold is set to 400).
                          input_directory="")

example1_mapped <- string_db$map(ppi, 'gene', removeUnmappedRows = TRUE ) # Check the dataframe genes'

hits <- example1_mapped$STRING_id

string_db$plot_network( hits )

modulePPIrow['p.value'] = string_db$get_ppi_enrichment(hits)$enrichment

modulePPIrow['expected.interactions'] = string_db$get_ppi_enrichment(hits)$lambda

```

do GO analysis

```{r}
enrichment <- string_db$get_enrichment( hits )
head(enrichment, n=20)
```

```{r}
openxlsx::write.xlsx(enrichment, here("output", paste0(prefixData,"_", prefixRun,  "_ribo_PPI_GO_results_", params$date,".xlsx")))
```

