---
title: 'Liver module gene ontology analysis (using gProfiler)'
author: "Jon Thompson, Pers lab"
date: "`r Sys.time()`" 
params: 
  date: !r substr(gsub("-","",as.character(Sys.Date())),3,1000)
  randomSeed: !r 12345
  pValThreshold: !r 0.05
  padjMethod: "bonferroni"
output:
  html_notebook: 
  df_print: paged
number_sections: yes
toc: yes
toc_depth: 3
toc_float: yes
html_document:
  df_print: paged
---

Sources
https://www.bioconductor.org/packages/release/bioc/vignettes/topGO/inst/doc/topGO.pdf
http://bioconductor.org/packages/release/BiocViews.html#___OrgDb
GOSemSim: https://bioconductor.org/packages/release/bioc/vignettes/GOSemSim/inst/doc/GOSemSim.html

## Setup

## Set options


```{r}
options(stringsAsFactors = F, 
        use="pairwise.complete.obs", 
        warn=1)
```
## source functions

```{r}
source(file=here("perslab-sc-library", "utility_functions.R"))
source(file=here("perslab-sc-library","functions_sc.R"))
```

## Load packages

```{r}

library("dplyr")
library("Matrix")
library("parallel")
library("readr")
library("openxlsx")
library("data.table")
library("here")
library("gprofiler2")
# ipak(c("dplyr", "ggplot2", "Matrix", "parallel", "RColorBrewer", "readr", "pheatmap", "GOSemSim", "AnnotationHub","org.Hs.eg.db", "GSEABase"))#,"ComplexHeatmap", "liger", "WGCNA", "circlize", "xlsx"))
```

### constants 

```{r}
prefixData <- "liver_perslab_int"
prefixRun <-  "seurat_7"

colGeneNames = "genes"
colGeneWeights ="pkMs"
colModule = "module_assoc"
colCellClust = "cell_cluster_assoc"

dirWGCNA_outs = "/projects/jonatan/pub-perslab/18-liver-wgcna/"
dirWGCNA_plots = paste0(dirWGCNA_outs, "plots/" )
dirWGCNA_RObjects = paste0(dirWGCNA_outs, "RObjects/" )
dirWGCNA_tables = paste0(dirWGCNA_outs, "tables/" )

```

```{r}
params = list("randomSeed"=12345, 
              "date"=substr(gsub("-","",as.character(Sys.Date())),3,1000), 
              "nrepBootstrap"=1e4, 
              "pValThreshold" =0.05)
padjMethod ="bonferroni"
set.seed(params$randomSeed)
nrepBootstrap = 10000
```

## load data

Load modules dataframe

```{r}
dt_geneMod_perslab <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_perslab_int_wgcna3_geneMod_merged_200325.csv.gz")
head(dt_geneMod_perslab)
```

```{r}
dt_geneMod_gerhard2018 <- fread("/projects/jonatan/pub-perslab/18-liver-wgcna/tables/liver_gerhard2018_wgcna2_geneMod_merged.csv.gz")
head(dt_geneMod_gerhard2018)
```


## Analysis

prepare lists of vectors of module gene weights

```{r}
vec_mods_perslab <- unique(na.omit(dt_geneMod_perslab$module_assoc))
vec_mods_perslab <- vec_mods_perslab[nchar(vec_mods_perslab)>0]
vec_mods_perslab
```


```{r}
list_pkMs_perslab <- lapply(vec_mods_perslab, function(module) {
  vecOut<-dt_geneMod_perslab[[colGeneWeights]][dt_geneMod_perslab[[colModule]]==module]
  names(vecOut) <- dt_geneMod_perslab[["ensembl"]][dt_geneMod_perslab[[colModule]]==module]
  vecOut <- vecOut[!is.na(names(vecOut))]
  vecOut <- sort(vecOut, decreasing = T)
  vecOut
})
names(list_pkMs_perslab)<-vec_mods_perslab
head(list_pkMs_perslab[[1]])
```

```{r}
vec_mods_gerhard2018 <- unique(na.omit(dt_geneMod_gerhard2018$module_assoc))
vec_mods_gerhard2018 <- vec_mods_gerhard2018[nchar(vec_mods_gerhard2018)>0]
vec_mods_gerhard2018
```


```{r}
list_pkMs_gerhard2018 <- lapply(vec_mods_gerhard2018, function(module) {
  vecOut<-dt_geneMod_gerhard2018[[colGeneWeights]][dt_geneMod_gerhard2018[[colModule]]==module]
  names(vecOut) <- dt_geneMod_gerhard2018[["ensembl"]][dt_geneMod_gerhard2018[[colModule]]==module]
  vecOut <- sort(vecOut, decreasing = T)
  vecOut <- vecOut[!is.na(names(vecOut)) & nchar(names(vecOut))>0]
  vecOut
})
names(list_pkMs_gerhard2018)<-vec_mods_gerhard2018
head(list_pkMs_gerhard2018[[1]])
```

```{r}
list_pkMs_comb <- append(list_pkMs_perslab, list_pkMs_gerhard2018)
head(list_pkMs_comb[[1]])
head(names(list_pkMs_comb))
tail(names(list_pkMs_comb))
```

### run gprofiler 

run together since gprofiler2 maps to ensembl anyway

combine genes in both experiments as background

```{r}

path_genesBackground_perslab_ensembl <- here("data", "liver_perslab_genesBackground_ensembl.RDS.gz")

if (!file.exists(path_genesBackground_perslab_ensembl)) {
  dt_datExpr_perslab <- fread(here("output","liver_perslab_int_seurat_7_datExpr_combat_ensembl.csv.gz"))
  vec_genesBackground_perslab <- dt_datExpr_perslab$ensembl_gene_id
  saveRDS(vec_genesBackground_perslab, path_genesBackground_perslab_ensembl, compress="gzip")
  rm(dt_datExpr_perslab)
} else {
  vec_genesBackground_perslab_ensembl <- readRDS(path_genesBackground_perslab_ensembl)
}
```

```{r}
path_genesBackground_gerhard2018 <- here("data", "liver_gerhard2018_genesBackground.RDS.gz")

if (!file.exists(path_genesBackground_gerhard2018)) {
  dt_datExpr_gerhard2018 <- fread(here("data","liver_gerhard2018_norm.counts.csv.gz"))
  vec_genesBackground_gerhard2018 <- dt_datExpr_gerhard2018$gene
  saveRDS(vec_genesBackground_gerhard2018, path_genesBackground_gerhard2018, compress="gzip")
rm(dt_datExpr_gerhard2018)
} else {
  vec_genesBackground_gerhard2018 <- readRDS(path_genesBackground_gerhard2018)
}
```

```{r}
vec_genesBackground_comb <- union(vec_genesBackground_perslab_ensembl, vec_genesBackground_gerhard2018)
```

```{r}
length(vec_genesBackground_comb)
```

run gprofiler 
* with bonferroni
* return all results including not significant

```{r}
list_gost <- gprofiler2::gost(query=lapply(list_pkMs_comb, names),
                                 organism = "hsapiens", 
                                 ordered_query = T, 
                                 significant=T,
                                 exclude_iea = T, 
                                 measure_underrepresentation= FALSE,
                                 correction_method= "g_SCS",
                                 user_threshold=params$pValThreshold,
                                 custom_bg = vec_genesBackground_comb,
                                 sources=c("GO:BP", "GO:CC","GO:MF", "KEGG", "REACTOME", "CORUM", "HP", "WP"))
```

add cell cluster to results

```{r}
dt_out <- data.table("cell_cluster"=ifelse(grepl("perslab",list_gost$result$query),dt_geneMod_perslab[[colCellClust]][match(list_gost$result$query, dt_geneMod_perslab[[colModule]])], dt_geneMod_gerhard2018[[colCellClust]][match(list_gost$result$query, dt_geneMod_gerhard2018[[colModule]])]), list_gost$result)
```


## write out outputs

```{r}
fwrite(dt_out, paste0(dirWGCNA_tables, prefixData,"_",prefixRun,"_", params$date, "_sc_bulk_modules_GOgprofiler2_full.csv"))
```

```{r}
openxlsx::write.xlsx(x=dt_out, file = paste0(dirWGCNA_tables, prefixData,"_",prefixRun,"_",params$date,"_sc_bulk_modules_GOgprofiler2_full.xlsx"))
```
